### 2022-01-21 scr/gbm_proteomics_data_analysis_proteolytic_products_prep_figures_v3_gh_format.Rmd
first commit

---
title: "GBM Proteomics and Proteolytic processing data analysis - Reproducible report"
author: "Miguel Cosenza"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document
always_allow_html: true
---

```{r}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

```{r results='hide'}
## Required packages ----
library(tidyverse)
library(mixOmics)
library(fs)
library(kableExtra)
library(sva)
library(limma)
library(naniar)
library(missForest)
library(clusterProfiler)
library(ReactomePA)
library(org.Hs.eg.db)
library(DT)
library(here)
library(janitor)
library(drawProteins)
library(seqinr)
library(ggpubr)

source(here("scr/helper_functions.R"))

theme_set(theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 6, 
                                           angle = 90),
                axis.text.y = element_text(hjust = 0.5, vjust = 0, size = 6),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", fill=NA, size=1.5),
                axis.title=element_text(size=8),
                legend.text = element_text(size = 7),
                legend.title = element_text(size = 8),
                legend.key.height= unit(3, 'mm'),
                legend.key.width= unit(3, 'mm'),
                legend.position="bottom"))
```

The following report holds the code, results and interpretation of the proteomics data analysis of Initial-Recurrent glioblastoma samples of matched patients.

# Sample annotation info

```{r echo=TRUE}
sample_annotation <- read_csv(here("data/sample_annotation.csv"))

# correct annotation

sample_annotation2 <- sample_annotation %>%
  mutate(patient = paste("x",patient, sep = ""),
         recurrence = case_when(recurrence == "initial" ~ "prim",
                                recurrence == "recurrent" ~ "rec",
                                TRUE ~ recurrence)) %>%
  mutate(paired_id = paste(patient, recurrence, sep = "_")) %>%
  filter(recurrence %in% c("prim", "rec"))

sample_annotationwo6 <- sample_annotation2 %>%
  filter(patient != "x6")
```

## Sample annotation

### Brief note on sample annotation:

Twenty-two patient matched samples (corresponding to 11 patients) were processed for protein extraction, TMT-labelled and submitted via high-pH fractionation and further analyzed via LC-MS/MS.

```{r echo=FALSE, error=TRUE}
kbl(sample_annotation, booktabs = TRUE) %>%
  kable_styling(font_size = 14, bootstrap_options = c("striped", "hover"))
```

# Brief note on data analysis and processing:

Three proteomics data analysis approaches were used to explore the spectral data and will be briefly described:

1. General large scale proteomics.
2. Proteolytic processing.
3. Proteogenomics

The FragPipe bioinformatic pipeline was used for each of these approaches with varying parameters depending on the type of explorative approach.

__1. General large scale proteomics analyses__ : In order to explore the differential abundance of proteins between Recurrent and Initial tumor, a fully-tryptic search was performed against the EBI Human canonical proteome (version 2021_03), appended with common contaminants and iRT peptides. A minimum percentage of  %5 of the total summed reporter ion intensities were required to consider a peptide and reporter ion for quantitation.

__2. Proteolytic processing__: The large scale evaluation of differential proteolytic activity between recurrent vs primary tumor was executed via a semi-tryptic search against the EBI Human canonical proteome (version 2021_03), appended with common contaminants and iRT peptides.  A minimum percentage of  %5 of the total summed reporter ion intensities were required to consider a peptide and reporter ion for quantitation.

__3. Proteogenomics__: The explorative proteogenomics analysis started with the generation of a GBM-specific database from publicly available RNA-seq sequencing data from tumor specimens. Fastq files were obtain from the SRA BioProject PRJNA627121 (Cell Rep, 2021 Mar 2;34(9):108787). Eight samples containing paired reads were selected, based on: 1) the higher number of reads per file; 2) having a similar proportion of males and females and 3) not bearing isocitrate-dehyderogenase (IDH)-1 mutation (IDH WT). None of these patients undergo previous treatment, therefore potential variants in the sequencing data are not expected to be associated to treatment-related mutagenic processes. Spectral data was then searched against this custom database using a fully-tryptic approac. No minimum intensity threshold was set for reporter ion intensity for quantitation.

Exceptionally when stated differently within each sub-approach described above, all these searches were executed using the following parameters: MSFragger was used as a search engine, setting a fully or semi-tryptic specificity allowing for 1 missed cleavages. Precursor mass tolerance was set to -20/20 and fragment mass tolerance of 20 ppm with mass calibration and parameter optimization resulting in 5 ppm. Peptide N-terminal acetylation and peptide N-terminal TMT labeling were set as variable modifications. TMT labelling at K and carbamidomethylation of C were set as fixed modifications. MSBooster was used for deep-learning based predictions of retention time and spectra. Predicted features were used by Percolator for post-processing scoring and false discovery rate (FDR) control via target-decoy competition of all peptide-to-spectrum matches (PSMs) obtained from MSFragger search. ProteinProphet was used as a protein inference algorithm keeping all PSMs with probability score bigger than 0.9. Final report and FDR estimation was based on the filtered PSM and protein lists before protein scoring.

Relative quantitation of identified peptides within each sample was performed via their reported ion intensities using TMT-integrator. Only PSMs coming from unique peptides with a minimum probability of 0.9 and an isotopic purity of at least 50% were considered for quantitation. Quantitative values were normalized via median centering and summarized by protein/gene and peptide using a virtual reference channel, to finally generate protein and peptide normalized abundance matrices of abundance used for further statistical processing in combination with the sample annotation.

# General proteomics analysis

## Initial data loading and wrangling

Load id/quant data and prep coverage information

```{r message=FALSE}
## tmt-integrator output loading 
tmt_protdata_frag <- read_tsv(here("data/specific_search_fragpipe17/specific_no_ptms_2/tmt-report/abundance_protein_MD.tsv")) %>% 
  janitor::clean_names()

identified_proteins <- read_tsv(here("data/specific_search_fragpipe17/specific_no_ptms_2/combined_protein.tsv")) %>% 
  janitor::clean_names()

identified_peptides <- read_tsv(here("data/specific_search_fragpipe17/specific_no_ptms_2/combined_peptide.tsv")) %>%
  janitor::clean_names()

fasta_ident_prots <- read.fasta(here("data/specific_search_fragpipe17/specific_no_ptms_2/mix_2/protein.fas"), seqtype = "AA", as.string = TRUE)

```

Wrangle and merge protein ID/quant with sample annotation

```{r}
# wide expression matrix with NAs
expr_matrix <- dplyr::select(tmt_protdata_frag,
                             Protein = index, starts_with("x"))

long_mat <- pivot_longer(expr_matrix,
                         cols = starts_with("x"),
                         names_sep = "_",
                         names_to = c("patient", "recurrence"),
                         values_to = "Abundance")

long_mat_2 <- long_mat %>%
  mutate(paired_id = paste(patient, recurrence, sep = "_"))

paired_annot_uniq <- long_mat_2 %>%
  dplyr::select(patient, recurrence, paired_id) %>%
  distinct()

# correct annotation

sample_annotation2 <- sample_annotation %>%
  mutate(patient = paste("x",patient, sep = ""),
         recurrence = case_when(recurrence == "initial" ~ "prim",
                                recurrence == "recurrent" ~ "rec",
                                TRUE ~ recurrence)) %>%
  mutate(paired_id = paste(patient, recurrence, sep = "_")) %>%
  filter(recurrence %in% c("prim", "rec"))

## Annotated abundance data in long format ----

quant_annotated <- left_join(long_mat_2,
                             sample_annotation2, 
                             by = c("paired_id", "patient","recurrence")) %>%
  mutate(Channel_mix = paste(mixture,channel,  sep = "_"))

## Wide xpression matrices without NAs
expr_matnona <- column_to_rownames(expr_matrix,
                            "Protein") %>%
          as.matrix() %>%
          na.omit()
```

## ID coverage

### Coverage plot

```{r}
ids_summary <- quant_annotated %>% 
  dplyr::select(Protein, mixture, Abundance) %>% 
  distinct(Protein, mixture, .keep_all = TRUE) %>%
  group_by(mixture) %>%
  summarise(`Nr of IDs` = sum(!is.na(Abundance))) %>% 
  bind_rows(.,
            tibble(mixture = "In all three",
                   `Nr of IDs` = nrow(expr_matnona))) %>%
  bind_rows(.,
            tibble(mixture = "Overall",
                   `Nr of IDs` = nrow(expr_matrix)))
```

* Number of identified proteins: `r nrow(expr_matrix)`

* Number of identified and quantified proteins in all samples: `r nrow(expr_matnona)`

The initial analysis with MaxQuant yielded 2090 IDs in all samples (without missing values) and 5176 overall.

```{r fig.height=4, fig.width=7}
ggplot(data = ids_summary,
       aes(y = `Nr of IDs`, x = mixture)) +
  geom_col(fill = "#008683") + 
  geom_text(size = 5, position = position_stack(vjust = 0.5), aes(label = `Nr of IDs`)) +
  coord_cartesian(ylim = c(0, 5000))+
  coord_flip() +
  labs(title = "Number of Identified and quantified Proteins by Mixture") +
  ylab("Nr of IDed and Quant Proteins") + 
  xlab("TMT Mixture") + 
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0.1, size = 10, angle = 0),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.1, size = 10, angle = 0),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey"),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5),
        axis.title=element_text(size=12,face="bold"))
```  
## Exploratory analysis and visualizations

### Visualization of missing values

```{r fig.height=7, fig.width=8}
vis_miss(expr_matrix) + 
          ggtitle("Distribution of missing values") + 
          theme(axis.text.x = element_text(hjust = 0.5, vjust = 0.1, size = 10, angle = 90),
                axis.text.y = element_text(hjust = 0.5, vjust = 0.1, size = 10),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", fill=NA, size=1.5),
                axis.title=element_text(size=12,face="bold"))
```

### QC plots after TMT integrator pre-processing and normalization (Protein level)  {.tabset .tabset-fade}

#### By TMT channel/sample {.unnumbered}

##### General Abudance distribution by TMT channel/sample {.unlisted .unnumbered}

```{r fig.height=6, fig.width=9, warning=FALSE}
qcplot <- ggplot(quant_annotated,
                 mapping = aes(x = Channel_mix, 
                               y = Abundance, 
                               fill = recurrence)) +
          geom_boxplot() + 
          #facet_grid(.~recurrence, scales = "free") +
          theme(axis.text.x = element_text(hjust = 0.5, vjust = 0.1, size = 10, angle = 90),
                axis.text.y = element_text(hjust = 0.5, vjust = 0.1, size = 10),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", fill=NA, size=1.5),
                axis.title=element_text(size=12,face="bold"))

print(qcplot)

```

The median-centering after TMT-integrator seem to show a slight mixture-dependent centered median per sample. The paired distribution of initial and recurrent tumor allows the biological effect to be more important than any potential TMT-mixture-based batch effect.

The different abundance distribution can also be explained by a different distribution of missing values.

PCA and sPLS-DA analysis allow to confirm this.

#### By Protein (General)

```{r}
# select/sample N proteins to observe their abundance distribution  
which_prots <- sample_proteins(x = quant_annotated$Protein,
                               size = 50,
                               seed = 156)
```

`r length(which_prots)` proteins were randomly sampled to check for the normality of their 'normalized' Abundance distribution.

```{r fig.height=5.5, fig.width=8, warning=FALSE}

qcplot_prots <- ggplot(quant_annotated %>% 
                         filter(Protein %in% which_prots),
                 mapping = aes(x = reorder(Protein,Abundance,na.rm = TRUE), 
                               y = Abundance)) +
          geom_boxplot() + 
          geom_jitter(color="black", size=0.4, alpha=0.9) +
          labs(subtitle = "Protein IDs arranged by median abudance",
               x = "Protein ID ") + 
          theme(axis.text.x = element_text(hjust = 0.5, vjust = 0.1, size = 10, angle = 90),
                axis.text.y = element_text(hjust = 0.5, vjust = 0.1, size = 10),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", fill=NA, size=1.5),
                axis.title=element_text(size=12,face="bold"))

print(qcplot_prots)
```

#### By Protein and Batch {.unnumbered}

`r length(which_prots)` proteins were randomly sampled to check for the normality of their 'normalized' Abundance distribution.

```{r fig.height=5, fig.width=10, warning=FALSE}
qcplot_prots2 <- ggplot(quant_annotated %>% 
                         filter(Protein %in% which_prots),
                 mapping = aes(x = reorder(Protein,Abundance,na.rm = TRUE), 
                               y = Abundance, 
                               fill = mixture)) +
          geom_boxplot() + 
          geom_jitter(color="black", size=0.4, alpha=0.9) +
          labs(subtitle = "Protein IDs arranged by median abudance",
                       x = "Protein ID ") + 
          facet_grid(.~mixture, scales = "free") +
          theme(axis.text.x = element_text(hjust = 0.5, vjust = 0.1, size = 10, angle = 90),
                axis.text.y = element_text(hjust = 0.5, vjust = 0.1, size = 10),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", fill=NA, size=1.5),
                axis.title=element_text(size=12,face="bold"))

print(qcplot_prots2)
```

The abundance distribution of individual proteins can be considered as normaly distributed in general, which makes then suitable for inferential statistics via linear models.

### Exploratory PCA analisys

```{r echo=FALSE}
## Get predictors for mixtures
expr_matrix2 <- expr_matrix %>%
  dplyr::select(Protein, sample_annotation2$paired_id)

paired_id <- sample_annotation2$paired_id
recurrence <- sample_annotation2$recurrence
patient <- sample_annotation2$patient
mixture <- sample_annotation2$mixture

## get predictor for mixtures wo patient 6
expr_matrixwo6 <- expr_matrix %>%
  dplyr::select(Protein, sample_annotationwo6$paired_id)

paired_idwo6 <- sample_annotationwo6$paired_id
recurrencewo6 <- sample_annotationwo6$recurrence
patientwo6 <- sample_annotationwo6$patient
mixturewo6 <- sample_annotationwo6$mixture

## Prep matrix without missing values for PCA

expr_matnona <- column_to_rownames(expr_matrix2,
                            "Protein") %>%
          as.matrix() %>%
          na.omit()

colnames(expr_matnona) <- paste(sample_annotation2$paired_id, 
                                sample_annotation2$mixture,
                                sep = "_")

t_expr_matnona <- t(expr_matnona)

expr_mat <- column_to_rownames(expr_matrix2,
                               "Protein") %>%
          as.matrix()

t_expr_mat <- t(expr_mat)

## Matrix without patient 6

expr_matnonawo6 <- column_to_rownames(expr_matrixwo6,
                            "Protein") %>%
          as.matrix() %>%
          na.omit()

colnames(expr_matnonawo6) <- paste(sample_annotationwo6$paired_id, 
                                sample_annotationwo6$mixture,
                                sep = "_")

t_expr_matnonawo6 <- t(expr_matnonawo6)

expr_matwo6 <- column_to_rownames(expr_matrixwo6,
                               "Protein") %>%
          as.matrix()

t_expr_matwo6 <- t(expr_matwo6)
```

#### General PCA

##### Percentage of explained variance per component

The PCA was applied on the expression matrix without missing values

```{r}
pca_res = pca(t_expr_matnona, ncomp = 10, center = TRUE, scale = TRUE)
```

```{r fig.width=7, fig.height=5}
plot(pca_res)
```

~45% of variance based on protein abundance can be explained by components 1 and 2 after PCA.

##### PCA plot on samples based on protein abundance

```{r}
# preprocess pca results 
pca_variates <- pca_res$variates$X %>% 
  as.data.frame() %>%
  rownames_to_column("Sample") %>%
  separate(col = "Sample",
           into = c("sample", "recurrence", "mixture"),
           sep = "\\_", extra = "merge") %>%
  mutate(paired_id = paste(sample, recurrence, sep = "_")) %>%
  dplyr::select(-c(mixture, recurrence)) %>%
  left_join(. , sample_annotation2,
            by = "paired_id") %>%
  mutate(Stage = if_else(condition = recurrence == "prim", 
                         true = "Initial", 
                         false = "Recurrent"),
         Mix = case_when(mixture == "tmt_1" ~ "Mix1",
                         mixture == "tmt_2" ~ "Mix2",
                         mixture == "tmt_3" ~ "Mix3"))

pca_variateswopat6 <- pca_variates %>% 
  filter(patient != "x6")

```

```{r fig.width=7.3, fig.height=5}
ggplot(data = pca_variates,
       aes(x = PC1, y = PC2)) +
  geom_point(size = 4, aes(shape = mixture, color = recurrence)) +  
  geom_text(aes(label = patient), position = position_nudge(x = 4.5)) +
  labs(title = "PCA plot of samples\nBased on the abundance values of proteins") +
  xlab(paste("PC1", round(pca_res$prop_expl_var$X[1]*100), "% var explained")) + 
  ylab(paste("PC2", round(pca_res$prop_expl_var$X[2]*100), "% var explained")) + 
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 10, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 10),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")
```

There is no evident batch effect after PCA analysis.

There is a good number of individual patients that show important differences in their PCA variates values in recurrent vs primary.

```{r fig.width=7.3, fig.height=5}
ggplot(data = pca_variateswopat6,
       aes(x = PC1, y = PC2)) +
  geom_point(size = 4, aes(shape = mixture, color = recurrence)) +  
  geom_text(aes(label = patient), position = position_nudge(x = 4.5)) +
  labs(title = "PCA plot of samples\nBased on the abundance values of proteins (patient 6 excluded)") +
  xlab(paste("PC1", round(pca_res$prop_expl_var$X[1]*100), "% var explained")) + 
  ylab(paste("PC2", round(pca_res$prop_expl_var$X[2]*100), "% var explained")) + 
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 10, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 10),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")
```

```{r}
pca_plot <- ggplot(data = pca_variateswopat6,
       aes(x = PC1, y = PC2)) +
  geom_point(size = 4, aes(shape = Mix, color = Stage)) +  
  geom_text(aes(label = patient), position = position_nudge(x = 4.5), size = 2) +
  xlab(paste("PC1", round(pca_res$prop_expl_var$X[1]*100), "% var explained")) + 
  ylab(paste("PC2", round(pca_res$prop_expl_var$X[2]*100), "% var explained")) +
theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 7, angle = 360),
      axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 7),
      panel.background = element_blank(),
      panel.grid.major = element_blank(),
      panel.border = element_rect(colour = "black", fill=NA, size=0.5),
      axis.title = element_text(size = 7),
      legend.text = element_text(size = 7),
      legend.title = element_text(size = 8),
      legend.key.height= unit(3, 'mm'),
      legend.key.width= unit(3, 'mm'),
      legend.position="bottom")
```

```{r}
ggsave(plot = pca_plot, 
       filename = here::here("figures/pca_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 50)

ggsave(plot = pca_plot, 
       filename = here::here("figures/pca_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 63,
       height = 50)
```

### Exploratory sPLS-DA

```{r}
plsda_rest1 <- plsda(X = t_expr_matnona,
                     Y = recurrence,
                     ncomp = 3)
```

```{r fig.width=6, fig.height=5}
plotIndiv(plsda_rest1,
          nd.names = TRUE, ellipse = TRUE, legend = FALSE, title = "sPLS-DA_Rec-Prim", 
          size.title = rel(2.2), size.xlabel = rel(1.5), size.ylabel = rel(1.5), 
          siz.axis = rel(1.2), point.lwd = 0.8, cex = 4)

```

Patient 6 in recurrence behaves suspiciously similar to initial samples. We consider this to be an outlier associated to a potential problem during sample prep.

Therefore we decided to exclude patient 6 from further analyses.

```{r}
plsda_rest2 <- plsda(X = t_expr_matnona,
                     Y = mixture,
                     ncomp = 3)
```

```{r fig.width=6, fig.height=5}
plotIndiv(plsda_rest2,
          nd.names = TRUE, ellipse = TRUE, legend = FALSE, title = "sPLS-DA_Rec-Prim", 
          size.title = rel(2.2), size.xlabel = rel(1.5), size.ylabel = rel(1.5), 
          siz.axis = rel(1.2), point.lwd = 0.8, cex = 4)

```

We can say that the biological differences between initial and recurrent status are stronger than the differences that could come from potential batch effects.

### Exploratory sPLS-DA without patient 6

```{r}
plsda_restwo6 <- plsda(X = t_expr_matnonawo6,
                     Y = recurrencewo6,
                     ncomp = 3)
```

```{r}
plotIndiv(plsda_restwo6,
          nd.names = TRUE, ellipse = TRUE, legend = FALSE, title = "sPLS-DA_Rec-Prim wo Pat6", 
          size.title = rel(2.2), size.xlabel = rel(1.5), size.ylabel = rel(1.5), 
          siz.axis = rel(1.2), point.lwd = 0.8, cex = 4)

```

#### GGplot-based visualization of Exploratory sPLS-DA

```{r}
# preprocess pca results 
plsda_variates <- plsda_restwo6$variates$X %>% 
  as.data.frame() %>%
  rownames_to_column("Sample") %>%
  separate(col = "Sample",
           into = c("sample", "recurrence", "mixture"),
           sep = "\\_", extra = "merge") %>%
  mutate(paired_id = paste(sample, recurrence, sep = "_")) %>%
  dplyr::select(-c(mixture, recurrence)) %>%
  left_join(. , sample_annotation2,
            by = "paired_id")

plsda_variateswpat6 <- plsda_rest1$variates$X %>% 
  as.data.frame() %>%
  rownames_to_column("Sample") %>%
  separate(col = "Sample",
           into = c("sample", "recurrence", "mixture"),
           sep = "\\_", extra = "merge") %>%
  mutate(paired_id = paste(sample, recurrence, sep = "_")) %>%
  dplyr::select(-c(mixture, recurrence)) %>%
  left_join(. , sample_annotation2,
            by = "paired_id")

```

```{r fig.width=7.3, fig.height=5}
ggplot(data = plsda_variateswpat6,
       aes(x = comp1, y = comp2)) +
  geom_point(size = 4, aes(shape = mixture, color = recurrence)) +  
  geom_text(aes(label = patient), position = position_nudge(x = 2)) +
  labs(title = "sPLS-DA plot of samples\nBased on the abundance values of proteins") +
  xlab(paste("sPLS-DA Comp-1", round(plsda_restwo6$prop_expl_var$X[1]*100), "% var explained")) + 
  ylab(paste("sPLS-DA Comp-2", round(plsda_restwo6$prop_expl_var$X[2]*100), "% var explained")) + 
  ggforce::geom_mark_ellipse(mapping = aes(color = recurrence)) +
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 10, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 10),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")
```

Patient 6 in recurrence behaves suspiciously similar to initial samples. We consider this to be an outlier associated to a potential problem during sample prep.

Therefore we decided to exclude patient 6 from further analyses.

```{r fig.width=7.3, fig.height=5}
ggplot(data = plsda_variates,
       aes(x = comp1, y = comp2)) +
  geom_point(size = 4, aes(shape = mixture, color = recurrence)) +  
  geom_text(aes(label = patient), position = position_nudge(x = 2)) +
  labs(title = "sPLS-DA plot of samples - pat6 excluded \nBased on the abundance values of proteins") +
  xlab(paste("sPLS-DA Comp-1", round(plsda_restwo6$prop_expl_var$X[1]*100), "% var explained")) + 
  ylab(paste("sPLS-DA Comp-2", round(plsda_restwo6$prop_expl_var$X[2]*100), "% var explained")) + 
  ggforce::geom_mark_ellipse(mapping = aes(color = recurrence)) +
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 10, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 10),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")
```

The difference between initial and recurrent tumor is better resolved in the sPLS-DA plot when excluding patient 6.

## Sparcity reduction

Only proteins that are identified and quantified in 2 out of 3 TMT mixtures will be kept for the inferential analysis. The proteins missing in 1 mixture will be imputed with `missForest`.

__Function to select proteins to exclude:__

```{r}
sel_proteins_missing <- function(long_matrix,
                                 threshold) {

na_count <- group_by(long_matrix,
                     Protein, mixture) %>%
    summarise(na_count = sum(is.na(Abundance)),
              total = n()) %>% 
    ungroup() %>% 
    mutate(NA_fraction = na_count/total)

na_count_perbatch <- na_count %>%
  group_by(Protein) %>%
  summarise(na_per_batch = sum(NA_fraction)) %>%
  ungroup()

proteins2exclude <- na_count_perbatch %>%
  filter(na_per_batch > threshold) %>%
  pull(Protein)

return(proteins2exclude)
}

```

__Excluded proteins:__

```{r}
proteins2exclude <- sel_proteins_missing(quant_annotated,
                                         threshold = 1)
```

Exclude proteins from initial abundance matrix.

```{r}
expr_matrix_filt <- expr_matrix2 %>%
  filter(!Protein %in% proteins2exclude)
```

### Visualize missing values (before and after filtering)

```{r}
nofiltnas_plot <- vis_miss(expr_matrix) + 
          ggtitle("Distribution of missing values",
                  subtitle = paste("All quant proteins. Total =", nrow(expr_matrix))) + 
          theme(axis.text.x = element_text(hjust = 0.5, vjust = 0.1, size = 10, angle = 90),
                axis.text.y = element_text(hjust = 0.5, vjust = 0.1, size = 10),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", fill=NA, size=1.5),
                axis.title=element_text(size=12,face="bold"))

filtnas_plot <- vis_miss(expr_matrix_filt) + 
          ggtitle("Distribution of missing values",
                  subtitle = paste("Only present in 2/3 TMT mixtures. Total =", nrow(expr_matrix_filt))) + 
          theme(axis.text.x = element_text(hjust = 0.5, vjust = 0.1, size = 10, angle = 90),
                axis.text.y = element_text(hjust = 0.5, vjust = 0.1, size = 10),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", fill=NA, size=1.5),
                axis.title=element_text(size=12,face="bold"))
```

```{r fig.height=6, fig.width=12}
cowplot::plot_grid(nofiltnas_plot, filtnas_plot, ncol = 2)
```

## `missForest` imputation:

```{r}
mat_filt <- expr_matrix_filt %>%
  column_to_rownames("Protein") %>%
  as.matrix()

t_mat_filt <- t(mat_filt)

if(!file.exists(here("rds/missforest_imp_mat_filt.Rds"))){

 t_imp_mat_filt <- missForest::missForest(t_mat_filt)
 t_imp_filt_mat <- t_imp_mat_filt$ximp

 mat_filt_imp <- t(t_imp_filt_mat)

 saveRDS(t_imp_mat_filt,
         file = here("rds/missforest_imp_mat_filt.Rds"))

} else {

 t_imp_mat_filt <- readRDS(here("rds/missforest_imp_mat_filt.Rds"))
 t_imp_filt_mat <- t_imp_mat_filt$ximp

 mat_filt_imp <- t(t_imp_filt_mat)

}

```

## Inferential analysis

__Protein to Gene annotation__

```{r}
prot2gene <- tmt_protdata_frag %>%
  dplyr::select(Protein = index, Gene = gene)
```

__Prep protein abundance matrix__

```{r}
mat_filt_impwo6 <- as.data.frame(mat_filt_imp) %>%
  rownames_to_column("Protein") %>%
  dplyr::select(-starts_with("x6")) %>%
  column_to_rownames("Protein") %>%
  as.matrix()

sample_annotationwo6 <- sample_annotation2 %>%
  filter(patient != "x6")
```

### Set up design matrix

Initial vs recurrent samples are patient-matched, therefore the design matrix is setup so the linear model can account for the patient random effect.

```{r}
design_wo6 <- model.matrix(~patientwo6+recurrencewo6)

rownames(design_wo6) <- sample_annotationwo6$paired_id
```

Check design matrix:

```{r}
print(design_wo6)
```

```{r}
mat_filt_impwo6[1:3,1:12]
```

### Fit limma

```{r}
source(here("scr/fit_limmawo6.R"))
```

```{r}
# fit limma
limma_tab_wo6 <- fit_limmawo6(mat_filt_impwo6, 
                         design_wo6, 
                         method = 'robust', 
                         Limma = "Robust - w Patient effect wo Pat6",
                         prot2gene = prot2gene)

#get proteins increased in recurrence
increased_in_rec <- limma_tab_wo6 %>%
  filter(logFC > 0,
         adj.P.Val < 0.05) %>%
  pull(Protein)

#get proteins decreased in recurrence
decreased_in_rec <- limma_tab_wo6 %>%
  filter(logFC < 0,
         adj.P.Val < 0.05) %>%
  pull(Protein)
```

### Volcano plots

```{r fig.width=8, fig.height=6}
our_volcano(limma_tab_wo6, 
            FC_cutoff = 0, 
            pval_cutoff = 0.05, 
            color_diffex = "red", 
            color_nondifex = "#2a9d8f", 
            interesting_proteins = NULL, 
            vert_line_col = "red",
            hline_col = "red", 
            hline_pos = 0.05, 
            linetype = "dashed",
            increased_in = "Recurrent", 
            comparison_title = "Recurrent vs Initial") +
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 10, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 10),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom") 
```

```{r fig.width=8, fig.height=6, echo=FALSE}
protlabels <- limma_tab_wo6 %>%
      filter(logFC > 0 | logFC < 0,
             adj.P.Val < 0.05)

signi_hits <- nrow(protlabels)

signi_up <- limma_tab_wo6 %>%
      filter(logFC > 0 ,
             adj.P.Val < 0.05) %>%
  nrow()

signi_down <- limma_tab_wo6 %>%
      filter(logFC < 0 ,
             adj.P.Val < 0.05) %>%
  nrow()

size <- 0.4

volcano <- ggplot(data = limma_tab_wo6,
                      mapping = aes(x = logFC, y = -log10(adj.P.Val))) +
      geom_point(data = limma_tab_wo6 %>% filter(logFC > 0,
                                           adj.P.Val < 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), 
                 color = "red",
                 size = size)+
      geom_point(data = limma_tab_wo6 %>% filter(logFC < -0,
                                           adj.P.Val < 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "red",
                 size = size) +
      geom_point(data = limma_tab_wo6 %>% filter(logFC > 0,
                                           adj.P.Val > 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "#2a9d8f",
                 size = size) +
      geom_point(data = limma_tab_wo6 %>% filter(logFC < -0,
                                           adj.P.Val > 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "#2a9d8f",
                 size = size) +
      ggrepel::geom_text_repel(data = protlabels,
                               aes(label = Gene), size = 1.5) +
      geom_hline(yintercept = -log10(0.05),
                 color = "red", linetype = "dashed") +
      xlab("logFC - Recurrent / Primary")+
      theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 6, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 8),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")

```

```{r echo=FALSE}
ggsave(plot = volcano, 
       filename = here::here("figures/volcano_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 50)

ggsave(plot = volcano, 
       filename = here::here("figures/volcano_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 63,
       height = 50)
```

```{r fig.width=8, fig.height=6, echo=FALSE}
protlabels <- limma_tab_wo6 %>%
      filter(logFC > 0 | logFC < 0,
             adj.P.Val < 0.05,
             Gene %in% c("ASAH1", "GPNMB", "SYNM"))

signi_hits <- nrow(protlabels)

size <- 0.4

volcano2 <- ggplot(data = limma_tab_wo6,
                      mapping = aes(x = logFC, y = -log10(adj.P.Val))) +
      geom_point(data = limma_tab_wo6 %>% filter(logFC > 0,
                                           adj.P.Val < 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), 
                 color = "red",
                 size = size)+
      geom_point(data = limma_tab_wo6 %>% filter(logFC < -0,
                                           adj.P.Val < 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "red",
                 size = size) +
      geom_point(data = limma_tab_wo6 %>% filter(logFC > 0,
                                           adj.P.Val > 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "#2a9d8f",
                 size = size) +
      geom_point(data = limma_tab_wo6 %>% filter(logFC < -0,
                                           adj.P.Val > 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "#2a9d8f",
                 size = size) +
      ggrepel::geom_text_repel(data = protlabels,
                               aes(label = Gene), size = 3.5) +
      geom_hline(yintercept = -log10(0.05),
                 color = "red", linetype = "dashed") +
      xlab("logFC - Recurrent / Primary")+
      theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 6, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 8),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")

```

```{r echo=FALSE}
ggsave(plot = volcano2, 
       filename = here::here("figures/volcano2_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 50)

ggsave(plot = volcano2, 
       filename = here::here("figures/volcano2_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 63,
       height = 50)
```

### Boxplots specific proteins

```{r}
quant_anwgene <- left_join(quant_annotated, prot2gene)

tab_2_paired_boxplot <-  quant_anwgene %>%
  filter(Gene %in% c("ASAH1", "SYNM", "GPNMB"))
# prep data
  slim_data <- pivot_longer(expr_matrix_filt, 
                          cols = 2:length(names(expr_matrix_filt)), 
                          names_to = c("patient","condition"),
                          values_to = "Intensity",
                          names_sep = "\\_") %>% 
  group_by(Protein)

  slim_data2 <- pivot_wider(slim_data, 
                          values_from = Intensity, 
                          names_from = condition) %>%
    filter(patient != "x6")

  slim_data3 <- slim_data2 %>% 
    mutate(Abs_diff = rec-prim) %>%
    mutate(rank = row_number(Abs_diff)) %>% 
    dplyr::rename(Initial = prim, Recurrent = rec) %>%
    ungroup() %>%
    filter(Protein %in% c("Q13510", "Q14956", "O15061")) %>%
    mutate(Gene = case_when(Protein == "Q13510" ~ "ASAH1",
                              Protein == "Q14956" ~ "GPNMB",
                              Protein == "O15061" ~ "SYNM"))

```

```{r fig.width=7, fig.height=3.5}
ggpaired(slim_data3,
         cond1 = "Initial",
         cond2 = "Recurrent", fill = "condition",
         ylab = "Normalized Abundance",
         label = NULL,
         repel = TRUE,
         facet.by = "Gene", point.size = 0.2, line.size = 0.1) +
      theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 5, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 5),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom",
        strip.text = element_text(hjust = 0.5, vjust = 0, size = 8, angle = 360)) 
```

```{r echo=FALSE}
paired_boxplots <- ggpaired(slim_data3,
         cond1 = "Initial",
         cond2 = "Recurrent", fill = "condition",
         ylab = "Normalized Abundance",
         label = NULL,
         repel = TRUE,
         facet.by = "Gene", point.size = 0.2, line.size = 0.1) +
      theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 5, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 5),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title = element_text(size = 6),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 6),
        legend.key.height= unit(2, 'mm'),
        legend.key.width= unit(2, 'mm'),
        legend.position="bottom",
        strip.text = element_text(hjust = 0.5, vjust = 0, size = 6, angle = 360)) 
```

```{r echo=FALSE}
ggsave(plot = paired_boxplots, 
       filename = here::here("figures/paired_box_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 61,
       height = 45)

ggsave(plot = paired_boxplots, 
       filename = here::here("figures/paired_box_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 61,
       height = 45)
```

## ID coverage (final)

### Coverage plot

```{r}
ids_summary2 <- quant_annotated %>% 
  dplyr::select(Protein, mixture, Abundance) %>% 
  distinct(Protein, mixture, .keep_all = TRUE) %>%
  group_by(mixture) %>%
  summarise(`Nr of IDs` = sum(!is.na(Abundance))) %>% 
  bind_rows(.,
            tibble(mixture = "Quant_in_all",
                   `Nr of IDs` = nrow(expr_matnona))) %>%
  bind_rows(.,
            tibble(mixture = "Quant_2/3_mixtures",
                   `Nr of IDs` = nrow(mat_filt_impwo6))) %>%
  bind_rows(.,
            tibble(mixture = "Overall",
                   `Nr of IDs` = nrow(expr_matrix)))

ids_summary2 <- mutate(ids_summary2,
                       Mixture = case_when(mixture == "tmt_1" ~ "Mix_1",
                                           mixture == "tmt_2" ~ "Mix_2",
                                           mixture == "tmt_3" ~ "Mix_3",
                                           mixture == "Quant_in_all" ~ "Quant. in all",
                                           mixture == "Overall" ~ "Overall",
                                           mixture == "Quant_2/3_mixtures" ~ "Quant. in 2/3")) %>%
  mutate(Mixture = factor(Mixture, levels = c("Overall","Quant. in 2/3", "Quant. in all",
                                               "Mix_3", "Mix_2", "Mix_1")))
```

Number of identified proteins: `r nrow(expr_matrix)`

Number of identified and quantified proteins in all samples: `r nrow(expr_matnona)`

Number of identified and quantified proteins 2 out of 3 batches: `r nrow(mat_filt_impwo6)`

```{r fig.width=7, fig.height=4}
ggplot(data = ids_summary2,
       aes(y = `Nr of IDs`, x = mixture)) +
  geom_col(fill = "#008683") +

  geom_text(size = 5, position = position_stack(vjust = 0.5), aes(label = `Nr of IDs`)) +
  coord_cartesian(ylim = c(0, 5000))+
  coord_flip() +
  labs(title = "Number of Identified and quantified Proteins by Mixture") +
  ylab("Nr of IDed and Quant Proteins") + 
  xlab("TMT Mixture") + 
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0.1, size = 14, angle = 0),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.1, size = 12, angle = 0),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey"),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5),
        axis.title=element_text(size=12, face="bold"))
```
```{r echo=FALSE}
barplot_cover <- ggplot(data = ids_summary2,
       aes(y = `Nr of IDs`, x = Mixture)) +
  geom_col(fill = "#008683") + 
  geom_text(size = 2, position = position_stack(vjust = 0.5), aes(label = `Nr of IDs`)) +
  coord_cartesian(ylim = c(0, 5000))+
  coord_flip() +
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 7, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 7),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")
```

```{r echo=FALSE}
ggsave(plot = barplot_cover, 
       filename = here::here("figures/barplot_coverage_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 50)

ggsave(plot = barplot_cover, 
       filename = here::here("figures/barplot_coverage_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 63,
       height = 50)
```

## Median abundance distribution of filtered and inputed samples

```{r}
expr_matrix_filtwo6 <- as.data.frame(mat_filt_imp) %>%
  rownames_to_column("Protein") %>%
  dplyr::select(-starts_with("x6"))

long_mat_f <- pivot_longer(expr_matrix_filtwo6,
                         cols = starts_with("x"),
                         names_sep = "_",
                         names_to = c("patient", "recurrence"),
                         values_to = "Abundance")

long_mat_3 <- long_mat_f %>%
  mutate(paired_id = paste(patient, recurrence, sep = "_"))

paired_annot_uniq2 <- long_mat_3 %>%
  dplyr::select(patient, recurrence, paired_id) %>%
  distinct()

paired_annot_uniq2$paired_id
## Annotated abundance data in long format ----

quant_annotated_f <- left_join(long_mat_3,
                             sample_annotation2, 
                             by = c("paired_id", "patient","recurrence")) %>%
  mutate(Channel_mix = paste(mixture,channel,  sep = "_"),
         paired_id = factor(paired_id, levels = paired_annot_uniq2$paired_id)) %>%
  mutate(Sample = str_replace_all(paired_id, pattern = "prim", replacement = "init"),
         Stage = if_else(recurrence == "prim",
                         true = "Initial",
                         false = "Recurrent"))

```

### By TMT channel/sample {.unnumbered}

#### General Abudance distribution by TMT channel/sample {.unlisted .unnumbered}

```{r}
qcplot_f <- ggplot(quant_annotated_f,
                 mapping = aes(x = Sample, 
                               y = Abundance, 
                               fill = Stage)) +
          geom_boxplot(outlier.size=0.1, size = 0.25) + 
          theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 7, angle = 90),
                axis.text.y = element_text(hjust = 0.5, vjust = 0, size = 6),
                panel.background = element_blank(),
                panel.grid.major = element_blank(),
                panel.border = element_rect(colour = "black", fill=NA, size=0.5),
                axis.title=element_text(size=8),
                axis.title.x = element_blank(),
                legend.text = element_text(size = 7),
                legend.title = element_text(size = 8),
                legend.key.height= unit(3, 'mm'),
                legend.key.width= unit(3, 'mm'),
                legend.position="bottom")

```

```{r fig.height=4, fig.width=5, warning=FALSE, echo=FALSE}
print(qcplot_f)
```
```{r echo=FALSE}
ggsave(plot = qcplot_f, 
       filename = here::here("figures/boxplot_abund_dist_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 50)

ggsave(plot = qcplot_f, 
       filename = here::here("figures/boxplot_abund_dist_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 63,
       height = 50)
```

## Over-representation analyses/tests (ORAs)

### Prep tabular info for ORAs

```{r}
tab_sig_prots_tryptic <- bind_rows(tibble(protein = increased_in_rec,
                                  characteristic = "up-regulated"),
                           tibble(protein = decreased_in_rec,
                                  characteristic = "down-regulated"))

limma_tab_wo62_tryptic <- limma_tab_wo6

unip2symb_tryptic <- bitr(limma_tab_wo62_tryptic$Protein, 
                  fromType = "UNIPROT", 
                  toType = c("SYMBOL", "ENTREZID"),
                  OrgDb = org.Hs.eg.db) %>%
  dplyr::rename(protein = UNIPROT) %>%
  mutate(Protein = protein)

tab_sig_prots_tryptic <- left_join(tab_sig_prots_tryptic, unip2symb_tryptic)

unip2symbIDed_tryptic <- bitr(identified_proteins$protein_id, 
                  fromType = "UNIPROT", 
                  toType = c("SYMBOL", "ENTREZID"),
                  OrgDb = org.Hs.eg.db)

limma_tab_wo62_tryptic <- left_join(limma_tab_wo62_tryptic, unip2symb_tryptic)

geneList <- limma_tab_wo62_tryptic$logFC
names(geneList) <- limma_tab_wo62_tryptic$Protein
geneList <-  sort(geneList, decreasing = TRUE)

geneListr <- limma_tab_wo62_tryptic$logFC
names(geneListr) <- limma_tab_wo62_tryptic$ENTREZID
geneListr <-  sort(geneListr, decreasing = TRUE)
```

### Reactome

```{r}
group_comparison_react <- compareCluster(ENTREZID~characteristic, 
                                              data=tab_sig_prots_tryptic, 
                                              fun="enrichPathway",
                                              organism = "human",
                                              pvalueCutoff = 0.05,
                                              pAdjustMethod = "BH",
                                              qvalueCutoff = 0.2,
                                              universe = unip2symbIDed_tryptic$ENTREZID,
                                              minGSSize = 10,
                                              maxGSSize = 1000,
                                              readable = TRUE)
```

#### Dotplot Reactome

```{r}
enrichplot::dotplot(group_comparison_react, x = "characteristic") + 
          xlab("Quant in Recurrent status") +
  scale_color_continuous(low="red", high="blue",
            guide=guide_colorbar(reverse=TRUE),breaks = c(0.01, 0.04)) +
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 8, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 7),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")
```

```{r}
dotplot_react <- enrichplot::dotplot(group_comparison_react, x = "characteristic") + 
          xlab("Quant in Recurrent status") +
  scale_color_continuous(low="red", high="blue",
            guide=guide_colorbar(reverse=TRUE),breaks = c(0.01, 0.04)) +
  scale_size_continuous(breaks = c(0.10, 0.25)) +
  theme(axis.text.x = element_text(hjust = 0.4, vjust = 0.1, size = 6, angle = 340),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 6),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 6),
        legend.key.height= unit(2, 'mm'),
        legend.key.width= unit(2, 'mm'),
        legend.position="bottom")
```
```{r}
ggsave(plot = dotplot_react, 
       filename = here::here("figures/dotplot_reactome_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 91,
       height = 90)

ggsave(plot = dotplot_react, 
       filename = here::here("figures/dotplot_reactome_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 91,
       height = 90)
```

#### Cnet plot Reactome

```{r}
enrichplot::cnetplot(group_comparison_react) + 
  theme(legend.text = element_text(size = 7),
                legend.title = element_text(size = 8),
                legend.key.height= unit(3, 'mm'),
                legend.key.width= unit(3, 'mm'),
                legend.position="bottom")
```

```{r}
cnet_plot <- enrichplot::cnetplot(group_comparison_react) + 
  theme(legend.text = element_text(size = 7),
        text = element_text(size = 4),
                legend.title = element_text(size = 8),
                legend.key.height= unit(3, 'mm'),
                legend.key.width= unit(3, 'mm'),
                legend.position="bottom")
```

```{r}
ggsave(plot = cnet_plot, 
       filename = here::here("figures/cnet_plot_reactome_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 110,
       height = 110)

ggsave(plot = cnet_plot, 
       filename = here::here("figures/cnet_plot_reactome_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 110,
       height = 110)

ggsave(plot = cnet_plot, 
       filename = here::here("figures/cnet_plot_reactome_proteomics_gbm.svg"), 
       device = "svg",
       units = "mm",
       width = 110,
       height = 110)
```

## Protein coverage plots

```{r}
source(here("scr/protein_peptide_coverage_plot.R"))
```

### Annotate peptides from interesting proteins

```{r}
source(here("scr/annotate_peptides.R"))
```

```{r}
# get protein to peptide pairs
pept2prot_inter <- identified_peptides %>%
  filter(gene %in% c("ASAH1", "SYNM", "GPNMB")) %>%
  dplyr::select(Peptide = sequence, Genes = protein_id)

annotated_peptides <- annotate_peptides(expr_mat = pept2prot_inter,
                                        fasta = fasta_ident_prots)

annotated_peptides2 <- annotated_peptides %>%
  mutate(peptide_length = str_length(Peptide))
```

### ASAH1

```{r}
feat_asah1 <- drawProteins::get_features("Q13510")

feat_asah1df <- drawProteins::feature_to_dataframe(feat_asah1)
#Get peptides as coverage features for plotting

cov_feat_asah1 <- get_coverage(annotated_peptides = annotated_peptides2, 
                               id = "Q13510")
#Merge coverage feature info with protein sequence features info:

fullfeat_asah1 <- add_peptides(feat_asah1df,
                               peptide_coverage_data = cov_feat_asah1)
# drawCanvas
canvas <- draw_canvas(fullfeat_asah1)

# draw the protein chain
wchain <- draw_chains(canvas, fullfeat_asah1)

# draw protein domains
wdomainsasah1 <- draw_domains(wchain, data = fullfeat_asah1)
#pasa <- drawProteins::draw_folding(wdomainsasah1, fullfeat_asah1)
cov_feat_asah1$order <- 1.4

pepcov_asah1 <- draw_peptides(wdomainsasah1, cov_feat_asah1)
pepcov_asah1
```

```{r}
ggsave(plot = pepcov_asah1, 
       filename = here::here("figures/asah1_coverage_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 150,
       height = 100)

ggsave(plot = pepcov_asah1, 
       filename = here::here("figures/asah1_coverage_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 150,
       height = 100)
```

### SYNM coverage plot
Get features from Uniprot

```{r}
feat_SYNM <- drawProteins::get_features("O15061")
feat_SYNMdf <- drawProteins::feature_to_dataframe(feat_SYNM)
#Get peptides as coverage features for plotting

cov_feat_SYNM <- get_coverage(annotated_peptides = annotated_peptides2, 
                               id = "O15061")
#Merge coverage feature info with protein sequence features info:

fullfeat_SYNM <- add_peptides(feat_SYNMdf,
                               peptide_coverage_data = cov_feat_SYNM)
# drawCanvas
canvasSYNM <- draw_canvas(fullfeat_SYNM)

# draw the protein chain
wchainSYNM <- draw_chains(canvasSYNM, fullfeat_SYNM)
cov_feat_SYNM$order <- 1.4

pepcov_SYNM <- draw_peptides(wchainSYNM, cov_feat_SYNM)
pepcov_SYNM
```

```{r}
ggsave(plot = pepcov_SYNM, 
       filename = here::here("figures/SYNM_coverage_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 150,
       height = 100)

ggsave(plot = pepcov_SYNM, 
       filename = here::here("figures/SYNM_coverage_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 150,
       height = 100)
```

# Analysis of proteolytic processing

## Load semi-specific search results

```{r}
tmt_peptdata_frag <- read_tsv(here("data/semi_specific_search_fragpipe17/tmt-report/abundance_peptide_MD.tsv")) %>% 
  janitor::clean_names()

pept_ids_mix1 <- read_delim(here("data/semi_specific_search_fragpipe17/mix_1/peptide.tsv")) %>%
                    clean_names() %>%
          dplyr::select(-c(starts_with("empty")))

pept_ids_mix2 <- read_delim(here("data/semi_specific_search_fragpipe17/mix_2/peptide.tsv")) %>%
                    clean_names() %>%
          dplyr::select(-c(starts_with("x"), starts_with("empty"), "master"))

pept_ids_mix3 <- read_delim(here("data/semi_specific_search_fragpipe17/mix_3/peptide.tsv")) %>%
                    clean_names() %>%
          dplyr::select(-c(starts_with("x"), starts_with("empty"), "master"))

identified_proteins_semi <- read_tsv(here("data/semi_specific_search_fragpipe17/combined_protein.tsv")) %>% 
  janitor::clean_names()

```

## Wrangle and merge peptide ID/quant with sample annotation

```{r}
# wide expression matrix with NAs
pept_matrix <- dplyr::select(tmt_peptdata_frag,
                             index, starts_with("x"))

pept_long_mat <- pivot_longer(pept_matrix,
                         cols = starts_with("x"),
                         names_to = c("patient"),
                         values_to = "Abundance") %>%
  mutate(recurrence = str_extract(patient, "[^0-9]+$")) %>%
  mutate(patient = str_remove_all(patient, "[^0-9]+$"))

pept_long_mat_2 <- pept_long_mat %>%
  mutate(paired_id_semi = paste(patient, recurrence, sep = "_"))

sample_annotation2 <- sample_annotation2 %>%
  mutate(paired_id_semi = paste(patient, recurrence, sep = "_"))

## Annotated abundance data in long format ----

pept_quant_annotated <- left_join(pept_long_mat_2,
                             sample_annotation2, 
                             by = c("paired_id_semi", "patient","recurrence")) %>%
  mutate(Channel_mix = paste(mixture,channel,  sep = "_"))

## Wide xpression matrices without NAs
pept_expr_matnona <- column_to_rownames(pept_matrix,
                            "index") %>%
          as.matrix() %>%
          na.omit()
```

Wrangle and merge peptide ids for further annotation based on they proteolytic specificity

```{r}
peptide_joined1 <- full_join(pept_ids_mix1, pept_ids_mix2) %>%
                    mutate(duplicated = duplicated(peptide))

peptide_joined <- full_join(peptide_joined1, pept_ids_mix3) %>%
                    mutate(duplicated = duplicated(peptide))

peptide_ids_filtered <- filter(peptide_joined,
                               peptide %in% peptide_joined$peptide) %>%
                    distinct(peptide, assigned_modifications,observed_modifications, .keep_all = TRUE) %>%
                    dplyr::rename(Peptide = peptide) %>%
                    dplyr::mutate(Genes = protein_id)

```

### Annotate peptides with cleavage information

```{r}
fasta1 <- seqinr::read.fasta(here("data/semi_specific_search_fragpipe17/mix_1/protein.fas"),
                            seqtype = "AA", as.string = TRUE)

fasta2 <- seqinr::read.fasta(here("data/semi_specific_search_fragpipe17/mix_2/protein.fas"),
                             seqtype = "AA", as.string = TRUE)

fasta3 <- seqinr::read.fasta(here("data/semi_specific_search_fragpipe17/mix_3/protein.fas"),
                             seqtype = "AA", as.string = TRUE)

fasta22 <- append(fasta1, fasta2)

fasta <- append(fasta22, fasta3)

fasta <- discard(fasta,
                 duplicated(names(fasta)))

attr(fasta[[1]], "Annot")

extr_annt <- function(x){
  ext <- attr(x, "Annot")
}

names <- purrr::map_chr(fasta, .f = extr_annt)

seqinr::write.fasta(fasta,
                    file.out = here("data/semi_specific_search_fragpipe17/protein_combined.fas"),
                    names = names)

if(!file.exists(here("results/semi_tryptic/cleavage_annoated_peptides_final.tsv"))){
  cleavage_annoated_peptides <- annotate_peptides(expr_mat = peptide_ids_filtered, 
                                                  fasta = fasta,
                                                  decoy_tag = "rev_")

  write_tsv(cleavage_annoated_peptides, here("results/semi_tryptic/cleavage_annoated_peptides_final.tsv"))
} else {
  cleavage_annoated_peptides <- read_tsv(here("results/semi_tryptic/cleavage_annoated_peptides_final.tsv"))
}

cleavage_annoated_peptides <- dplyr::mutate(cleavage_annoated_peptides,
                                            Genes = protein_id)

peptides_annotated <- left_join(peptide_ids_filtered, cleavage_annoated_peptides,
                                by = c("Peptide", "protein_id"))

```

## Sparcity reduction

Only peptides that were identified and quantified in all 3 mixtures would be used for further inferential analyses for proteolytic processing.

```{r}
pept_matrix_nona <- pept_matrix %>% 
  na.omit()

pept_nona <- pept_matrix_nona %>%
  dplyr::select(index) %>%
  separate(col = index, 
           sep = "_", into = c("protein", "peptide"))

pept_matrix_nonawop6 <- pept_matrix %>% 
  dplyr::select(-starts_with("x6")) %>%
  na.omit()

```

## Coverage information (peptide-level)

```{r}
ids_summarypep <- pept_quant_annotated %>% 
  dplyr::select(index, mixture, Abundance) %>% 
  distinct(index, mixture, .keep_all = TRUE) %>%
  group_by(mixture) %>%
  summarise(`Nr of IDs` = sum(!is.na(Abundance))) %>% 
  bind_rows(.,
            tibble(mixture = "In all three",
                   `Nr of IDs` = nrow(pept_matrix_nona)))# %>%
  #bind_rows(.,
  #          tibble(mixture = "Overall",
  #                `Nr of IDs` = nrow(pept_matrix)))

ids_summarypep <- mutate(ids_summarypep,
                       Mixture = case_when(mixture == "tmt_1" ~ "Mix_1",
                                           mixture == "tmt_2" ~ "Mix_2",
                                           mixture == "tmt_3" ~ "Mix_3",
                                           mixture == "In all three" ~ "Quant. in all",
                                           mixture == "Overall" ~ "Overall",
                                           mixture == "Quant_2/3_mixtures" ~ "Quant. in 2/3")) %>%
  mutate(Mixture = factor(Mixture, levels = c("Overall", "Quant. in all",
                                               "Mix_3", "Mix_2", "Mix_1")))
```

```{r fig.height=4, fig.width=7}
ggplot(data = ids_summarypep,
       aes(y = `Nr of IDs`, x = Mixture)) +
  geom_col(fill = "#008683") + 
  geom_text(size = 5, position = position_stack(vjust = 0.5), 
            aes(label = `Nr of IDs`)) +
  coord_cartesian(ylim = c(0, 5000))+
  coord_flip() +
  labs(title = "Number of Identified and quantified Peptides by Mixture") +
  ylab("Nr of IDed and Quant Peptides") + 
  xlab("TMT Mixture") + 
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 7, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 7),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")
```
```{r}
barplot_pept_cover <- ggplot(data = ids_summarypep,
       aes(y = `Nr of IDs`, x = Mixture)) +
  geom_col(fill = "#008683") + 
  geom_text(size = 2, position = position_stack(vjust = 0.5), aes(label = `Nr of IDs`)) +
  coord_flip() +
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 7, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 7),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")
```  
```{r}
ggsave(plot = barplot_pept_cover, 
       filename = here::here("figures/barplot_coverage_pept_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 38,
       height = 50)

ggsave(plot = barplot_pept_cover, 
       filename = here::here("figures/barplot_coverage_pept_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 38,
       height = 50)
```

## Label N-terminal peptides

```{r echo=TRUE}
source(here("scr/annotate_nterm.R"))
```

```{r echo=TRUE}
nterannot <- annotate_nterm(peptides_annotated,
                            tmtmass = 229.1629) %>%
  clean_names()
```

### Merge peptide annotation with TMT integrator quant report

```{r}
# all quantified peptides
tmt_reprt_annot <- left_join(tmt_peptdata_frag, nterannot)

check_na <- filter(tmt_reprt_annot, is.na(semi_type))

excluded_tmt_reprt_annot <- anti_join(nterannot, tmt_reprt_annot) %>% 
                    clean_names()

#only those present in all mixtures
tmt_peptdata_fragnona <- tmt_peptdata_frag %>%
  filter(peptide %in% pept_nona$peptide)

tmt_reprt_annotnona <- left_join(tmt_peptdata_fragnona, nterannot) 
```

## Summarize feature counts per feature

```{r echo=TRUE}
source(here("scr/summarize_peptide_counts.R"))
```

```{r echo=TRUE}
summary_count <- summarize_peptide_counts(tmt_reprt_annot) 
```

### Barplot

These are the ones that would be used for inferential statistics.

```{r echo=TRUE}
to_count_info <- tmt_reprt_annot %>% 
   filter(index %in% pept_matrix_nona$index) %>%
                      dplyr::select(index, specificity, nterm, semi_type, tmt_tag,
                                    aa_after, aa_before, following_10_resid, previous_10_resid, start_position, end_position) %>%
  mutate(specificity = if_else(str_detect(semi_type, "unspecific"), 
                               true = "semi_specific",
                               false = specificity),
         duplicated = duplicated(index)) %>%
  distinct()

to_count_info_semi <- to_count_info %>%
                           dplyr::select(index,specificity) %>%
                           distinct() %>%
  mutate(duplicated = duplicated(index))

false_duplication <- to_count_info_semi %>%
  filter(duplicated) %>%
  pull(index)

to_count_info_semi <- to_count_info_semi %>%
  mutate(specificity = if_else(index %in% false_duplication, 
                               true = "semi_specific",
                               false = specificity)) %>%
  dplyr::select(-duplicated) %>%
  distinct()

  n_semi <- dplyr::count(to_count_info_semi, specificity) %>% 
                      dplyr::rename(feature_type = specificity) %>%
                      dplyr::mutate(category = "specificity")

  n_total <- tibble(feature_type = "Total",
                    n = length(unique(to_count_info$index)),
                    category = "Total")

  n_term <- dplyr::count(to_count_info, nterm) %>% 
                      dplyr::rename(feature_type = nterm) %>%
                      dplyr::mutate(category = "N-term")

  summary_count_pept_nona <- bind_rows(n_semi,
                                       n_term,
                                       n_total)

total_sumcount <- summary_count_pept_nona %>%
  filter(feature_type == "Total") %>%
  pull(n)

summary_countnona_no_tot <- summary_count_pept_nona %>%
  dplyr::filter(category != "Total") %>%
  dplyr::rename(`Feature category` = category)
```

```{r fig.width=5, fig.height=2, echo=FALSE}
ggplot(summary_countnona_no_tot,
       aes(x = feature_type, y = n, fill = `Feature category`)) + 
  coord_flip()+ 
  geom_col() +
  geom_text(aes(label = n), hjust = -0.2, size = 2.5) + 
  scale_y_continuous(limits = c(0, 17000), breaks = c(0, 7000, 15000)) + 
  labs(y = "Number of IDs", x = "Feature type") +
  facet_wrap(~`Feature category`, nrow = 1, scales="free") +
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 7, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 7),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom", 
        strip.background = element_blank(),
        strip.text = element_blank())
```

```{r echo=FALSE}
summary_pept_annot <- ggplot(summary_countnona_no_tot,
       aes(x = feature_type, y = n, fill = `Feature category`)) + 
  coord_flip()+ 
  geom_col() +
  geom_text(aes(label = n), hjust = -0.2, size = 2) + 
  scale_y_continuous(limits = c(0, 20000), breaks = c(0, 7000, 15000)) + 
  labs(y = "Number of IDs", x = "Feature type", 
       caption = paste0("Total = ",total_sumcount)) +
  facet_wrap(~`Feature category`, nrow = 1, scales="free") +
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 7, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 7),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom", 
        strip.background = element_blank(),
        strip.text = element_blank())

```
```{r}
ggsave(plot = summary_pept_annot, 
       filename = here::here("figures/barplot_coverage_peptidesfeatures_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 76,
       height = 50)

ggsave(plot = summary_pept_annot, 
       filename = here::here("figures/barplot_coverage_peptidesfeatures_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 76,
       height = 50)
```

## Inferential analysis (peptide-level)

### Prep design matrix

```{r}
design_limmawo6 <- model.matrix(~patientwo6+recurrencewo6)

sample_annotationwo6 <- sample_annotationwo6 %>%
  mutate(paired_id_semi = paste(patient, recurrence, sep = ""))

rownames(design_limmawo6) <- sample_annotationwo6$paired_id_semi
```

### Prep abundance matrices

```{r}
pep_matwo6 <- dplyr::select(pept_matrix_nonawop6,
                             index, rownames(design_limmawo6))  %>%
  column_to_rownames("index") %>%
  as.matrix()

```

```{r}
colnames(pep_matwo6)
```
```{r}
rownames(design_limmawo6)
```

The column names in our peptide abundance matrix have the same order as the row names in our design matrix.

### Peptide-level model fit

```{r}
fit_limmapeptwo6 <- function(mat, design, method,Limma){
  limmafit <- lmFit(mat, design, method = method)
  limmafit <- eBayes(limmafit)
  limma_tab <- topTable(limmafit, coef = "recurrencewo6rec", number = Inf, adjust.method = "BH") %>%
  mutate(Protein = rownames(.),
         Limma = Limma)

  return(limma_tab)
}
```

```{r echo=TRUE, warning=FALSE, message=FALSE}
fit_mod_peptwo6 <- fit_limmapeptwo6(pep_matwo6, 
                         design_limmawo6, 
                         method = 'robust', 
                         Limma = "Robust - w Patient effect")

compar_tab_peptwo6 <-  fit_mod_peptwo6 %>%
                    separate(Protein, 
                             into = c("protein", "peptide"), 
                             remove = FALSE) %>%
  dplyr::rename(index = Protein)

```

### Extract coefficients and feature specific FDR correction

#### Interesting features

```{r}
features1 <- tmt_reprt_annotnona %>%
                    dplyr::select(peptide, index, specificity, nterm, 
                                  semi_type, is_terminal)

interesting_features1 <- features1 %>%
                    filter(specificity == "semi_specific",
                           is_terminal == "not_terminal") 
```

```{r}
# load function for feature specific fdr correction
source(here("scr/feature_specific_fdr_correction.R"))
```

```{r}
# load function for volcano plot
source(here("scr/plot_volcano.R"))
```

```{r}
# excluding patient 6
compar_tab_correctedwo6_pept <- feature_fdr_correction(compar_tab_peptwo6,
                                               features1,
                                               interesting_features1)

compar_tab_interestingwo6_pept <- compar_tab_correctedwo6_pept %>% 
  dplyr::rename(Protein = protein) %>%
  filter(fdr_correction == 'feature-specific') %>%
  left_join(.,prot2gene)

increased_recwo6_pept <- compar_tab_interestingwo6_pept %>%
                    filter(logFC>0,
                           adj.P.Val < 0.05) %>% 
  pull(peptide)

decreased_recwo6_pept <- compar_tab_interestingwo6_pept %>%
                    filter(logFC<0,
                           adj.P.Val < 0.05) %>% 
  pull(peptide)

check_increased_pept <- cleavage_annoated_peptides %>%
  filter(Peptide %in% increased_recwo6_pept)

```

```{r}
length(increased_recwo6_pept)
```

```{r}
length(decreased_recwo6_pept)
```

`r length(increased_recwo6_pept)` semi-specific peptides are increased in Recurrent tumor and `r length(decreased_recwo6_pept)` are decreased in Recurrent tumor.

### Volcano peptides

```{r fig.width=8, fig.height=6}
recwo6_pept22 <- compar_tab_interestingwo6_pept %>%
  filter(adj.P.Val < 0.05)

size <- 0.4

volcano_pept <- ggplot(data = compar_tab_interestingwo6_pept,
                      mapping = aes(x = logFC, y = -log10(adj.P.Val))) +
      geom_point(data = compar_tab_interestingwo6_pept %>% filter(logFC > 0,
                                           adj.P.Val < 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), 
                 color = "red",
                 size = size)+
      geom_point(data = compar_tab_interestingwo6_pept %>% filter(logFC < -0,
                                           adj.P.Val < 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "red",
                 size = size) +
      geom_point(data = compar_tab_interestingwo6_pept %>% filter(logFC > 0,
                                           adj.P.Val > 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "#2a9d8f",
                 size = size) +
      geom_point(data = compar_tab_interestingwo6_pept %>% filter(logFC < -0,
                                           adj.P.Val > 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "#2a9d8f",
                 size = size) +
      geom_hline(yintercept = -log10(0.05),
                 color = "red", linetype = "dashed") +
      xlab("logFC - Recurrent / Primary") + 
      theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 6, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 8),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")

```

```{r}
ggsave(plot = volcano_pept, 
       filename = here::here("figures/volcano_proteolysis_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 50)

ggsave(plot = volcano_pept, 
       filename = here::here("figures/volcano_proteolysis_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 63,
       height = 50)
```

## Proportional abundance of proteolytic peptides

### Prepare data

__Wrangle data__

```{r}
pept_semi_annot <- peptides_annotated %>%
  dplyr::select(Peptide, specificity, is_terminal)

pept_quant_to_summary <- pept_quant_annotated %>%
  separate(col = index, 
           into = c("Protein", "Peptide"), 
           sep = "\\_", 
           remove = FALSE) %>%
  left_join(.,pept_semi_annot)

pep_quant_presummary <- pept_quant_to_summary %>%
  dplyr::select(Protein, Peptide, recurrence, paired_id, 
                Abundance, specificity, is_terminal)

```

__Prepare data-frame with abudance percentage of semi-specific peptides per sample__

```{r}
pept_summary_int_all <- pep_quant_presummary %>%
  group_by(paired_id, recurrence) %>%
  summarise(Sum_All = sum(Abundance, na.rm = TRUE))

pept_summary_int_semi <- pep_quant_presummary %>%
  filter(specificity == "semi_specific",
         is_terminal == "not_terminal") %>%
  group_by(paired_id, recurrence) %>%
  summarise(Sum_Semi = sum(Abundance, na.rm = TRUE))

pept_sum_summary <- left_join(pept_summary_int_all, pept_summary_int_semi) %>%
  mutate(Percentage = Sum_Semi/Sum_All * 100,
         Stage = if_else(recurrence == "prim",
                         true = "Initial",
                         false = "Recurrence"))

```

### Generate plot % of abundance of semi-specific peptides

```{r}
plotSumEM <- ggplot(pept_sum_summary, 
                    aes(x = Stage, 
                        y = Percentage, fill = Stage, 
                        cex.axis = 1.5)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 0.5) +
  # Box plot with jittered points
  # 0.2 : degree of jitter in x direction
  # geom_jitter(shape=16, position=position_jitter(0.2))
  ylab("Proportional intensity of semi-tryptic peptides [%]") +
  geom_signif(
    comparisons = list(c("Initial", "Recurrence")),
    map_signif_level = TRUE
  ) + 
  stat_compare_means(method="t.test") +
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 6, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 8),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")

```

```{r}
plotSumEM
```

```{r}
ggsave(plot = plotSumEM, 
       filename = here::here("figures/box_perc_semispec_pepts.tiff"), 
       device = "tiff",
       units = "mm",
       width = 40,
       height = 50)

ggsave(plot = plotSumEM, 
       filename = here::here("figures/box_perc_semispec_pepts.eps"), 
       device = "eps",
       units = "mm",
       width = 40,
       height = 50)
```

## Comparative analysis of semi-specific peptides vs protein abundance

### Prepare data

#### Get log2FC Summary of abundances of semi-specific peptides

```{r}
#taking output from limma semi-specific
log2_semi_tryptic <- dplyr::select(compar_tab_interestingwo6_pept,
                                   logFC, Protein, peptide, specificity, 
                                   is_terminal,
                                   Gene) %>%
  filter(specificity == "semi_specific",
         is_terminal == "not_terminal") %>%
  dplyr::rename(logFC_semi_peptide = logFC) %>%
  dplyr::select(-c(specificity, is_terminal))

```

#### Get log2 Summary of protein abundances based only on fully-tryptic peptides

```{r}
med_prot_quant_specific <- pep_quant_presummary %>%
  group_by(Protein, paired_id) %>%
  summarise(`median(Abundance)` = median(Abundance, na.rm = TRUE)) %>%
  ungroup()
```

##### Prep wide matrix of median protein abundances from only fully-tryptic peptides

```{r}
# prep wide matrix for median proteins abundances (arousing only from fully-tryptic peptides)
wide_premat_prot_only_tryp <- med_prot_quant_specific %>%
  pivot_wider(id_cols = c("Protein"),
              values_from = `median(Abundance)`, 
              names_from = paired_id) %>%
  # keep only proteins present in 2 / 3 of the samples
  filter(!Protein %in% proteins2exclude)

mat_prot_only_tryp <- wide_premat_prot_only_tryp %>%
  column_to_rownames("Protein") %>%
  as.matrix()
```

__missForest-based imputation__

```{r}
t_mat_prot_only_tryp <- t(mat_prot_only_tryp)

if(!file.exists(here("rds/missforest_imp_mat_prot_only_tryp.Rds"))){

 imp_prot_only_tryp <- missForest::missForest(t_mat_prot_only_tryp)
 t_mat_imp_prot_only_tryp <- imp_prot_only_tryp$ximp

 mat_prot_only_tryp <- t(t_mat_prot_only_tryp)

 saveRDS(imp_prot_only_tryp,
         file = here("rds/missforest_imp_mat_prot_only_tryp.Rds"))

} else {

 imp_prot_only_tryp <- readRDS(here("rds/missforest_imp_mat_prot_only_tryp.Rds"))
 t_mat_imp_prot_only_tryp <- imp_prot_only_tryp$ximp

 mat_imp_prot_only_tryp <- t(t_mat_imp_prot_only_tryp)

}

```

__Run limma on this new data to generate log2 FCs of proteins from only fully-tryptic peptides__

```{r}
split <- str_split_fixed(colnames(mat_imp_prot_only_tryp), pattern = "\\_", n = 2)

patient_tryp <- split[,1]
recurrence <- split[,2]

design_tryp <- model.matrix(~patient_tryp+recurrence)

rownames(design_tryp) <- colnames(mat_imp_prot_only_tryp)
```

```{r}
rownames(design_tryp) 
```

```{r}
colnames(mat_prot_only_tryp)
```

```{r}
fit_limmawo6_tryp <- function(mat, design, method,Limma, prot2gene){
  limmafit <- lmFit(mat, design, method = method)
  limmafit <- eBayes(limmafit)
  limma_tab <- topTable(limmafit, coef = "recurrencerec", number = Inf, adjust.method = "BH") %>%
  mutate(Protein = rownames(.),
         Limma = Limma)

  limma_tab <- left_join(limma_tab, prot2gene)

  return(limma_tab)
}
```

```{r}
limma_tab_wo6tryp <- fit_limmawo6_tryp(mat_imp_prot_only_tryp, 
                         design_tryp, 
                         method = 'robust', 
                         Limma = "Robust - Proteins w only tryptic peptides",
                         prot2gene = prot2gene)

```

```{r}
log2_fully_tryptic <- dplyr::select(limma_tab_wo6tryp,
                                   logFC, Protein, Gene) %>%
  dplyr::rename(logFC_fully_tryp_protein = logFC) 
```

#### Merge log2FCs of semi-tryptic peptides vs proteins with fully tryptic peptides

```{r}
log2semipept2_log2protein_spec <- left_join(log2_fully_tryptic, log2_semi_tryptic) %>%
  na.omit() %>%
  mutate(DA_Protein = if_else(Protein %in% tab_sig_prots_tryptic$protein,
                              true = TRUE,
                              false = FALSE),
         DA_peptide = if_else(peptide %in% recwo6_pept22$peptide,
                              true = TRUE,
                              false = FALSE))

```

#### log2FC Proteins vs Semi-tryp peptides (scatter plot)

```{r}
inter_scater <- log2semipept2_log2protein_spec %>%
  filter(Protein %in% check_increased_pept$protein_id)

inter_scater$Gene[duplicated(inter_scater$Gene)] <- NA

#geom_point(data = dataarg %>% filter(logFC < -FC_cutoff,
#                                           adj.P.Val < pval_cutoff),
#                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "red") +
#      geom_point(data = dataarg %>% filter(logFC > FC_cutoff,
#                                           adj.P.Val > pval_cutoff),
#                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "#2a9d8f") +

scatter_proteolysis_plot <- ggplot(log2semipept2_log2protein_spec, 
                                   aes(x = logFC_fully_tryp_protein, 
                                       y = logFC_semi_peptide)) + 
  geom_smooth(method=lm, se = FALSE, linetype="dashed", size = 0.5, 
              color = "black") + 
  geom_point(aes(color = DA_Protein, shape = DA_peptide), size = 0.6) +
  scale_color_manual(values = c("#2a9d8f", "red")) +
  xlab("log2(FC) - Protein abundances") + 
  ylab("log2(FC) - Semi-specific peptides") + 
  ggrepel::geom_text_repel(data = inter_scater,
                               aes(label = Gene), 
                           size = 1,
                           box.padding = 0.5,
                           max.overlaps = 25) +
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 8, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 8),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")

cor.test(log2semipept2_log2protein_spec$logFC_fully_tryp_protein,
    log2semipept2_log2protein_spec$logFC_semi_peptide, method = "pearson")
```

```{r}
ggsave(plot = scatter_proteolysis_plot, 
       filename = here::here("figures/scatter_proteolysis_plot_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 63)

ggsave(plot = scatter_proteolysis_plot, 
       filename = here::here("figures/scatter_proteolysis_plot_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 63,
       height = 63)
```

## Analysis of differential amino acid usage (iceLogo)

```{r}
source(here("scr/get_cleavage_area.R"))
```

```{r}
cleave_areas <- get_cleave_area(cleavage_annoated_peptides)
```

```{r}
library(dagLogo)
```

**Load background proteome**

Used the identified protein sequences as background proteome.

```{r}
proteome_ided <- prepareProteome(fasta = here("data/semi_specific_search_fragpipe17/protein_combined.fas"), 
                                 species = "Homo sapiens")
```

```{r}
increased_Rec <- filter(cleave_areas$cleave_area20,
                            Peptide %in% increased_recwo6_pept)

decreased_Rec <- filter(cleave_areas$cleave_area20,
                            Peptide %in% decreased_recwo6_pept)

write_tsv(x = increased_Rec,
          file = here("results/semi_tryptic/proteolytic_products_increased_in_Recurrent.tsv"))

write_tsv(x = decreased_Rec,
          file = here("results/semi_tryptic/proteolytic_products_increased_in_Recurrent.tsv"))

increased_Rec_4ice <- filter(cleave_areas$cleave_area20,
                            Peptide %in% increased_recwo6_pept) %>%
                    pull(cleave_area20)

decreased_Rec_4ice <- filter(cleave_areas$cleave_area20,
                            Peptide %in% decreased_recwo6_pept) %>%
                    pull(cleave_area20)
```

**Format peptide sequences**

```{r}
if(!file.exists(here("results/semi_tryptic/formated_pept_daglogo_increasedREC.rds"))){

  form_peptidesincreased_4ice <- formatSequence(increased_Rec_4ice, 
                                  proteome = proteome_ided)

  write_rds(form_peptidesincreased_4ice, file = here("results/semi_tryptic/formated_pept_daglogo_increasedREC.rds"))
} else {

  form_peptidesincreased_4ice <- read_rds(here("results/semi_tryptic/formated_pept_daglogo_increasedREC.rds"))

}
```

```{r}
if(!file.exists(here("results/semi_tryptic/formated_pept_daglogo_decreasedREC.rds"))){

  form_peptidesdecreased_4ice <- formatSequence(decreased_Rec_4ice, 
                                  proteome = proteome_ided)

  write_rds(form_peptidesdecreased_4ice, file = here("results/semi_tryptic/formated_pept_daglogo_decreasedREC.rds"))
} else {

  form_peptidesdecreased_4ice <- read_rds(here("results/semi_tryptic/formated_pept_daglogo_decreasedREC.rds"))

}
```

#### DAU increased

```{r}
bg_mod_ztest_increased <- buildBackgroundModel(form_peptidesincreased_4ice,
                                           proteome = proteome_ided,
                           background = "wholeProteome",
                           testType = "ztest")
```

```{r}
dau_nogroup_increased <- testDAU(form_peptidesincreased_4ice, 
                           dagBackground = bg_mod_ztest_increased)
```

```{r}
dagHeatmap(dau_nogroup_increased) 
```

#### DAU decreased

```{r}
bg_mod_ztest_decreased <- buildBackgroundModel(form_peptidesdecreased_4ice,
                                           proteome = proteome_ided,
                           background = "wholeProteome",
                           testType = "ztest")
```

```{r}
dau_nogroup_decreased <- testDAU(form_peptidesdecreased_4ice, 
                           dagBackground = bg_mod_ztest_decreased)
```

```{r}
dagHeatmap(dau_nogroup_decreased) 
```

# Proteogenomics

## Initial data loading and wrangling

Load id/quant data and prep coverage information

```{r message=FALSE}
## tmt-integrator output loading 
tmt_protdata_frag_progen <- read_tsv(here("data/proteogenomics/sras_cellreports_2021_paper_razor/tmt-report/abundance_protein_MD.tsv")) %>% 
  janitor::clean_names()

tmt_peptdata_frag_progen <- read_tsv(here("data/proteogenomics/sras_cellreports_2021_paper_razor/tmt-report/abundance_peptide_MD.tsv")) %>% 
  janitor::clean_names()

identified_proteins_progen <- read_tsv(here("data/proteogenomics/sras_cellreports_2021_paper_razor/combined_protein.tsv")) %>% 
  janitor::clean_names()

identified_peptides_progen <- read_tsv(here("data/proteogenomics/sras_cellreports_2021_paper_razor/combined_peptide.tsv")) %>% 
  janitor::clean_names()

```

```{r}
# wide expression matrix with NAs
expr_matrix_progen <- dplyr::select(tmt_protdata_frag_progen,
                             Protein = index, 
                             starts_with("x"))

long_mat_progen <- pivot_longer(expr_matrix_progen,
                         cols = starts_with("x"),
                         names_to = c("paired_id_semi"),
                         values_to = "Abundance") %>%
  mutate(recurrence = str_extract(paired_id_semi, "[^0-9]+$")) %>%
  mutate(patient = str_remove_all(paired_id_semi, "[^0-9]+$"))

long_mat_progen_2 <- long_mat_progen %>%
  mutate(paired_id = paste(patient, recurrence, sep = "_"))

paired_annot_uniq_progen <- long_mat_progen_2 %>%
  dplyr::select(patient, recurrence, paired_id) %>%
  distinct()

# wide peptide abundace matrix with NAs
expr_matrix_pept_progen <- dplyr::select(tmt_peptdata_frag_progen,
                             index, peptide, starts_with("x")) %>%
  dplyr::select(!starts_with("x6"))

expr_matrixnona_pept_progen <- expr_matrix_pept_progen %>%
  na.omit()

long_mat_pept_progen <- pivot_longer(expr_matrix_pept_progen,
                         cols = starts_with("x"),
                         names_to = c("paired_id_semi"),
                         values_to = "Abundance") %>%
  mutate(recurrence = str_extract(paired_id_semi, "[^0-9]+$")) %>%
  mutate(patient = str_remove_all(paired_id_semi, "[^0-9]+$"))

long_mat_pept_progen_2 <- long_mat_pept_progen %>%
  mutate(paired_id = paste(patient, recurrence, sep = "_"))

paired_annot_uniq_pept_progen <- long_mat_pept_progen_2 %>%
  dplyr::select(patient, recurrence, paired_id) %>%
  distinct()

## Annotated abundance data in long format ----

quant_annotated_progen <- left_join(long_mat_progen_2,
                             sample_annotation2) %>%
  mutate(Channel_mix = paste(mixture,channel,  sep = "_"))

quant_annotated_pept_progen <- left_join(long_mat_pept_progen_2,
                             sample_annotation2) %>%
  mutate(Channel_mix = paste(mixture,channel,  sep = "_"))

## Wide xpression matrices without NAs
expr_matnona_progen <- column_to_rownames(expr_matrix_progen,
                            "Protein") %>%
          as.matrix() %>%
          na.omit()

# peptide to protein mapping from combined_peptides.tsv 
protein2peptideall <- dplyr::select(identified_peptides_progen,
                                    protein, 
                                    Peptide = sequence) %>% 
          separate(col = protein,
                   into = c("sp", "Genes", "ensembl"),
                   sep = "\\|") %>%
          dplyr::select(Genes, Peptide) %>%
          distinct() %>%
          mutate(Peptide = str_remove_all(Peptide, "\\[.+?\\]")) %>%
          mutate(Peptide = str_remove_all(Peptide, "^n"))

# peptide to protein mappong from tmt abundance report (quantified and normalized)
protein2peptide_tmtrept <- tmt_peptdata_frag_progen %>%
  dplyr::select(index, protein_id, peptide)

```

## Localization of identified peptides and mapping to variant regions

```{r}
# list of identified peptides associated with 'variant' proteins
var_prots <- protein2peptideall %>%
                    mutate(variant = if_else(str_detect(Genes, "\\_"),
                           true = "variant",
                           false = "canonical")) %>%
                    filter(variant == "variant") %>% 
                    pull(Genes) %>%
                    unique()
```

```{r echo=TRUE}
variant_peptides_1 <- filter(protein2peptideall,
                             Genes %in% var_prots)

# prep fasta containing only identified protein sequences
fastavar <- seqinr::read.fasta(here("data/proteogenomics/sras_cellreports_2021_paper_razor/mix_1/protein.fas"), 
                            seqtype = "AA", 
                            as.string = TRUE)

fastavar2 <- seqinr::read.fasta(here("data/proteogenomics/sras_cellreports_2021_paper_razor/mix_2/protein.fas"), 
                            seqtype = "AA", 
                            as.string = TRUE)

fastavar3 <- seqinr::read.fasta(here("data/proteogenomics/sras_cellreports_2021_paper_razor/mix_3/protein.fas"), 
                            seqtype = "AA", 
                            as.string = TRUE)

fasta1 <- append(fastavar,
                fastavar2)

fasta <- append(fasta1,
                fastavar3)

fasta <- discard(fasta,
                 duplicated(names(fasta)))

```

```{r echo=TRUE}
if(!file.exists(here("results/proteogenomics/cellreps_2021/cleavage_annoated_peptides_razor.tsv"))){
  cleavage_annoated_peptides <- annotate_peptides(expr_mat = variant_peptides_1, 
                                                  fasta = fasta,
                                                  decoy_tag = "rev_")

  write_tsv(cleavage_annoated_peptides, "results/proteogenomics/cellreps_2021/cleavage_annoated_peptides_razor.tsv")
} else {
  cleavage_annoated_peptides <- read_tsv(here("results/proteogenomics/cellreps_2021/cleavage_annoated_peptides_razor.tsv"))
}

```

### Localized peptides within protein sequence and confirm identification of SAAVs

```{r echo=TRUE}
# tabulate location of the variant
annotated_w_variant_location <- cleavage_annoated_peptides %>%
          as_tibble() %>%
          mutate(protein_id_var = protein_id,
                 variant_type = if_else(str_detect(protein_id, "\\:"), 
                                        true = "other",
                                        false = "saav")) %>%
          mutate(protein_id = str_replace_all(protein_id, "\\.", "\\_")) %>%
          mutate(protein_id = str_replace_all(protein_id, "\\:", "\\_")) %>%
          separate(protein_id, 
                   into = c("ense_id", "variant_location"), 
                   sep = "\\_",
                   extra = "merge") %>%
                    na.omit() %>%
          mutate(variant_type = if_else(condition = str_detect(variant_location, "[0-9]$"),
                                        true = "other",
                                        false = variant_type))

# generate a long list of variant locations

locat2prot <- str_split(annotated_w_variant_location$variant_location,
                        pattern = "\\_")

names(locat2prot) <- annotated_w_variant_location$ense_id %>%
                    na.omit()

locat2protdf <- list2DF(locat2prot)

long2prot <- locat2protdf %>%
          pivot_longer(everything(),
                       values_to = "variant_location",
                       names_to = "ense_id") %>% 
          distinct() %>%
          mutate(variant_num_location = readr::parse_number(variant_location))

long2prot_worivar <- left_join(long2prot, 
                               dplyr::select(annotated_w_variant_location,
                                             protein_id_var, ense_id)) %>%
  distinct()

annotated_wo_variant_location <- annotated_w_variant_location %>%
          dplyr::select(-variant_location)

# merge long table of variant locations with the annotation of peptide position within 
# their associated protein sequences

protein2varlocat <- left_join(annotated_wo_variant_location, long2prot_worivar) %>%
          distinct() %>%
          mutate(start_num = as.numeric(start_position),
                 end_num = as.numeric(end_position)) %>%
          mutate(variant_peptide = variant_num_location >= as.numeric(start_position) & variant_num_location <= as.numeric(end_position)) %>% # evaluate if the falls on the position of a variant sequence
          dplyr::select(protein_id_var, ense_id, protein_description, Peptide, start_position, 
                        start_num,end_position,end_num, variant_location, 
                        variant_num_location, variant_peptide, variant_type)

pept2variant <- protein2varlocat %>%
  dplyr::select(Peptide, variant_peptide, variant_type) %>%
  distinct() %>%
  na.omit() %>%
  filter(variant_type == "saav")
```

```{r echo=TRUE}
table(pept2variant$variant_peptide)
```  
232 peptides were identified as mapping to single amino acid variant regions of their corresponding annotated proteins sequence.

## Visualize quantitative features of identified SAAVs

### Prep abundance matrices

```{r}
pep_prematwo6 <- dplyr::select(expr_matrix_pept_progen,
                               index, peptide, rownames(design_limmawo6))

pep_matwo6 <- dplyr::select(expr_matrix_pept_progen,
                             index, rownames(design_limmawo6))  %>%
  column_to_rownames("index") %>%
  as.matrix()

pep_matnona_wo6 <- pep_matwo6 %>%
  na.omit()

```

### Filter, merge and summarize SAAV abundances

```{r}
# extract indexing, peptide and sequence information from normalized data
index2genepeptide <- tmt_peptdata_frag_progen %>% 
  dplyr::select(index, peptide, gene, protein_id)

# table of interesting features (saavs)
features1 <- pept2variant %>%
  dplyr::rename(peptide = Peptide) %>%
  left_join(., index2genepeptide) %>%
  na.omit()

interesting_features1 <- features1 %>%
                    filter(variant_peptide == TRUE)

# data frame of identified SAAVs and their annotation
protein2varlocat_saav <- protein2varlocat %>%
  dplyr::rename(peptide = Peptide) %>%
  filter(peptide %in% interesting_features1$peptide,
         variant_peptide == TRUE) %>%
  left_join(., index2genepeptide) %>%
  mutate(gene_variant = paste(gene, variant_location, sep = "_"))

# mapping variant to index id
genevariant2index <- protein2varlocat_saav %>%
  dplyr::select(index, gene_variant)

# abundance matrix of peptides mapping to variant locations
pep_premat_variantwo6 <- filter(pep_prematwo6,
                                peptide %in% protein2varlocat_saav$peptide) %>%
  left_join(., genevariant2index) %>%
  relocate(gene_variant)

# summarize SAAVs (gene+variant location)
# i.e. take the median abundance of all the peptides mapping to the same SAAV

pep_premat_sumgenarwo6 <- pep_premat_variantwo6 %>%
  group_by(gene_variant) %>%
  summarise_at(vars(starts_with("x")), .funs = function(x){median(x, na.rm = TRUE)}) %>%
  ungroup()

full_info_sumgenvar <- pep_premat_sumgenarwo6 %>%
  left_join(., protein2varlocat_saav) %>%
  filter(variant_type == "saav")

write_tsv(protein2varlocat_saav,
          here("results/proteogenomics/cellreps_2021/all_peptides_mapping_to_saavs_gbm.tsv"))

write_tsv(pep_premat_sumgenarwo6,
          here("results/proteogenomics/cellreps_2021/summarized_median_abundance_of_saavs.tsv"))

write_tsv(full_info_sumgenvar,
          here("results/proteogenomics/cellreps_2021/summarized_median_abundance_of_saavs_full.tsv"))

pep_matvariantswo6or <- pep_premat_sumgenarwo6 %>%
  column_to_rownames("gene_variant") %>%
  as.matrix()

pep_matvariantswo6 <- pep_premat_sumgenarwo6 %>%
  filter(str_detect(gene_variant, "CAP1_", 
                    negate = TRUE)) %>%
  column_to_rownames("gene_variant") %>%
  as.matrix()

pep_matvariantswo6nona <- pep_matvariantswo6 %>%
  na.omit()
```

The variant-containing peptides identified are able to identify `r nrow(pep_premat_sumgenarwo6)` single amino acid mutations (SAAVs).

Among these identified SAAVs, `r nrow(pep_matvariantswo6nona)` were identified in all mixtures and samples.

### Heatmap of all SAAVs observed

```{r}
library(circlize)
```

```{r}
library(ComplexHeatmap)
```

```{r}
split_patstage <- str_split_fixed(string = colnames(pep_matvariantswo6nona), 
                                  pattern = "[:digit:]+", 
                                  n = 2)

patient <- str_extract(string = colnames(pep_matvariantswo6nona),
                       pattern = "[:digit:]+")

patientstage <- colnames(pep_matvariantswo6nona) %>%
  str_replace(., pattern = "prim", 
              replacement = "Init") %>%
  str_replace(., pattern = "rec", 
              replacement = "Rec")

stage <- split_patstage[,2] %>%
  if_else(condition = . == "prim",
          true = "Initial",
          false = "Recurrent")

pep_matvariantswo6or_hm <- pep_matvariantswo6or
pep_matvariantswo6nona_hm <- pep_matvariantswo6nona

colnames(pep_matvariantswo6or_hm) <- patientstage
colnames(pep_matvariantswo6nona_hm) <- patientstage
```

```{r fig.height=20, fig.width=7}
col_fun_or = colorRamp2(c(min(pep_matvariantswo6or_hm, na.rm = TRUE), 
                       median(pep_matvariantswo6or_hm, na.rm = TRUE), 
                       max(pep_matvariantswo6or_hm, na.rm = TRUE)), 
                     c("#2a9d8f", "white", "red"))

Heatmap(pep_matvariantswo6or_hm, 
        col = col_fun_or,
        column_split = stage, 
        row_names_gp = gpar(fontsize = 5.5),
        column_names_gp = gpar(fontsize = 7),
        cluster_rows = FALSE, cluster_columns = FALSE,
        heatmap_legend_param = list(
        title = "log2-\nNormalized\nAbundance"
    ))
```

### Heatmap of SAAVs observed on every sample

```{r fig.height=7.5, fig.width=7}
col_fun_sel = colorRamp2(c(min(pep_matvariantswo6or_hm, na.rm = TRUE), 
                       median(pep_matvariantswo6or_hm, na.rm = TRUE), 
                       max(pep_matvariantswo6or_hm, na.rm = TRUE)), 
                     c("#2a9d8f", "white", "red"))

Heatmap(pep_matvariantswo6nona_hm, 
        col = col_fun_sel,
        column_split = stage, 
        row_names_gp = gpar(fontsize = 7),
        column_names_gp = gpar(fontsize = 7),
        heatmap_legend_param = list(
        title = "log2-\nNormalized\nAbundance"
    ))
```

### Limma on SAAVs

### Prep design matrix

```{r}
design_limmawo6_saav <- model.matrix(~patient+stage)

sample_annotationwo6 <- sample_annotationwo6 %>%
  mutate(paired_id_semi = paste(patient, recurrence, sep = ""))

rownames(design_limmawo6_saav) <- sample_annotationwo6$paired_id_semi
```

### Prep abundance matrices

```{r}
colnames(pep_matvariantswo6nona)
```

```{r}
rownames(design_limmawo6_saav)
```

#### SAAV-level model fit

```{r}
fit_limmapeptwo6 <- function(mat, design, method,Limma){
  limmafit <- lmFit(mat, design, method = method)
  limmafit <- eBayes(limmafit)
  limma_tab <- topTable(limmafit, coef = "stageRecurrent", number = Inf, adjust.method = "BH") %>%
  mutate(Protein = rownames(.),
         Limma = Limma)

  return(limma_tab)
}
```

```{r echo=TRUE, warning=FALSE, message=FALSE}
fit_mod_saavs <- fit_limmapeptwo6(pep_matvariantswo6nona, 
                                  design_limmawo6_saav, 
                                  method = 'robust', 
                                  Limma = "Robust - w Patient effect")

compar_tab_saavs <-  fit_mod_saavs %>% 
  rownames_to_column("gene_variant") 
```

```{r}
increased_recwo6_saavs <- compar_tab_saavs %>%
                    filter(logFC>0,
                           adj.P.Val < 0.05) %>% 
  pull(gene_variant)

decreased_recwo6_saavs <- compar_tab_saavs %>%
                    filter(logFC<0,
                           adj.P.Val < 0.05) %>% 
  pull(gene_variant)

```

```{r}
length(increased_recwo6_saavs)
```
```{r}
length(decreased_recwo6_saavs)
```

#### Volcano SAAVs

```{r fig.width=8, fig.height=6}
diff_saavs <- compar_tab_saavs %>%
                    filter(adj.P.Val < 0.05)

size <- 1.5

volcano_saavs <- ggplot(data = compar_tab_saavs,
                      mapping = aes(x = logFC, y = -log10(adj.P.Val))) +
      geom_point(data = compar_tab_saavs %>% filter(logFC > 0,
                                           adj.P.Val < 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), 
                 color = "red",
                 size = size)+
      geom_point(data = compar_tab_saavs %>% filter(logFC < -0,
                                           adj.P.Val < 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "red",
                 size = size) +
      geom_point(data = compar_tab_saavs %>% filter(logFC > 0,
                                           adj.P.Val > 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "#2a9d8f",
                 size = size) +
      geom_point(data = compar_tab_saavs %>% filter(logFC < -0,
                                           adj.P.Val > 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "#2a9d8f",
                 size = size) +
      geom_hline(yintercept = -log10(0.05),
                 color = "red", linetype = "dashed") +
      ggrepel::geom_text_repel(data = diff_saavs,
                               aes(label = gene_variant), size = 4) +
      xlab("logFC - Recurrent / Primary") + 
      theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 6, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 8),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom") 
```

```{r echo=FALSE}
ggsave(plot = volcano_saavs, 
       filename = here::here("figures/volcanosaavs_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 50)

ggsave(plot = volcano_saavs, 
       filename = here::here("figures/volcanosaavs_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 63,
       height = 50)
```

### Boxplot of differentially abundant SAAV

```{r}
prelongsaav <- pep_matvariantswo6nona %>%
  as.data.frame() %>%
  rownames_to_column("gene_variant") %>%
   pivot_longer(cols = starts_with("x"),
                         names_to = c("patient"),
                         values_to = "Abundance") %>%
  mutate(Stage = str_extract(patient, "[^0-9]+$")) %>%
  mutate(Patient = str_remove_all(patient, "[^0-9]+$")) %>%
  filter(gene_variant == increased_recwo6_saavs)
```

```{r}
diff_abund_saav <- ggplot(prelongsaav, 
                    aes(x = Stage, 
                        y = Abundance, fill = Stage, 
                        cex.axis = 1.5)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 0.2) +
  # Box plot with jittered points
  # 0.2 : degree of jitter in x direction
  # geom_jitter(shape=16, position=position_jitter(0.2))
  ylab("log2-Normalized Abundance") +
  geom_signif(
    comparisons = list(c("Initial", "Recurrence")),
    map_signif_level = TRUE
  ) + 
  stat_compare_means(method="t.test") +
  ggtitle(increased_recwo6_saavs) +
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 6, angle = 360),
        axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 8),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        axis.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")

```

### Boxplot DA SAAV

```{r}
  slim_data2saav <- pivot_wider(prelongsaav %>% dplyr::select(-patient), 
                          values_from = Abundance, 
                          names_from = Stage) #%>%
  #dplyr::select(-patient)

  slim_data3saav <- slim_data2saav %>% 
    mutate(Abs_diff = rec-prim) %>%
    mutate(rank = row_number(Abs_diff)) %>% 
    dplyr::rename(Initial = prim, Recurrent = rec) %>%
    ungroup()

```

```{r fig.width=7, fig.height=3.5}
ggpaired(slim_data3saav,
         cond1 = "Initial",
         cond2 = "Recurrent", 
         fill = "condition",
         ylab = "log2-Normalized Abundance",
         label = NULL,
         repel = TRUE, point.size = 0.2, line.size = 0.1) +
      theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 5, angle = 360),
            axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 5),
            panel.background = element_blank(),
            panel.grid.major = element_blank(),
            panel.border = element_rect(colour = "black", fill=NA, size=0.5),
            axis.title = element_text(size = 8),
            legend.text = element_text(size = 6),
            legend.title = element_text(size = 8),
            legend.key.height= unit(3, 'mm'),
            legend.key.width= unit(3, 'mm'),
            legend.position="bottom",
            strip.text = element_text(hjust = 0.5, vjust = 0, size = 8, angle = 360)) + 
  ggtitle(increased_recwo6_saavs)
```

```{r}
paired_box_saavs <- ggpaired(slim_data3saav,
         cond1 = "Initial",
         cond2 = "Recurrent", 
         fill = "condition",
         ylab = "log2-Normalized Abundance",
         label = NULL,
         repel = TRUE, point.size = 0.2, line.size = 0.1) +
      theme(axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 5, angle = 360),
            axis.text.y = element_text(hjust = 0.95, vjust = 0.2, size = 5),
            panel.background = element_blank(),
            panel.grid.major = element_blank(),
            panel.border = element_rect(colour = "black", fill=NA, size=0.5),
            axis.title = element_text(size = 8),
            legend.text = element_text(size = 6),
            legend.title = element_text(size = 8),
            legend.key.height= unit(3, 'mm'),
            legend.key.width= unit(3, 'mm'),
            legend.position="bottom",
            strip.text = element_text(hjust = 0.5, vjust = 0, size = 8, angle = 360)) 
```

```{r echo=FALSE}
ggsave(plot = paired_box_saavs, 
       filename = here::here("figures/paired_box_saav_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 61,
       height = 50)

ggsave(plot = paired_box_saavs, 
       filename = here::here("figures/paired_box_saav_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 61,
       height = 50)
```

```{r}
sessionInfo()
```