### 2022-12-15 tables/elisa_plasma_proteomics/supp_table_2.tex

\setlength{\LTpost}{0mm}
\begin{longtable}{llrrrr}
\caption*{
{\large \textbf{Statistical Comparisons Between Population Groups per Genes}}
} \\ 
\toprule
\textbf{Comparison} & \textbf{Statistical Test}\textsuperscript{1} & \textbf{ASAH1}\textsuperscript{2} & \textbf{GPNMB}\textsuperscript{2} & \textbf{MMP-9}\textsuperscript{2} & \textbf{SYNEMIN}\textsuperscript{2} \\ 
\midrule
\multicolumn{6}{l}{\textbf{Control vs Pre-Surgery vs Post-Surgery Plasma Samples}} \\ 
\midrule
Ctrl vs post-
surgery & Permutation Test & 0.7485000 & 0.1685000 & 0.1231000 & 0.2574000 \\ 
Ctrl vs pre-
surgery & Permutation Test & 0.0208000 & 0.9939500 & 0.0411000 & 0.0617000 \\ 
pre-
surgery vs post-
surgery & Paired Permutation Test & 0.0089720 & 0.2421880 & 0.4755860 & 0.5427860 \\ 
\midrule
\multicolumn{6}{l}{\textbf{Control vs Initial GBM vs Recurrent GBM Plasma Samples}} \\ 
\midrule
Ctrl vs iGBM & Permutation Test & 0.1836500 & 0.3994500 & 0.0830000 & 0.0022500 \\ 
Ctrl vs rGBM & Permutation Test & 0.2484000 & 0.2455000 & 0.4659500 & 0.1344500 \\ 
iGBM vs rGBM & Paired Permutation Test & 0.1992188 & 0.6123047 & 0.1171875 & 0.1054688 \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
\textsuperscript{1}Permutation test using t-test statisics for non-paired samples, or paired permutation test using pair t-test statistics for paired samples.\\
\textsuperscript{2}Values in cells represent the non-adjusted p-value from the corresponding statistical test.\\
\end{minipage}