### 2023-02-14 reports/gbm_recurrence_general_proteomics_manuscript_report.qmd
proteomics report clean

---
title: "GBM recurrence - Shotgun proteomics analysis"
author: "Miguel Cosenza-Contreras"
format:
  gfm: 
    toc-depth: 4 # default is 3
    toc-title: Contents
    toc-location: left # default is right, option is 'body'
    number-sections: true
    number-depth: 3
    embed-resources: true
always_allow_html: true
editor: visual
---

```{r knirt_helper_funcs}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

source(here::here("scr/helper_functions.R"))
```

```{r load_packages}
## Required packages ----
library(tidyverse)
library(mixOmics)
library(fs)
library(kableExtra)
library(sva)
library(limma)
library(naniar)
library(missForest)
library(clusterProfiler)
library(ReactomePA)
library(org.Hs.eg.db)
library(DT)
library(here)
library(janitor)
library(drawProteins)
library(seqinr)
library(ggpubr)
library(ggrepel)
library(extrafont)
```

```{r theme_set}
theme_set(theme(axis.text.x = element_text(hjust = 0.5, 
                                           vjust = 0, 
                                           size = 6, 
                                           angle = 360),
                axis.text.y = element_text(hjust = 0.5, 
                                           vjust = 0, 
                                           size = 6),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", 
                                            fill = NA, 
                                            size = 1.5),
                axis.title=element_text(size = 8),
                legend.title = element_text(size = 8),
                legend.key.height = unit(3, 
                                         'mm'),
                legend.key.width = unit(3, 
                                        'mm'),
                legend.position = "bottom"))
```

The following report holds the code, results and interpretation of the proteomics data analysis of Initial-Recurrent glioblastoma (GBM) samples of matched patients.

# Sample annotation info

```{r echo=TRUE}
sample_annotation <- read_csv(here("data/sample_annotation.csv"))

# correct annotation

sample_annotation2 <- sample_annotation %>%
  mutate(patient = paste("x",
                         patient,
                         sep = ""),
         recurrence = case_when(recurrence == "initial" ~ "prim",
                                recurrence == "recurrent" ~ "rec",
                                TRUE ~ recurrence)) %>%
  mutate(paired_id = paste(patient, 
                           recurrence, 
                           sep = "_")) %>%
  filter(recurrence %in% c("prim", 
                           "rec"))

sample_annotationwo6 <- sample_annotation2 %>%
  filter(patient != "x6")
```

## Sample annotation

### Brief note on sample annotation: {.unnumbered .unlisted}

Twenty-two patient matched samples (corresponding to 11 patients) were processed for protein extraction, TMT-labelled and submitted via high-pH fractionation and further analyzed via LC-MS/MS.

```{r echo=FALSE, error=TRUE}
kbl(sample_annotation, booktabs = TRUE) %>%
  kable_styling(font_size = 14, 
                bootstrap_options = c("striped", 
                                      "hover"))
```

# Brief note on data analysis and processing:

Three proteomics data analysis approaches were used to explore the spectral data and will be briefly described:

1.  General large scale proteomics.
2.  Proteolytic processing.
3.  Proteogenomics

The FragPipe bioinformatic pipeline was used for each of these approaches with varying parameters depending on the type of explorative approach.

**This report holds the code, analysis, visualizations and results for the General large scale proteomics approach.**

**1. General large scale proteomics analyses** : In order to explore the differential abundance of proteins between Recurrent and Initial tumor, a fully-tryptic search was performed against the EBI Human canonical proteome (version 2021_03), appended with common contaminants and iRT peptides. A minimum percentage of %5 of the total summed reporter ion intensities were required to consider a peptide and reporter ion for quantitation.

**2. Proteolytic processing**: The large scale evaluation of differential proteolytic activity between recurrent vs primary tumor was executed via a semi-tryptic search against the EBI Human canonical proteome (version 2021_03), appended with common contaminants and iRT peptides. A minimum percentage of %5 of the total summed reporter ion intensities were required to consider a peptide and reporter ion for quantitation.

**3. Proteogenomics**: The explorative proteogenomics analysis started with the generation of a GBM-specific database from publicly available RNA-seq sequencing data from tumor specimens. Fastq files were obtain from the SRA BioProject PRJNA627121 (Cell Rep, 2021 Mar 2;34(9):108787). Eight samples containing paired reads were selected, based on: 1) the higher number of reads per file; 2) having a similar proportion of males and females and 3) not bearing isocitrate-dehyderogenase (IDH)-1 mutation (IDH WT). None of these patients undergo previous treatment, therefore potential variants in the sequencing data are not expected to be associated to treatment-related mutagenic processes. Spectral data was then searched against this custom database using a fully-tryptic approac. No minimum intensity threshold was set for reporter ion intensity for quantitation.

Exceptionally when stated differently within each sub-approach described above, all these searches were executed using the following parameters: MSFragger was used as a search engine, setting a fully or semi-tryptic specificity allowing for 1 missed cleavages. Precursor mass tolerance was set to -20/20 and fragment mass tolerance of 20 ppm with mass calibration and parameter optimization resulting in 5 ppm. Peptide N-terminal acetylation and peptide N-terminal TMT labeling were set as variable modifications. TMT labelling at K and carbamidomethylation of C were set as fixed modifications. MSBooster was used for deep-learning based predictions of retention time and spectra. Predicted features were used by Percolator for post-processing scoring and false discovery rate (FDR) control via target-decoy competition of all peptide-to-spectrum matches (PSMs) obtained from MSFragger search. ProteinProphet was used as a protein inference algorithm keeping all PSMs with probability score bigger than 0.9. Final report and FDR estimation was based on the filtered PSM and protein lists before protein scoring.

Relative quantitation of identified peptides within each sample was performed via their reported ion intensities using TMT-integrator. Only PSMs coming from unique peptides with a minimum probability of 0.9 and an isotopic purity of at least 50% were considered for quantitation. Quantitative values were normalized via median centering and summarized by protein/gene and peptide using a virtual reference channel, to finally generate protein and peptide normalized abundance matrices of abundance used for further statistical processing in combination with the sample annotation.

# Initial data loading and wrangling {.unnumbered .unlisted}

```{r message=FALSE}
## tmt-integrator output loading 
tmt_protdata_frag <- read_tsv(here("data/specific_search_fragpipe17/specific_no_ptms_2/tmt-report/abundance_protein_MD.tsv")) %>% 
  janitor::clean_names()

identified_proteins <- read_tsv(here("data/specific_search_fragpipe17/specific_no_ptms_2/combined_protein.tsv")) %>% 
  janitor::clean_names()

identified_peptides <- read_tsv(here("data/specific_search_fragpipe17/specific_no_ptms_2/combined_peptide.tsv")) %>%
  janitor::clean_names()

fasta_ident_prots <- read.fasta(here("data/specific_search_fragpipe17/specific_no_ptms_2/mix_2/protein.fas"), seqtype = "AA", as.string = TRUE)
```

```{r}
# wide expression matrix with NAs
expr_matrix <- dplyr::select(tmt_protdata_frag,
                             Protein = index, 
                             starts_with("x"))

long_mat <- pivot_longer(expr_matrix,
                         cols = starts_with("x"),
                         names_sep = "_",
                         names_to = c("patient", 
                                      "recurrence"),
                         values_to = "Abundance")

long_mat_2 <- long_mat %>%
  mutate(paired_id = paste(patient, 
                           recurrence, 
                           sep = "_"))

paired_annot_uniq <- long_mat_2 %>%
  dplyr::select(patient, 
                recurrence, 
                paired_id) %>%
  distinct()

# correct annotation

sample_annotation2 <- sample_annotation %>%
  mutate(patient = paste("x", 
                         patient, 
                         sep = ""),
         recurrence = case_when(recurrence == "initial" ~ "prim",
                                recurrence == "recurrent" ~ "rec",
                                TRUE ~ recurrence)) %>%
  mutate(paired_id = paste(patient, 
                           recurrence, 
                           sep = "_")) %>%
  filter(recurrence %in% c("prim", 
                           "rec"))

## Annotated abundance data in long format ----

quant_annotated <- left_join(long_mat_2,
                             sample_annotation2, 
                             by = c("paired_id", 
                                    "patient",
                                    "recurrence")) %>%
  mutate(Channel_mix = paste(mixture, 
                             channel,  
                             sep = "_"))

## Wide xpression matrices without NAs
expr_matnona <- column_to_rownames(expr_matrix,
                                   "Protein") %>%
          as.matrix() %>%
          na.omit()
```

# ID Coverage before sparicity reduction and filtering

These numbers are based on the results from the output of FragPipe's TMT-integrator, before sparicity reduction and filtering.

```{r}
ids_summary <- quant_annotated %>% 
  dplyr::select(Protein, mixture, Abundance) %>% 
  distinct(Protein, mixture, .keep_all = TRUE) %>%
  group_by(mixture) %>%
  summarise(`Nr of IDs` = sum(!is.na(Abundance))) %>% 
  bind_rows(.,
            tibble(mixture = "In all three",
                   `Nr of IDs` = nrow(expr_matnona))) %>%
  bind_rows(.,
            tibble(mixture = "Overall",
                   `Nr of IDs` = nrow(expr_matrix)))
```

-   Number of identified proteins: `r nrow(expr_matrix)`

-   Number of identified and quantified proteins in all samples: `r nrow(expr_matnona)`

```{r fig.height=4, fig.width=7, warning=FALSE}
ggplot(data = ids_summary,
       aes(y = `Nr of IDs`, x = mixture)) +
  geom_col(fill = "#008683") + 
  geom_text(size = 5, 
            position = position_stack(vjust = 0.5), 
            aes(label = `Nr of IDs`)) +
  coord_cartesian(ylim = c(0, 5000))+
  coord_flip() +
  labs(title = "Number of Identified and quantified Proteins by Mixture") +
  ylab("Nr of IDed and Quant Proteins") + 
  xlab("TMT Mixture") + 
  theme(axis.text.x = element_text(hjust = 0.5, 
                                   vjust = 0.1, 
                                   size = 10, 
                                   angle = 0),
        axis.text.y = element_text(hjust = 0.5, 
                                   vjust = 0.1, 
                                   size = 10, 
                                   angle = 0),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey"),
        panel.border = element_rect(colour = "black", 
                                    fill = NA, 
                                    size = 1.5),
        axis.title = element_text(size = 12,
                                face = "bold"))
```

# Exploratory analysis

## Visualization of missing values

```{r fig.height=7, fig.width=8, warning=FALSE}
vis_miss(expr_matrix) + 
          ggtitle("Distribution of missing values") + 
          theme(axis.text.x = element_text(hjust = 0.5, 
                                           vjust = 0.1, 
                                           size = 10, 
                                           angle = 90),
                axis.text.y = element_text(hjust = 0.5, 
                                           vjust = 0.1, 
                                           size = 10),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", 
                                            fill=NA, 
                                            size=1.5),
                axis.title = element_text(size = 12,
                                        face = "bold"))
```

There is an uneven distribution of missingness between TMT-batches.

## Plots of normalized abundance distribution

### Normalized ion reporter abundance distribution by TMT channel/sample {.unnumbered}

```{r fig.height=4, fig.width=7, warning=FALSE}
ggplot(quant_annotated,
                 mapping = aes(x = Channel_mix, 
                               y = Abundance, 
                               fill = recurrence)) +
          geom_boxplot() + 
          theme(axis.text.x = element_text(hjust = 0.5, 
                                           vjust = 0.1, 
                                           size = 10, 
                                           angle = 90),
                axis.text.y = element_text(hjust = 0.5, 
                                           vjust = 0.1, 
                                           size = 10),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", 
                                            fill = NA, 
                                            size = 1.5),
                axis.title = element_text(size = 12,
                                         face = "bold"))
```

The median-centering after TMT-integrator seem to show a slight mixture-dependent centered median per sample. The paired distribution of initial and recurrent tumor allows the biological effect to be more important than any potential TMT-mixture-based batch effect.

The different abundance distribution can also be explained by a different distribution of missing values.

PCA and sPLS-DA analysis allow to confirm this.

### Normalized ion reporter abundance distribution by protein {.unnumbered}

Select/sample *N* proteins to observe their abundance distributio.

```{r}
which_prots <- sample_proteins(x = quant_annotated$Protein,
                               size = 50,
                               seed = 156)
```

`r length(which_prots)` proteins were randomly sampled to check for the normality of their 'normalized' Abundance distribution.

```{r fig.height=4, fig.width=7, warning=FALSE}
ggplot(quant_annotated %>% 
          filter(Protein %in% which_prots),
                 mapping = aes(x = reorder(Protein,
                                           Abundance,
                                           na.rm = TRUE), 
                               y = Abundance)) +
          geom_boxplot() + 
          geom_jitter(color = "black", 
                      size = 0.4, 
                      alpha = 0.9) +
          labs(subtitle = "Protein IDs arranged by median abudance",
               x = "Protein ID ") + 
          theme(axis.text.x = element_text(hjust = 0.5, 
                                           vjust = 0.1, 
                                           size = 10, 
                                           angle = 90),
                axis.text.y = element_text(hjust = 0.5, 
                                           vjust = 0.1, 
                                           size = 10),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", fill=NA, size=1.5),
                axis.title=element_text(size=12,face="bold"))

```

### Normalized ion reporter abundance distribution by protein and batch {.unnumbered}

`r length(which_prots)` proteins were randomly sampled to check for the normality of their 'normalized' Abundance distribution.

```{r fig.height=4, fig.width=9, warning=FALSE}
ggplot(quant_annotated %>% 
                         filter(Protein %in% which_prots),
                 mapping = aes(x = reorder(Protein,
                                           Abundance,
                                           na.rm = TRUE), 
                               y = Abundance, 
                               fill = mixture)) +
          geom_boxplot() + 
          geom_jitter(color = "black", 
                      size = 0.4, 
                      alpha = 0.9) +
          labs(subtitle = "Protein IDs arranged by median abudance",
                       x = "Protein ID ") + 
          facet_grid(.~mixture, 
                     scales = "free") +
          theme(axis.text.x = element_text(hjust = 0.5, 
                                           vjust = 0.1, 
                                           size = 10, 
                                           angle = 90),
                axis.text.y = element_text(hjust = 0.5, 
                                           vjust = 0.1, 
                                           size = 10),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", 
                                            fill = NA, 
                                            size = 1.5),
                axis.title = element_text(size = 12,
                                          face = "bold"))
```

The abundance distribution of individual proteins can be considered as normaly distributed in general, which makes then suitable for inferential statistics via linear models.

## Exploratory PCA

```{r echo=FALSE}
## Get predictors for mixtures
expr_matrix2 <- expr_matrix %>%
  dplyr::select(Protein, 
                sample_annotation2$paired_id)

paired_id <- sample_annotation2$paired_id
recurrence <- sample_annotation2$recurrence
patient <- sample_annotation2$patient
mixture <- sample_annotation2$mixture

## get predictor for mixtures wo patient 6
expr_matrixwo6 <- expr_matrix %>%
  dplyr::select(Protein, 
                sample_annotationwo6$paired_id)

paired_idwo6 <- sample_annotationwo6$paired_id
recurrencewo6 <- sample_annotationwo6$recurrence
patientwo6 <- sample_annotationwo6$patient
mixturewo6 <- sample_annotationwo6$mixture

## Prep matrix without missing values for PCA

expr_matnona <- column_to_rownames(expr_matrix2,
                                   "Protein") %>%
          as.matrix() %>%
          na.omit()

colnames(expr_matnona) <- paste(sample_annotation2$paired_id, 
                                sample_annotation2$mixture,
                                sep = "_")

t_expr_matnona <- t(expr_matnona)

expr_mat <- column_to_rownames(expr_matrix2,
                               "Protein") %>%
          as.matrix()

t_expr_mat <- t(expr_mat)

## Matrix without patient 6

expr_matnonawo6 <- column_to_rownames(expr_matrixwo6,
                                      "Protein") %>%
          as.matrix() %>%
          na.omit()

colnames(expr_matnonawo6) <- paste(sample_annotationwo6$paired_id, 
                                   sample_annotationwo6$mixture,
                                   sep = "_")

t_expr_matnonawo6 <- t(expr_matnonawo6)

expr_matwo6 <- column_to_rownames(expr_matrixwo6,
                                  "Protein") %>%
          as.matrix()

t_expr_matwo6 <- t(expr_matwo6)
```

## Percentage of explained variance per component {.unnumbered .unlisted}

The PCA was applied on the expression matrix without missing values

```{r}
pca_res <- pca(t_expr_matnona, 
               ncomp = 10, 
               center = TRUE, 
               scale = TRUE)
```

```{r fig.width=4.7, fig.height=3.5}
plot(pca_res)
```

\~45% of variance based on protein abundance can be explained by components 1 and 2 after PCA.

## PCA plot on samples based on protein abundance

```{r}
# preprocess pca results 
pca_variates <- pca_res$variates$X %>% 
  as.data.frame() %>%
  rownames_to_column("Sample") %>%
  separate(col = "Sample",
           into = c("sample", 
                    "recurrence", 
                    "mixture"),
           sep = "\\_", 
           extra = "merge") %>%
  mutate(paired_id = paste(sample, 
                           recurrence, 
                           sep = "_")) %>%
  dplyr::select(-c(mixture, 
                   recurrence)) %>%
  left_join(. , sample_annotation2,
            by = "paired_id") %>%
  mutate(Stage = if_else(condition = recurrence == "prim", 
                         true = "Initial", 
                         false = "Recurrent"),
         Mix = case_when(mixture == "tmt_1" ~ "Mix1",
                         mixture == "tmt_2" ~ "Mix2",
                         mixture == "tmt_3" ~ "Mix3"))

pca_variateswopat6 <- pca_variates %>% 
  filter(patient != "x6")
```

```{r fig.width=5, fig.height=4, warning=FALSE}
ggplot(data = pca_variates,
       aes(x = PC1, 
           y = PC2)) +
  geom_point(size = 4, aes(shape = mixture, 
                           color = recurrence)) +  
  geom_text(aes(label = patient), 
            position = position_nudge(x = 4.5)) +
  labs(title = "PCA plot of samples\nBased on the abundance values of proteins") +
  xlab(paste("PC1", 
             round(pca_res$prop_expl_var$X[1]*100), 
             "% var explained")) + 
  ylab(paste("PC2", 
             round(pca_res$prop_expl_var$X[2]*100), 
             "% var explained")) + 
  theme(axis.text.x = element_text(hjust = 0.5, 
                                   vjust = 0, 
                                   size = 10, 
                                   angle = 360),
        axis.text.y = element_text(hjust = 0.95, 
                                   vjust = 0.2, 
                                   size = 10),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", 
                                    fill=NA, 
                                    size=0.5),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")
```

There is no evident batch effect after PCA analysis.

There is a good number of individual patients that show important differences in their PCA variates values in recurrent vs primary.

```{r fig.width=5.8, fig.height=4, warning=FALSE}
ggplot(data = pca_variateswopat6,
       aes(x = PC1, y = PC2)) +
  geom_point(size = 4, aes(shape = mixture, 
                           color = recurrence)) +  
  geom_text(aes(label = patient), 
            position = position_nudge(x = 4.5)) +
  labs(title = "PCA plot of samples\nBased on the abundance values of proteins (patient 6 excluded)") +
  xlab(paste("PC1", 
             round(pca_res$prop_expl_var$X[1]*100), 
             "% var explained")) + 
  ylab(paste("PC2", 
             round(pca_res$prop_expl_var$X[2]*100), 
             "% var explained")) + 
  theme(axis.text.x = element_text(hjust = 0.5, 
                                   vjust = 0, 
                                   size = 10, 
                                   angle = 360),
        axis.text.y = element_text(hjust = 0.95, 
                                   vjust = 0.2, 
                                   size = 10),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", 
                                    fill=NA, 
                                    size=0.5),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")
```

```{r pca_proteomics_figure}
pca_plot <- ggplot(data = pca_variateswopat6,
       aes(x = PC1, 
           y = PC2)) +
  geom_point(size = 4, aes(shape = Mix, 
                           color = Stage)) +  
  geom_text(aes(label = patient), 
            position = position_nudge(x = 4.5), 
            size = 2) +
  xlab(paste("PC1", 
             round(pca_res$prop_expl_var$X[1]*100), 
             "% var explained")) + 
  ylab(paste("PC2", 
             round(pca_res$prop_expl_var$X[2]*100), 
             "% var explained")) +
theme(axis.text.x = element_text(hjust = 0.5, 
                                 vjust = 0, 
                                 size = 7,
                                 angle = 360),
      axis.text.y = element_text(hjust = 0.95, 
                                 vjust = 0.2,
                                 size = 7),
      panel.background = element_blank(),
      panel.grid.major = element_blank(),
      panel.border = element_rect(colour = "black", 
                                  fill=NA, 
                                  size=0.5),
      axis.title = element_text(size = 7),
      legend.text = element_text(size = 7),
      legend.title = element_text(size = 8),
      legend.key.height= unit(3, 'mm'),
      legend.key.width= unit(3, 'mm'),
      legend.position="bottom")

ggsave(plot = pca_plot, 
       filename = here::here("figures/pca_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 50)

ggsave(plot = pca_plot, 
       filename = here::here("figures/pca_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 63,
       height = 50)
```

## Exploratory sPLS-DA

```{r}
plsda_rest1 <- plsda(X = t_expr_matnona,
                     Y = recurrence,
                     ncomp = 3)
```

```{r fig.width=5, fig.height=4}
plotIndiv(plsda_rest1,
          nd.names = TRUE, 
          ellipse = TRUE, 
          legend = FALSE, 
          title = "sPLS-DA_Rec-Prim", 
          size.title = rel(1.5), 
          size.xlabel = rel(1.5), 
          size.ylabel = rel(1.5), 
          siz.axis = rel(1.2), 
          point.lwd = 0.8, 
          cex = 3)  
```

Patient 6 in recurrence behaves suspiciously similar to initial samples. We consider this to be an outlier associated to a potential problem during sample prep.

Therefore we decided to exclude patient 6 from inferential analyses.

### sPLS-DA excluding 'patient 6' {.unnumbered .unlisted}

```{r}
plsda_restwo6 <- plsda(X = t_expr_matnonawo6,
                       Y = recurrencewo6,
                       ncomp = 3)
```

```{r fig.width=5, fig.height=4}
plotIndiv(plsda_restwo6,
          nd.names = TRUE, 
          ellipse = TRUE, 
          legend = FALSE, 
          title = "sPLS-DA_Rec-Prim wo Pat6", 
          size.title = rel(1.5), 
          size.xlabel = rel(1.5), 
          size.ylabel = rel(1.5), 
          siz.axis = rel(1.2), 
          point.lwd = 0.8, 
          cex = 3)  
```

### GGplot-based visualization of Exploratory sPLS-DA {.unnumbered .unlisted}

```{r}
# preprocess pca results 
plsda_variates <- plsda_restwo6$variates$X %>% 
  as.data.frame() %>%
  rownames_to_column("Sample") %>%
  separate(col = "Sample",
           into = c("sample", 
                    "recurrence", 
                    "mixture"),
           sep = "\\_", 
           extra = "merge") %>%
  mutate(paired_id = paste(sample, 
                           recurrence, 
                           sep = "_")) %>%
  dplyr::select(-c(mixture, 
                   recurrence)) %>%
  left_join(. , sample_annotation2,
            by = "paired_id") 
```

```{r fig.width=5.5, fig.height=4.5, warning=FALSE}
ggplot(data = plsda_variates,
       aes(x = comp1, 
           y = comp2)) +
  geom_point(size = 4, aes(shape = mixture, 
                           color = recurrence)) +  
  geom_text(aes(label = patient), 
            position = position_nudge(x = 2)) +
  labs(title = "sPLS-DA plot of samples - pat6 excluded \nBased on the abundance values of proteins") +
  xlab(paste("sPLS-DA Comp-1", 
             round(plsda_restwo6$prop_expl_var$X[1]*100), 
             "% var explained")) + 
  ylab(paste("sPLS-DA Comp-2", 
             round(plsda_restwo6$prop_expl_var$X[2]*100), 
             "% var explained")) + 
  ggforce::geom_mark_ellipse(mapping = aes(color = recurrence)) +
  theme(axis.text.x = element_text(hjust = 0.5, 
                                   vjust = 0, 
                                   size = 10, 
                                   angle = 360),
        axis.text.y = element_text(hjust = 0.95, 
                                   vjust = 0.2, 
                                   size = 10),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", 
                                    fill=NA, 
                                    size=0.5),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")
```

The difference between initial and recurrent tumor is better resolved in the sPLS-DA plot when excluding patient 6.

## Sparcity reduction

Only proteins that were identified and quantified in 2 out of 3 TMT mixtures were kept for the inferential analysis. The proteins missing in 1 mixture were imputed with `missForest`.

```{r}
# sel_proteins_missing function is listed in the helper functions
proteins2exclude <- sel_proteins_missing(quant_annotated,
                                         threshold = 1)
```

Exclude proteins from initial abundance matrix.

```{r}
expr_matrix_filt <- expr_matrix2 %>%
  filter(!Protein %in% proteins2exclude)
```

### Missing values before and after filtering {.unnumbered .unlisted}

```{r}
nofiltnas_plot <- vis_miss(expr_matrix) + 
          ggtitle("Distribution of missing values",
                  subtitle = paste("All quant proteins. Total =", 
                                   nrow(expr_matrix))) + 
          theme(axis.text.x = element_text(hjust = 0.5, 
                                           vjust = 0.1, 
                                           size = 10, 
                                           angle = 90),
                axis.text.y = element_text(hjust = 0.5, 
                                           vjust = 0.1, 
                                           size = 10),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", 
                                            fill = NA, 
                                            size = 1.5),
                axis.title=element_text(size = 12,
                                        face = "bold"))

filtnas_plot <- vis_miss(expr_matrix_filt) + 
          ggtitle("Distribution of missing values",
                  subtitle = paste("Only present in 2/3 TMT mixtures. Total =",
                                   nrow(expr_matrix_filt))) + 
          theme(axis.text.x = element_text(hjust = 0.5, 
                                           vjust = 0.1, 
                                           size = 10, 
                                           angle = 90),
                axis.text.y = element_text(hjust = 0.5, 
                                           vjust = 0.1, 
                                           size = 10),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", 
                                            fill = NA, 
                                            size = 1.5),
                axis.title=element_text(size = 12,
                                        face = "bold"))
```

```{r fig.height=6, fig.width=10}
cowplot::plot_grid(nofiltnas_plot, 
                   filtnas_plot, 
                   ncol = 2)
```

## Missing value imputation (`missForest`)

```{r}
mat_filt <- expr_matrix_filt %>%
  column_to_rownames("Protein") %>%
  as.matrix()

t_mat_filt <- t(mat_filt)

if(!file.exists(here("rds/missforest_imp_mat_filt.Rds"))){

 t_imp_mat_filt <- missForest::missForest(t_mat_filt)
 t_imp_filt_mat <- t_imp_mat_filt$ximp

 mat_filt_imp <- t(t_imp_filt_mat)

 saveRDS(t_imp_mat_filt,
         file = here("rds/missforest_imp_mat_filt.Rds"))

} else {

 t_imp_mat_filt <- readRDS(here("rds/missforest_imp_mat_filt.Rds"))
 t_imp_filt_mat <- t_imp_mat_filt$ximp

 mat_filt_imp <- t(t_imp_filt_mat)

}
```

# Differential abundance analysis (DAA)

**Protein to Gene annotation**

```{r}
prot2gene <- identified_proteins %>%
  dplyr::select(Protein = protein_id, Gene = gene,
                Description = description)
```

**Prep protein abundance matrix**

```{r}
mat_filt_impwo6 <- as.data.frame(mat_filt_imp) %>%
  rownames_to_column("Protein") %>%
  dplyr::select(-starts_with("x6")) %>%
  column_to_rownames("Protein") %>%
  as.matrix()

sample_annotationwo6 <- sample_annotation2 %>%
  filter(patient != "x6")
```

## Set up design matrix

Initial vs recurrent samples are patient-matched, therefore the design matrix is setup so the linear model can account for the patient random effect.

```{r}
design_wo6 <- model.matrix(~ patientwo6 + recurrencewo6)

rownames(design_wo6) <- sample_annotationwo6$paired_id
```

Check design matrix:

```{r}
print(design_wo6[1:4,])
```

Check expression matrix:

```{r}
mat_filt_impwo6[1:4,1:6]
```

## Fit linear model (`limma`)

```{r}
source(here("scr/fit_limmawo6.R"))
```

```{r, warning=FALSE}
# fit limma
limma_tab_wo6 <- fit_limmawo6(mat_filt_impwo6, 
                         design_wo6, 
                         method = 'robust', 
                         Limma = "Robust - w Patient effect",
                         prot2gene = prot2gene)

#get proteins increased in recurrence
increased_in_rec <- limma_tab_wo6 %>%
  filter(logFC > 0,
         adj.P.Val < 0.05) %>%
  pull(Protein)

#get proteins decreased in recurrence
decreased_in_rec <- limma_tab_wo6 %>%
  filter(logFC < 0,
         adj.P.Val < 0.05) %>%
  pull(Protein)
```

## Volcano plot (DAA visualization)

```{r fig.width=6.5, fig.height=5.5, warning=FALSE}
our_volcano(limma_tab_wo6, 
            FC_cutoff = 0, 
            pval_cutoff = 0.05, 
            color_diffex = "red", 
            color_nondifex = "#2a9d8f", 
            interesting_proteins = NULL, 
            vert_line_col = "red",
            hline_col = "red", 
            hline_pos = 0.05, 
            linetype = "dashed",
            increased_in = "Recurrent", 
            comparison_title = "Recurrent vs Initial") +
  theme(axis.text.x = element_text(hjust = 0.5, 
                                   vjust = 0, 
                                   size = 10, 
                                   angle = 360),
        axis.text.y = element_text(hjust = 0.95, 
                                   vjust = 0.2, 
                                   size = 10),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", 
                                    fill = NA, 
                                    size = 0.5),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom") 
```

### Generate manuscript-ready volcano-plots {.unnumbered .unlisted}

```{r generate_volcano, fig.width=8, fig.height=6, echo=FALSE, warning=FALSE}
protlabels <- limma_tab_wo6 %>%
      filter(logFC > 0 | logFC < 0,
             adj.P.Val < 0.05)

signi_hits <- nrow(protlabels)

signi_up <- limma_tab_wo6 %>%
      filter(logFC > 0 ,
             adj.P.Val < 0.05) %>%
  nrow()

signi_down <- limma_tab_wo6 %>%
      filter(logFC < 0 ,
             adj.P.Val < 0.05) %>%
  nrow()

size <- 0.4

volcano <- ggplot(data = limma_tab_wo6,
                  mapping = aes(x = logFC, 
                                y = -log10(adj.P.Val))) +
      geom_point(data = limma_tab_wo6 %>% 
                 filter(logFC > 0,
                        adj.P.Val < 0.05),
                 mapping = aes(x = logFC, 
                               y = -log10(adj.P.Val)), 
                 color = "red",
                 size = size)+
      geom_point(data = limma_tab_wo6 %>% 
                   filter(logFC < -0,
                          adj.P.Val < 0.05),
                 mapping = aes(x = logFC, 
                               y = -log10(adj.P.Val)), 
                 color = "red",
                 size = size) +
      geom_point(data = limma_tab_wo6 %>% 
                   filter(logFC > 0,
                          adj.P.Val > 0.05),
                 mapping = aes(x = logFC, 
                               y = -log10(adj.P.Val)), 
                 color = "#2a9d8f",
                 size = size) +
      geom_point(data = limma_tab_wo6 %>% 
                   filter(logFC < -0,
                          adj.P.Val > 0.05),
                 mapping = aes(x = logFC, 
                               y = -log10(adj.P.Val)), 
                 color = "#2a9d8f",
                 size = size) +
      ggrepel::geom_text_repel(data = protlabels,
                               aes(label = Gene), 
                               size = 1.5) +
      geom_hline(yintercept = -log10(0.05),
                 color = "red", 
                 linetype = "dashed") +
      xlab("logFC - Recurrent / Primary")

```

```{r echo=FALSE}
ggsave(plot = volcano, 
       filename = here::here("figures/volcano_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 50)

ggsave(plot = volcano, 
       filename = here::here("figures/volcano_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 63,
       height = 50)
```

```{r fig.width=8, fig.height=6, echo=FALSE}
protlabels <- limma_tab_wo6 %>%
      filter(logFC > 0 | logFC < 0,
             adj.P.Val < 0.05,
             Gene %in% c("ASAH1", "GPNMB", "SYNM", "CD14"))

size <- 0.4

volcano2 <- ggplot(data = limma_tab_wo6,
                      mapping = aes(x = logFC, y = -log10(adj.P.Val))) +
      geom_point(data = limma_tab_wo6 %>% filter(logFC > 0,
                                           adj.P.Val < 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), 
                 color = "red",
                 size = size)+
      geom_point(data = limma_tab_wo6 %>% filter(logFC < -0,
                                           adj.P.Val < 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "red",
                 size = size) +
      geom_point(data = limma_tab_wo6 %>% filter(logFC > 0,
                                           adj.P.Val > 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "#2a9d8f",
                 size = size) +
      geom_point(data = limma_tab_wo6 %>% filter(logFC < -0,
                                           adj.P.Val > 0.05),
                 mapping = aes(x = logFC, y = -log10(adj.P.Val)), color = "#2a9d8f",
                 size = size) +
      ggrepel::geom_text_repel(data = protlabels,
                               aes(label = Gene), size = 3.5) +
      geom_hline(yintercept = -log10(0.05),
                 color = "red", linetype = "dashed") +
      xlab("logFC - Recurrent / Primary")

```

```{r echo=FALSE}
ggsave(plot = volcano2, 
       filename = here::here("figures/volcano2_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 50)

ggsave(plot = volcano2, 
       filename = here::here("figures/volcano2_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 63,
       height = 50)
```

### Write tabular results (for supplementary tables) {.unnumbered .unlisted}

```{r}
write_tsv(x = limma_tab_wo6, 
          file = here("suppl_tables/diff_abundance_gbm_rec_vs_init_proteomics.tsv"))
```

## Boxplots specific proteins

```{r}
quant_anwgene <- left_join(quant_annotated, 
                           prot2gene)

tab_2_paired_boxplot <-  quant_anwgene %>%
  filter(Gene %in% c("ASAH1", 
                     "SYNM", 
                     "GPNMB",
                     "MPO"))
# prep data
slim_data <- pivot_longer(expr_matrix_filt, 
                          cols = 2:length(names(expr_matrix_filt)), 
                          names_to = c("patient",
                                       "condition"),
                          values_to = "Intensity",
                          names_sep = "\\_") %>% 
  group_by(Protein)

slim_data2 <- pivot_wider(slim_data, 
                          values_from = Intensity, 
                          names_from = condition) %>%
  filter(patient != "x6")

slim_data3 <- slim_data2 %>% 
  mutate(Abs_diff = rec-prim) %>%
  mutate(rank = row_number(Abs_diff)) %>% 
  dplyr::rename(Initial = prim, 
                Recurrent = rec) %>%
  ungroup() %>%
  filter(Protein %in% c("Q13510", 
                        "Q14956", 
                        "O15061",
                        "P05164")) %>%
  mutate(Gene = case_when(Protein == "Q13510" ~ "ASAH1",
                          Protein == "Q14956" ~ "GPNMB",
                          Protein == "O15061" ~ "SYNM",
                          Protein == "P05164" ~ "MPO"))
```

```{r fig.width=7.5, fig.height=5.9}
ggpaired(slim_data3,
         cond1 = "Initial",
         cond2 = "Recurrent", 
         fill = "condition",
         ylab = "Normalized Abundance",
         label = NULL,
         repel = TRUE,
         facet.by = "Gene", 
         point.size = 0.2, 
         line.size = 0.1) 
```

```{r echo=FALSE}
size <- 5

paired_boxplots <- ggpaired(slim_data3,
                            cond1 = "Initial",
                            cond2 = "Recurrent", 
                            fill = "condition",
                            ylab = "Normalized Abundance",
                            label = NULL,
                            repel = TRUE,
                            facet.by = "Gene", 
                            point.size = 0.2, 
                            line.size = 0.1) +
      theme(axis.text.x = element_text(hjust = 0.5, 
                                       vjust = 0, 
                                       size = 5, 
                                       angle = 360),
           axis.text.y = element_text(hjust = 0.95, 
                                      vjust = 0.2, 
                                      size = 5),
           panel.background = element_blank(),
           panel.grid.major = element_blank(),
           panel.border = element_rect(colour = "black", 
                                       fill = NA, 
                                       size = 0.5),
           axis.title = element_text(size = size),
           legend.position="none",
           strip.text = element_text(hjust = 0.5, 
                                     vjust = 0, 
                                     size = 4, 
                                     angle = 360,
                                     margin = margin(0.4,0,0.4,0, "mm"),
                                     lineheight = 0.3)) 
```

```{r echo=FALSE}
ggsave(plot = paired_boxplots, 
       filename = here::here("figures/paired_box_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 61,
       height = 45)

ggsave(plot = paired_boxplots, 
       filename = here::here("figures/paired_box_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 61,
       height = 45)
```

#### ASAH1 vs MPO abundance (Rec vs Init)

```{r}
tab_2_mporatio <- tab_2_paired_boxplot %>%
  filter(Gene %in% c("ASAH1", "MPO")) %>%
  dplyr::select(Gene,
                recurrence,
                Abundance,
                patient,
                sample_id) %>%
  mutate(Condition = case_when(recurrence == "prim" ~ "Initial", 
                               recurrence == "rec" ~ "Recurrent")) %>%
  dplyr::select(-recurrence)

mpo_ratio_prep <- tab_2_mporatio %>% 
  pivot_wider(values_from = Abundance,
              names_from = Gene) %>%
  filter(patient != "x6") %>%
  mutate(Ratio_MPO_vs_ASAH1 = ASAH1 - MPO) %>%
  dplyr::select(-c(MPO, ASAH1)) %>%
  pivot_wider(values_from = Ratio_MPO_vs_ASAH1,
              names_from = Condition) %>% 
  mutate(measure = "Delta(ASAH1 - MPO)")

mpo_ratio_prep2 <- tab_2_mporatio %>% 
  pivot_wider(values_from = Abundance,
              names_from = Gene) %>%
  filter(patient != "x6") 
```

```{r}
paired_boxplots_asah1mporatio <- ggpaired(mpo_ratio_prep,
                                          cond1 = "Initial",
                                          cond2 = "Recurrent", 
                                          fill = "condition",
                                          ylab = "Abundance Delta(ASAH1 - MPO)",
                                          label = NULL,
                                          repel = TRUE,
                                          point.size = 0.2, 
                                          line.size = 0.1) +
      stat_compare_means(paired = TRUE, 
                         vjust = 2, 
                         size = 1.5) +
      theme(axis.text.x = element_text(hjust = 0.5, 
                                       vjust = 0, 
                                       size = 5, 
                                       angle = 360),
           axis.text.y = element_text(hjust = 0.95, 
                                      vjust = 0.2, 
                                      size = 5),
           panel.background = element_blank(),
           panel.grid.major = element_blank(),
           panel.border = element_rect(colour = "black", 
                                       fill = NA, 
                                       size = 0.5),
           axis.title = element_text(size = size),
           legend.position="none",
           strip.text = element_text(hjust = 0.5, 
                                     vjust = 0, 
                                     size = 4, 
                                     angle = 360,
                                     margin = margin(0.4,
                                                     0,
                                                     0.4,
                                                     0, "mm"),
                                     lineheight = 0.3))  
```

##### Compare Deltas of abundance (Boxplot)

```{r fig.width=4.8, fig.height=3.6}
print(paired_boxplots_asah1mporatio) 
```

```{r echo=FALSE}
ggsave(plot = paired_boxplots_asah1mporatio, 
       filename = here::here("figures/paired_box_delta_asah1_mpo_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 60,
       height = 45)

ggsave(plot = paired_boxplots_asah1mporatio, 
       filename = here::here("figures/paired_box_delta_asah1_mpo_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 60,
       height = 45)
```

##### Scatter plot ASAH1 vs MPO

```{r}
scatter_asah1mpo <- ggplot(data = mpo_ratio_prep2,
                           aes(x = ASAH1,
                               y = MPO)) + 
                    geom_point(mapping = aes(color = Condition)) + 
                    geom_smooth(method = lm, 
                                se = FALSE, 
                                color = "black",
                                size = 0.5) 
```

```{r fig.width=4.8, fig.height=3.6}
print(scatter_asah1mpo)
```

```{r echo=FALSE}
ggsave(plot = scatter_asah1mpo, 
       filename = here::here("figures/scatterplot_asah1_mpo_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 60,
       height = 60)

ggsave(plot = scatter_asah1mpo, 
       filename = here::here("figures/scatterplot_asah1_mpo_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 60,
       height = 60)
```

## Differences in abundance of ASAH1 by type of resection

Load the patient clinical data

```{r echo=FALSE}
patient_table <- readxl::read_xlsx(path = here("data/Patients_information_FINAL.xlsx"))
```

```{r}
patients_proteomics <- patient_table %>%
  filter(`Mass Spectrometry` == "X") %>%
  dplyr::select(Sex, 
                `Age at surgery (years)`, 
                `Time-to-reccurence (days)`,
                `Type of resection`,
                `Tumor localization`, 
                Treatment,
                sample_id = `LAB-ID`) %>%
  mutate(`Time-to-reccurence (days)` = as.numeric(`Time-to-reccurence (days)`))
```

```{r}
sample2patient_id <- mpo_ratio_prep2 %>% 
  dplyr::select(sample_id, 
                patient) %>% 
  distinct()

patients_proteomics <- left_join(patients_proteomics, 
                                 sample2patient_id) %>%
  dplyr::select(-matches("sample"))
```

```{r}
resection_diff_asah1 <- left_join(mpo_ratio_prep2,
                                  patients_proteomics) %>% 
  dplyr::select(-MPO) %>%
  dplyr::rename(abundance_ASAH1 = ASAH1,
                type_resection = `Type of resection`) %>%
  na.exclude() %>%
  distinct()
```

```{r}
resection_diff_asah1plot <- ggplot(data = resection_diff_asah1, 
                                   mapping = aes(x = Condition,
                                                 y = abundance_ASAH1,
                                                 fill = type_resection)) + 
  geom_boxplot()
```

```{r}
resection_diff_asah1plot
```

```{r}
twowayanova_resection <- aov(abundance_ASAH1 ~ Condition * type_resection, 
                             data = resection_diff_asah1)
```

```{r}
summary(twowayanova_resection)
```

```{r}
TukeyHSD(twowayanova_resection)
```

There is no significant difference in ASAH1 abundance based on type of resection.

# ID coverage after sparcity reduction

```{r}
ids_summary2 <- quant_annotated %>% 
  dplyr::select(Protein, 
                mixture, 
                Abundance) %>% 
  distinct(Protein, 
           mixture, 
           .keep_all = TRUE) %>%
  group_by(mixture) %>%
  summarise(`Nr of IDs` = sum(!is.na(Abundance))) %>% 
  bind_rows(.,
            tibble(mixture = "Quant_in_all",
                   `Nr of IDs` = nrow(expr_matnona))) %>%
  bind_rows(.,
            tibble(mixture = "Quant_2/3_mixtures",
                   `Nr of IDs` = nrow(mat_filt_impwo6))) %>%
  bind_rows(.,
            tibble(mixture = "Overall",
                   `Nr of IDs` = nrow(expr_matrix)))

ids_summary2 <- mutate(ids_summary2,
                       Mixture = case_when(mixture == "tmt_1" ~ "Mix_1",
                                           mixture == "tmt_2" ~ "Mix_2",
                                           mixture == "tmt_3" ~ "Mix_3",
                                           mixture == "Quant_in_all" ~ "Quant. in all",
                                           mixture == "Overall" ~ "Overall",
                                           mixture == "Quant_2/3_mixtures" ~ "Quant. in 2/3")) %>%
  mutate(Mixture = factor(Mixture, levels = c("Overall","Quant. in 2/3", "Quant. in all",
                                               "Mix_3", "Mix_2", "Mix_1")))
```

Number of identified proteins: `r nrow(expr_matrix)`

Number of identified and quantified proteins in all samples: `r nrow(expr_matnona)`

Number of identified and quantified proteins 2 out of 3 batches: `r nrow(mat_filt_impwo6)`

```{r fig.height=4, fig.width=7, warning=FALSE}
ggplot(data = ids_summary2,
       aes(y = `Nr of IDs`, 
           x = mixture)) +
  geom_col(fill = "#008683") +

  geom_text(size = 5, 
            position = position_stack(vjust = 0.5), 
            aes(label = `Nr of IDs`)) +
  coord_cartesian(ylim = c(0, 5000))+
  coord_flip() +
  labs(title = "Number of Identified and quantified Proteins by Mixture") +
  ylab("Nr of IDed and Quant Proteins") + 
  xlab("TMT Mixture") + 
  theme(axis.text.x = element_text(hjust = 0.5, 
                                   vjust = 0.1, 
                                   size = 14, 
                                   angle = 0),
        axis.text.y = element_text(hjust = 0.5, 
                                   vjust = 0.1, 
                                   size = 12, 
                                   angle = 0),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey"),
        panel.border = element_rect(colour = "black", 
                                    fill = NA, 
                                    size = 1.5),
        axis.title = element_text(size = 12, 
                                  face = "bold"))
```

```{r echo=FALSE}
barplot_cover <- ggplot(data = ids_summary2,
       aes(y = `Nr of IDs`, 
           x = Mixture)) +
  geom_col(fill = "#008683") + 
  geom_text(size = 2, 
            position = position_stack(vjust = 0.5), 
            aes(label = `Nr of IDs`)) +
  coord_cartesian(ylim = c(0, 5000))+
  coord_flip() +
  theme(axis.text.x = element_text(hjust = 0.5, 
                                   vjust = 0, 
                                   size = 7, 
                                   angle = 360),
        axis.text.y = element_text(hjust = 0.95, 
                                   vjust = 0.2, 
                                   size = 7),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", 
                                    fill = NA, 
                                    size = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")
```

```{r echo=FALSE}
ggsave(plot = barplot_cover, 
       filename = here::here("figures/barplot_coverage_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 50)

ggsave(plot = barplot_cover, 
       filename = here::here("figures/barplot_coverage_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 63,
       height = 50)
```

# Exploratory analysis after sparcity reduction

## Plots of normalized abundance distribution

```{r}
expr_matrix_filtwo6 <- as.data.frame(mat_filt_imp) %>%
  rownames_to_column("Protein") %>%
  dplyr::select(-starts_with("x6"))

long_mat_f <- pivot_longer(expr_matrix_filtwo6,
                         cols = starts_with("x"),
                         names_sep = "_",
                         names_to = c("patient", "recurrence"),
                         values_to = "Abundance")

long_mat_3 <- long_mat_f %>%
  mutate(paired_id = paste(patient, recurrence, sep = "_"))

paired_annot_uniq2 <- long_mat_3 %>%
  dplyr::select(patient, recurrence, paired_id) %>%
  distinct()

paired_annot_uniq2$paired_id
## Annotated abundance data in long format ----

quant_annotated_f <- left_join(long_mat_3,
                             sample_annotation2, 
                             by = c("paired_id", "patient","recurrence")) %>%
  mutate(Channel_mix = paste(mixture,channel,  sep = "_"),
         paired_id = factor(paired_id, levels = paired_annot_uniq2$paired_id)) %>%
  mutate(Sample = str_replace_all(paired_id, pattern = "prim", replacement = "init"),
         Stage = if_else(recurrence == "prim",
                         true = "Initial",
                         false = "Recurrent"))

```

```{r}
expr_matrix_filtwo6 <- as.data.frame(mat_filt_imp) %>%
  rownames_to_column("Protein") %>%
  dplyr::select(-starts_with("x6"))

long_mat_f <- pivot_longer(expr_matrix_filtwo6,
                         cols = starts_with("x"),
                         names_sep = "_",
                         names_to = c("patient", "recurrence"),
                         values_to = "Abundance")

long_mat_3 <- long_mat_f %>%
  mutate(paired_id = paste(patient, recurrence, sep = "_"))

paired_annot_uniq2 <- long_mat_3 %>%
  dplyr::select(patient, recurrence, paired_id) %>%
  distinct()

paired_annot_uniq2$paired_id
## Annotated abundance data in long format ----

quant_annotated_f <- left_join(long_mat_3,
                             sample_annotation2, 
                             by = c("paired_id", "patient","recurrence")) %>%
  mutate(Channel_mix = paste(mixture,channel,  sep = "_"),
         paired_id = factor(paired_id, levels = paired_annot_uniq2$paired_id)) %>%
  mutate(Sample = str_replace_all(paired_id, pattern = "prim", replacement = "init"),
         Stage = if_else(recurrence == "prim",
                         true = "Initial",
                         false = "Recurrent"))

```

### Normalized ion reporter abundance distribution by TMT channel/sample {.unnumbered}

```{r warning=FALSE}
qcplot_f <- ggplot(quant_annotated_f,
                   mapping = aes(x = Sample, 
                                 y = Abundance, 
                                 fill = Stage)) +
                   geom_boxplot(outlier.size=0.1, size = 0.25) + 
                   theme(axis.text.x = element_text(hjust = 0.5, 
                                                    vjust = 0, 
                                                    size = 7, 
                                                    angle = 90),
                         axis.text.y = element_text(hjust = 0.5, 
                                                    vjust = 0, 
                                                    size = 6),
                         panel.background = element_blank(),
                         panel.grid.major = element_blank(),
                         panel.border = element_rect(colour = "black", 
                                                     fill = NA, 
                                                     size = 0.5),
                         axis.title=element_text(size = 8),
                         axis.title.x = element_blank(),
                         legend.text = element_text(size = 7),
                         legend.title = element_text(size = 8),
                         legend.key.height= unit(3, 'mm'),
                         legend.key.width= unit(3, 'mm'),
                         legend.position="bottom")
```

```{r fig.height=4, fig.width=7, warning=FALSE, echo=FALSE}
print(qcplot_f)
```

```{r echo=FALSE}
ggsave(plot = qcplot_f, 
       filename = here::here("figures/boxplot_abund_dist_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 50)

ggsave(plot = qcplot_f, 
       filename = here::here("figures/boxplot_abund_dist_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 63,
       height = 50)
```

# Over-representation analyses/tests (ORAs)

## Prep tabular info for ORAs {.unnumbered .unlisted}

```{r}
tab_sig_prots_tryptic <- bind_rows(tibble(protein = increased_in_rec,
                                  characteristic = "up-regulated"),
                           tibble(protein = decreased_in_rec,
                                  characteristic = "down-regulated"))

limma_tab_wo62_tryptic <- limma_tab_wo6

unip2symb_tryptic <- bitr(limma_tab_wo62_tryptic$Protein, 
                  fromType = "UNIPROT", 
                  toType = c("SYMBOL", "ENTREZID"),
                  OrgDb = org.Hs.eg.db) %>%
  dplyr::rename(protein = UNIPROT) %>%
  mutate(Protein = protein)

tab_sig_prots_tryptic <- left_join(tab_sig_prots_tryptic, unip2symb_tryptic)

unip2symbIDed_tryptic <- bitr(identified_proteins$protein_id, 
                  fromType = "UNIPROT", 
                  toType = c("SYMBOL", "ENTREZID"),
                  OrgDb = org.Hs.eg.db)

limma_tab_wo62_tryptic <- left_join(limma_tab_wo62_tryptic, unip2symb_tryptic)

geneList <- limma_tab_wo62_tryptic$logFC
names(geneList) <- limma_tab_wo62_tryptic$Protein
geneList <-  sort(geneList, decreasing = TRUE)

geneListr <- limma_tab_wo62_tryptic$logFC
names(geneListr) <- limma_tab_wo62_tryptic$ENTREZID
geneListr <-  sort(geneListr, decreasing = TRUE)
```

## Reactome

```{r}
group_comparison_react <- compareCluster(ENTREZID~characteristic, 
                                              data=tab_sig_prots_tryptic, 
                                              fun="enrichPathway",
                                              organism = "human",
                                              pvalueCutoff = 0.05,
                                              pAdjustMethod = "BH",
                                              qvalueCutoff = 0.2,
                                              universe = unip2symbIDed_tryptic$ENTREZID,
                                              minGSSize = 10,
                                              maxGSSize = 1000,
                                              readable = TRUE)
```

### Dotplot Reactome

```{r warning=FALSE, message=FALSE}
dotplot_react <- enrichplot::dotplot(group_comparison_react, 
                                     x = "characteristic") + 
          xlab("Quant in Recurrent status") +
  scale_color_continuous(low = "red", 
                         high = "blue",
            guide=guide_colorbar(reverse = TRUE),
            breaks = c(0.01, 
                       0.04)) +
  scale_size_continuous(breaks = c(0.10, 
                                   0.25)) +
  theme(axis.text.x = element_text(hjust = 0.4, 
                                   vjust = 0.1, 
                                   size = 6, 
                                   angle = 340),
        axis.text.y = element_text(hjust = 0.95, 
                                   vjust = 0.2, 
                                   size = 6),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "black", 
                                    fill = NA, 
                                    size = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 6),
        legend.key.height= unit(2, 'mm'),
        legend.key.width= unit(2, 'mm'),
        legend.position="bottom")
```

```{r}
print(dotplot_react)
```

```{r}
ggsave(plot = dotplot_react, 
       filename = here::here("figures/dotplot_reactome_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 91,
       height = 90)

ggsave(plot = dotplot_react, 
       filename = here::here("figures/dotplot_reactome_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 91,
       height = 90)
```

### Cnet plot Reactome

```{r}
cnet_plot <- enrichplot::cnetplot(group_comparison_react) + 
  theme(legend.text = element_text(size = 7),
        text = element_text(size = 4),
                legend.title = element_text(size = 8),
                legend.key.height = unit(3, 'mm'),
                legend.key.width = unit(3, 'mm'),
                legend.position = "bottom")
```

```{r}
print(cnet_plot)
```

```{r}
ggsave(plot = cnet_plot, 
       filename = here::here("figures/cnet_plot_reactome_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 110,
       height = 110)

ggsave(plot = cnet_plot, 
       filename = here::here("figures/cnet_plot_reactome_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 110,
       height = 110)

ggsave(plot = cnet_plot, 
       filename = here::here("figures/cnet_plot_reactome_proteomics_gbm.svg"), 
       device = "svg",
       units = "mm",
       width = 110,
       height = 110)
```

# Protein coverage plots

```{r}
source(here("scr/protein_peptide_coverage_plot.R"))
source(here("scr/annotate_peptides.R"))
```

## Annotate peptides from interesting proteins {.unnumbered .unlist}

```{r}
# get protein to peptide pairs
pept2prot_inter <- identified_peptides %>%
  filter(gene %in% c("ASAH1", 
                     "SYNM", 
                     "GPNMB")) %>%
  dplyr::select(Peptide = sequence, 
                Genes = protein_id)

# map peptides to protein sequence
annotated_peptides <- annotate_peptides(expr_mat = pept2prot_inter,
                                        fasta = fasta_ident_prots)

annotated_peptides2 <- annotated_peptides %>%
  mutate(peptide_length = str_length(Peptide))
```

## ASAH1

```{r}
feat_asah1 <- drawProteins::get_features("Q13510")

feat_asah1df <- drawProteins::feature_to_dataframe(feat_asah1)
#Get peptides as coverage features for plotting

cov_feat_asah1 <- get_coverage(annotated_peptides = annotated_peptides2, 
                               id = "Q13510")
#Merge coverage feature info with protein sequence features info:

fullfeat_asah1 <- add_peptides(feat_asah1df,
                               peptide_coverage_data = cov_feat_asah1)
# drawCanvas
canvas <- draw_canvas(fullfeat_asah1)

# draw the protein chain
wchain <- draw_chains(canvas, fullfeat_asah1)

# draw protein domains
wdomainsasah1 <- draw_domains(wchain, data = fullfeat_asah1)
#pasa <- drawProteins::draw_folding(wdomainsasah1, fullfeat_asah1)
cov_feat_asah1$order <- 1.4

pepcov_asah1 <- draw_peptides(wdomainsasah1, cov_feat_asah1)
```

```{r}
print(pepcov_asah1)
```

```{r}
ggsave(plot = pepcov_asah1, 
       filename = here::here("figures/asah1_coverage_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 150,
       height = 100)

ggsave(plot = pepcov_asah1, 
       filename = here::here("figures/asah1_coverage_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 150,
       height = 100)
```

## SYMN

```{r}
feat_SYNM <- drawProteins::get_features("O15061")
feat_SYNMdf <- drawProteins::feature_to_dataframe(feat_SYNM)
#Get peptides as coverage features for plotting

cov_feat_SYNM <- get_coverage(annotated_peptides = annotated_peptides2, 
                               id = "O15061")
#Merge coverage feature info with protein sequence features info:

fullfeat_SYNM <- add_peptides(feat_SYNMdf,
                               peptide_coverage_data = cov_feat_SYNM)
# drawCanvas
canvasSYNM <- draw_canvas(fullfeat_SYNM)

# draw the protein chain
wchainSYNM <- draw_chains(canvasSYNM, fullfeat_SYNM)
cov_feat_SYNM$order <- 1.4

pepcov_SYNM <- draw_peptides(wchainSYNM, cov_feat_SYNM)
```

```{r}
print(pepcov_SYNM)
```

```{r}
ggsave(plot = pepcov_SYNM, 
       filename = here::here("figures/SYNM_coverage_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 150,
       height = 100)

ggsave(plot = pepcov_SYNM, 
       filename = here::here("figures/SYNM_coverage_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 150,
       height = 100)
```

# Associate protein expression with Time to Recurrence

```{r}
library(readxl)
```

```{r echo=FALSE}
latency_data <- read_excel(path = here("data/ELISA_LATENCY_KM.xlsx"), 
                           skip = 1
                           ) %>%
  clean_names()
```

```{r echo=FALSE}
patient_table <- readxl::read_xlsx(path = here("data/Patients_information_FINAL.xlsx"))
```

```{r echo=FALSE}
latency_data %>%
  dplyr::mutate(ASAH1.ng.ml = log2(asah1_ng_ml_6+1),
                SYNM.ng.ml = log2(synm_ng_ml_7+1),
                GPNMB.ng.ml = log2(gpnmb_ng_ml_8+1),
                MMP.9.ng.ml = log2(mmp_9_ng_ml_9+1)) %>%
  dplyr::mutate(ASAH1 = ifelse(ASAH1.ng.ml>=mean(asah1_ng_ml_6), "high", "low"),
                SYNM = ifelse(SYNM.ng.ml>=mean(synm_ng_ml_7), "high", "low"),
                GPNMB = ifelse(GPNMB.ng.ml>=mean(gpnmb_ng_ml_8), "high", "low"),
                MMP9 = ifelse(MMP.9.ng.ml>=mean(mmp_9_ng_ml_9, na.rm=T), "high", "low")) -> latency_data
```

```{r}
patients_proteomics <- patient_table %>%
  filter(`Mass Spectrometry` == "X") %>%
  dplyr::select(Sex, 
                `Age at surgery (years)`, 
                `Time-to-reccurence (days)`,
                `Type of resection`,
                `Tumor localization`, 
                Treatment,
                sample_id = `LAB-ID`) %>%
  mutate(`Time-to-reccurence (days)` = as.numeric(`Time-to-reccurence (days)`))
```

## What's the association of ASAH1/SYNM/GPNNMB/MMP9 at Initial stage with time to recurrence?

```{r}
quant_annot_intr_init <- quant_annotated %>%
  filter(Protein %in% c("Q13510", 
                        "Q14956", 
                        "O15061",
                        "P14780"),
         recurrence == "prim") %>%
  mutate(Gene = case_when(Protein == "Q13510" ~ "ASAH1a",
                          Protein == "Q14956" ~ "GPNMBa",
                          Protein == "O15061" ~ "SYNMa",
                          Protein == "P14780" ~ "MMP9a")) %>% 
  left_join(patients_proteomics) %>% 
  dplyr::select(sample_id,
                patient,
                Gene,
                Abundance, 
                `Time-to-reccurence (days)`) %>%
  pivot_wider(names_from = Gene, 
              values_from = Abundance) %>%
  dplyr::mutate(ASAH1 = ifelse(ASAH1a >= mean(ASAH1a), 
                               "high", 
                               "low"),
                SYNM = ifelse(SYNMa >= mean(SYNMa), 
                              "high", 
                              "low"),
                GPNMB = ifelse(GPNMBa >= mean(GPNMBa), 
                               "high", 
                               "low"),
                MMP9 = ifelse(MMP9a >= mean(MMP9a, 
                                            na.rm=T), 
                              "high", 
                              "low"))

labid2patientn <- quant_annot_intr_init %>% 
  dplyr::select(patient, sample_id) %>% 
  dplyr::distinct()

patients_proteomics <- left_join(labid2patientn, 
                                 patients_proteomics)

```

```{r}
library(survival)
library(survminer)
```

```{r}
fit1 <- survfit(Surv(`Time-to-reccurence (days)`) ~ ASAH1, 
               data = quant_annot_intr_init)
print(fit1)
# Change color, linetype by strata, risk.table color by strata
surv_p1 <- ggsurvplot(fit1,
          pval = TRUE, conf.int = TRUE,
          risk.table = FALSE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          ) +
  labs(x="Time to Recurrence (Days)", y = "Time to Recurrence Probability")

fit2 <- survfit(Surv(`Time-to-reccurence (days)`) ~ SYNM, 
               data = quant_annot_intr_init)
print(fit2)
# Change color, linetype by strata, risk.table color by strata
surv_p2 <- ggsurvplot(fit2,
          pval = TRUE, conf.int = TRUE,
          risk.table = FALSE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          ) +
  labs(x="Time to Recurrence (Days)", y = "Time to Recurrence Probability")

fit3 <- survfit(Surv(`Time-to-reccurence (days)`) ~ GPNMB, 
               data = quant_annot_intr_init)
print(fit3)
# Change color, linetype by strata, risk.table color by strata
surv_p3 <- ggsurvplot(fit3,
          pval = TRUE, 
          conf.int = TRUE,
          risk.table = FALSE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          ) +
  labs(x="Time to Recurrence (Days)", y = "Time to Recurrence Probability")

fit4 <- survfit(Surv(`Time-to-reccurence (days)`) ~ MMP9, 
               data = quant_annot_intr_init)
print(fit4)
# Change color, linetype by strata, risk.table color by strata
surv_p4 <- ggsurvplot(fit4,
          pval = TRUE, conf.int = TRUE,
          risk.table = FALSE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          ) +
  labs(x="Time to Recurrence (Days)", y = "Time to Recurrence Probability")
```

```{r}
surv_p1
surv_p2
surv_p3
surv_p4
```

## What's the association of ASAH1/SYNM/GPNNMB/MMP9 at Recurrent stage with time to recurrence?

```{r}
quant_annot_intr_rec <- quant_annotated %>%
  filter(Protein %in% c("Q13510", 
                        "Q14956", 
                        "O15061",
                        "P05164"),
         recurrence == "rec") %>%
  mutate(Gene = case_when(Protein == "Q13510" ~ "ASAH1a",
                          Protein == "Q14956" ~ "GPNMBa",
                          Protein == "O15061" ~ "SYNMa",
                          Protein == "P05164" ~ "MMP9a"))  %>% 
  left_join(patients_proteomics, by = "patient") %>% 
  dplyr::select(patient,
                Gene,
                Abundance, 
                `Time-to-reccurence (days)`) %>%
  pivot_wider(names_from = Gene, 
              values_from = Abundance) %>%
  dplyr::mutate(ASAH1 = ifelse(ASAH1a >= mean(ASAH1a), 
                               "high", 
                               "low"),
                SYNM = ifelse(SYNMa >= mean(SYNMa), 
                              "high", 
                              "low"),
                GPNMB = ifelse(GPNMBa >= mean(GPNMBa), 
                               "high", 
                               "low"),
                MMP9 = ifelse(MMP9a >= mean(MMP9a, 
                                            na.rm=T), 
                              "high", 
                              "low")) 
```

```{r}
fit12 <- survfit(Surv(`Time-to-reccurence (days)`) ~ ASAH1, 
               data = quant_annot_intr_rec)
print(fit12)
# Change color, linetype by strata, risk.table color by strata
surv_p12 <- ggsurvplot(fit12,
          pval = TRUE, conf.int = TRUE,
          risk.table = FALSE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          ) +
  labs(x="Time to Recurrence (Days)", y = "Time to Recurrence Probability")

fit22 <- survfit(Surv(`Time-to-reccurence (days)`) ~ SYNM, 
               data = quant_annot_intr_rec)
print(fit22)
# Change color, linetype by strata, risk.table color by strata
surv_p22 <- ggsurvplot(fit22,
          pval = TRUE, conf.int = TRUE,
          risk.table = FALSE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          ) +
  labs(x="Time to Recurrence (Days)", y = "Time to Recurrence Probability")

fit32 <- survfit(Surv(`Time-to-reccurence (days)`) ~ GPNMB, 
               data = quant_annot_intr_rec)
print(fit32)
# Change color, linetype by strata, risk.table color by strata
surv_p32 <- ggsurvplot(fit32,
          pval = TRUE, 
          conf.int = TRUE,
          risk.table = FALSE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          ) +
  labs(x="Time to Recurrence (Days)", y = "Time to Recurrence Probability")

fit42 <- survfit(Surv(`Time-to-reccurence (days)`) ~ MMP9, 
               data = quant_annot_intr_rec)
print(fit42)
# Change color, linetype by strata, risk.table color by strata
surv_p42 <- ggsurvplot(fit42,
          pval = TRUE, conf.int = TRUE,
          risk.table = FALSE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          ) +
  labs(x="Time to Recurrence (Days)", y = "Time to Recurrence Probability")
```

```{r}
surv_p12
surv_p22
surv_p32
surv_p42
```