### 2023-02-16 reports/gbm_recurrence_proteogenomics_manuscript_report.qmd
proteogenomics_clean_up

---
title: "GBM Recurrence Proteogenomics"
author: "Miguel Cosenza-Contreras"
format:
  gfm:
    toc: true 
    toc-depth: 4 # default is 3
    number-sections: true
    execute:
      output: true
always_allow_html: true
editor: source
---

```{r}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

source(here::here("scr/helper_functions.R"))

## Required packages ----
library(tidyverse)
library(mixOmics)
library(fs)
library(kableExtra)
library(sva)
library(limma)
library(naniar)
library(missForest)
library(clusterProfiler)
library(ReactomePA)
library(org.Hs.eg.db)
library(DT)
library(here)
library(janitor)
library(drawProteins)
library(seqinr)
library(ggpubr)
library(ggrepel)
library(extrafont)
extrafont::loadfonts(device = "win")

theme_set(theme(axis.text.x = element_text(hjust = 0.5, 
                                           vjust = 0, 
                                           size = 6, 
                                           angle = 360),
                axis.text.y = element_text(hjust = 0.5, 
                                           vjust = 0, 
                                           size = 6),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", 
                                            fill = NA, 
                                            size = 1.5),
                axis.title=element_text(size = 8),
                legend.title = element_text(size = 8),
                legend.key.height = unit(3, 
                                         'mm'),
                legend.key.width = unit(3, 
                                        'mm'),
                legend.position = "bottom",
                text = element_text(family = "Helvetica")))

```

# Initial data loading and wrangling

Load id/quant data and prep coverage information

```{r message=FALSE}
## tmt-integrator output loading 
tmt_protdata_frag_progen <- read_tsv(here("data/proteogenomics/sras_cellreports_2021_paper_razor/tmt-report/abundance_protein_MD.tsv")) %>% 
  janitor::clean_names()

tmt_peptdata_frag_progen <- read_tsv(here("data/proteogenomics/sras_cellreports_2021_paper_razor/tmt-report/abundance_peptide_MD.tsv")) %>% 
  janitor::clean_names()

identified_proteins_progen <- read_tsv(here("data/proteogenomics/sras_cellreports_2021_paper_razor/combined_protein.tsv")) %>% 
  janitor::clean_names()

identified_peptides_progen <- read_tsv(here("data/proteogenomics/sras_cellreports_2021_paper_razor/combined_peptide.tsv")) %>% 
  janitor::clean_names()

index2genepeptide <- tmt_peptdata_frag_progen %>%
  dplyr::select(index, gene, peptide, protein_id)
```

```{r}
sample_annotation <- read_csv(here("data/sample_annotation.csv"))

# correct annotation

sample_annotation2 <- sample_annotation %>%
  mutate(patient = paste("x",
                         patient, 
                         sep = ""),
         recurrence = case_when(recurrence == "initial" ~ "prim",
                                recurrence == "recurrent" ~ "rec",
                                TRUE ~ recurrence)) %>%
  mutate(paired_id = paste(patient, 
                           recurrence, 
                           sep = "_")) %>%
  filter(recurrence %in% c("prim", 
                           "rec"))

sample_annotationwo6 <- sample_annotation2 %>%
  filter(patient != "x6")
```

```{r}
# wide expression matrix with NAs
expr_matrix_progen <- dplyr::select(tmt_protdata_frag_progen,
                             Protein = index, 
                             starts_with("x"))

long_mat_progen <- pivot_longer(expr_matrix_progen,
                         cols = starts_with("x"),
                         names_to = c("paired_id_semi"),
                         values_to = "Abundance") %>%
  mutate(recurrence = str_extract(paired_id_semi, 
                                  "[^0-9]+$")) %>%
  mutate(patient = str_remove_all(paired_id_semi, 
                                  "[^0-9]+$"))

long_mat_progen_2 <- long_mat_progen %>%
  mutate(paired_id = paste(patient, 
                           recurrence, 
                           sep = "_"))

paired_annot_uniq_progen <- long_mat_progen_2 %>%
  dplyr::select(patient, 
                recurrence, 
                paired_id) %>%
  distinct()

# wide peptide abundace matrix with NAs
expr_matrix_pept_progen <- dplyr::select(tmt_peptdata_frag_progen,
                             index,
                             peptide,
                             starts_with("x")) %>%
  dplyr::select(!starts_with("x6"))

expr_matrixnona_pept_progen <- expr_matrix_pept_progen %>%
  na.omit()

long_mat_pept_progen <- pivot_longer(expr_matrix_pept_progen,
                         cols = starts_with("x"),
                         names_to = c("paired_id_semi"),
                         values_to = "Abundance") %>%
  mutate(recurrence = str_extract(paired_id_semi, 
                                  "[^0-9]+$")) %>%
  mutate(patient = str_remove_all(paired_id_semi, 
                                  "[^0-9]+$"))

long_mat_pept_progen_2 <- long_mat_pept_progen %>%
  mutate(paired_id = paste(patient, 
                           recurrence, 
                           sep = "_"))

paired_annot_uniq_pept_progen <- long_mat_pept_progen_2 %>%
  dplyr::select(patient, 
                recurrence, 
                paired_id) %>%
  distinct()

## Annotated abundance data in long format ----

quant_annotated_progen <- left_join(long_mat_progen_2,
                             sample_annotation2) %>%
  mutate(Channel_mix = paste(mixture,
                             channel, 
                             sep = "_"))

quant_annotated_pept_progen <- left_join(long_mat_pept_progen_2,
                             sample_annotation2) %>%
  mutate(Channel_mix = paste(mixture,
                             channel, 
                             sep = "_"))

## Wide xpression matrices without NAs
expr_matnona_progen <- column_to_rownames(expr_matrix_progen,
                            "Protein") %>%
          as.matrix() %>%
          na.omit()

# peptide to protein mapping from combined_peptides.tsv 
protein2peptideall <- dplyr::select(identified_peptides_progen,
                                    protein, 
                                    Peptide = sequence) %>% 
          separate(col = protein,
                   into = c("sp", 
                            "Genes", 
                            "ensembl"),
                   sep = "\\|") %>%
          dplyr::select(Genes, 
                        Peptide) %>%
          distinct() %>%
          mutate(Peptide = str_remove_all(Peptide, 
                                          "\\[.+?\\]")) %>%
          mutate(Peptide = str_remove_all(Peptide, 
                                          "^n"))

# peptide to protein mappong from tmt abundance report (quantified and normalized)
protein2peptide_tmtrept <- tmt_peptdata_frag_progen %>%
  dplyr::select(index, 
                protein_id, 
                peptide)
```

# Localization of identified peptides and mapping to variant regions

```{r}
# list of identified peptides associated with 'variant' proteins
var_prots <- protein2peptideall %>%
                    mutate(variant = if_else(str_detect(Genes, 
                                                        "\\_"),
                           true = "variant",
                           false = "canonical")) %>%
                    filter(variant == "variant") %>% 
                    pull(Genes) %>%
                    unique()
```

```{r echo=TRUE}
variant_peptides_1 <- filter(protein2peptideall,
                             Genes %in% var_prots)

# prep fasta containing only identified protein sequences
fastavar <- seqinr::read.fasta(here("data/proteogenomics/sras_cellreports_2021_paper_razor/mix_1/protein.fas"), 
                            seqtype = "AA", 
                            as.string = TRUE)

fastavar2 <- seqinr::read.fasta(here("data/proteogenomics/sras_cellreports_2021_paper_razor/mix_2/protein.fas"), 
                            seqtype = "AA", 
                            as.string = TRUE)

fastavar3 <- seqinr::read.fasta(here("data/proteogenomics/sras_cellreports_2021_paper_razor/mix_3/protein.fas"), 
                            seqtype = "AA", 
                            as.string = TRUE)

fasta1 <- append(fastavar,
                fastavar2)

fasta <- append(fasta1,
                fastavar3)

fasta <- discard(fasta,
                 duplicated(names(fasta)))

```

```{r echo=TRUE}
if(!file.exists(here("results/proteogenomics/cellreps_2021/cleavage_annoated_peptides_razor.tsv"))){
  cleavage_annoated_peptides <- annotate_peptides(expr_mat = variant_peptides_1, 
                                                  fasta = fasta,
                                                  decoy_tag = "rev_")

  write_tsv(cleavage_annoated_peptides, "results/proteogenomics/cellreps_2021/cleavage_annoated_peptides_razor.tsv")
} else {
  cleavage_annoated_peptides <- read_tsv(here("results/proteogenomics/cellreps_2021/cleavage_annoated_peptides_razor.tsv"))
}

```

## Localize peptides within protein sequence and confirm identification of SAAVs

```{r echo=TRUE}
# tabulate location of the variant
annotated_w_variant_location <- cleavage_annoated_peptides %>%
          as_tibble() %>%
          mutate(protein_id_var = protein_id,
                 variant_type = if_else(str_detect(protein_id, 
                                                   "\\:"), 
                                        true = "other",
                                        false = "saav")) %>%
          mutate(protein_id = str_replace_all(protein_id, 
                                              "\\.", 
                                              "\\_")) %>%
          mutate(protein_id = str_replace_all(protein_id, 
                                              "\\:", 
                                              "\\_")) %>%
          separate(protein_id, 
                   into = c("ense_id", "variant_location"), 
                   sep = "\\_",
                   extra = "merge") %>%
                    na.omit() %>%
          mutate(variant_type = if_else(condition = str_detect(variant_location, 
                                                               "[0-9]$"),
                                        true = "other",
                                        false = variant_type))

# generate a long list of variant locations

locat2protdf <- annotated_w_variant_location %>%
  separate_rows(variant_location)

long2prot <- locat2protdf %>%
          mutate(variant_num_location = readr::parse_number(variant_location))

long2prot_worivar <- left_join(long2prot, 
                               dplyr::select(annotated_w_variant_location,
                                             protein_id_var, 
                                             ense_id)) %>%
  distinct()

annotated_wo_variant_location <- annotated_w_variant_location %>%
          dplyr::select(-variant_location)

# merge long table of variant locations with the annotation of peptide position within 
# their associated protein sequences

protein2varlocat <- left_join(annotated_wo_variant_location, 
                              long2prot_worivar) %>%
          distinct() %>%
          mutate(start_num = as.numeric(start_position),
                 end_num = as.numeric(end_position)) %>%
  # evaluate if the SAAV falls on the position of a variant sequence
          mutate(variant_peptide = variant_num_location >= as.numeric(start_position) & variant_num_location <= as.numeric(end_position)) %>% 
          dplyr::select(protein_id_var, 
                        ense_id, 
                        protein_description, 
                        Peptide, 
                        start_position, 
                        start_num,
                        end_position,
                        end_num, 
                        variant_location, 
                        variant_num_location, 
                        variant_peptide, 
                        variant_type)

pept2variant <- protein2varlocat %>%
  dplyr::select(Peptide, 
                variant_peptide, 
                variant_type,
                protein_id_var,
                ense_id,
                variant_location,
                peptide_start_position = start_position,
                peptide_end_position = end_position,
                protein_description) %>%
  distinct() %>%
  na.omit() %>%
  filter(variant_type == "saav") %>% 
  mutate(duplicated_peptide = duplicated(Peptide)) %>% 
  relocate(duplicated_peptide,
           .after = variant_type)
```

## Generate table of identified peptides mapping to SAAVs

```{r}
saavs_table <- pept2variant %>%
  filter(variant_peptide == TRUE) %>% 
  mutate(duplicated_peptide = duplicated(Peptide)) %>% 
  dplyr::rename(peptide = Peptide) %>%
  left_join(., index2genepeptide) %>%
  mutate(gene_variant = paste(gene, 
                              variant_location, 
                              sep = "_"))

```

```{r}
length(unique(saavs_table$Peptide))
```

```{r}
length(unique(saavs_table$protein_id_var))
```

```{r}
saavs_suppl_table <- saavs_table %>% 
  dplyr::select(gene_variant,
                variant_type,
                variant_location,
                peptide,
                peptide_start_position,
                peptide_end_position,
                protein_id_var,
                ense_id,
                gene,
                protein_description)
```

```{r}
write_tsv(x = saavs_suppl_table,
          file = here("suppl_tables/supplementary_table_5_proteogenomics_identified_saavs.tsv"))
```

# Visualize quantitative features of identified SAAVs

## Prep abundance matrices

```{r}
recurrencewo6 <- sample_annotationwo6$recurrence
patientwo6 <- sample_annotationwo6$patient
```

```{r echo=FALSE}
design_limmawo6 <- model.matrix(~patientwo6+recurrencewo6)

sample_annotationwo6 <- sample_annotationwo6 %>%
  mutate(paired_id_semi = paste(patient, 
                                recurrence, 
                                sep = ""))

rownames(design_limmawo6) <- sample_annotationwo6$paired_id_semi
```

```{r}
pep_prematwo6 <- dplyr::select(expr_matrix_pept_progen,
                               index, 
                               peptide, 
                               rownames(design_limmawo6))

pep_matwo6 <- dplyr::select(expr_matrix_pept_progen,
                            index, 
                            rownames(design_limmawo6))  %>%
  column_to_rownames("index") %>%
  as.matrix()

pep_matnona_wo6 <- pep_matwo6 %>%
  na.omit()
```

## Filter, merge and summarize SAAV abundances

```{r}
# extract indexing, peptide and sequence information from normalized data
index2genepeptide <- tmt_peptdata_frag_progen %>% 
  dplyr::select(index, 
                peptide, 
                gene, 
                protein_id)

# table of interesting features (saavs)
features1 <- pept2variant %>%
  dplyr::rename(peptide = Peptide) %>%
  left_join(., index2genepeptide) %>%
  na.omit()

interesting_features1 <- features1 %>%
                    filter(variant_peptide == TRUE)

# data frame of identified and TMT-integrator summarized SAAVs + annotation
protein2varlocat_saav <- protein2varlocat %>%
  dplyr::rename(peptide = Peptide) %>%
  filter(peptide %in% interesting_features1$peptide,
         variant_peptide == TRUE,
         variant_type == "saav") %>%
  left_join(., index2genepeptide) %>%
  mutate(gene_variant = paste(gene, 
                              variant_location, 
                              sep = "_"))

# mapping variant to index id
genevariant2index <- protein2varlocat_saav %>%
  dplyr::select(index, 
                gene_variant)

# abundance matrix of peptides mapping to variant locations
pep_premat_variantwo6 <- filter(pep_prematwo6,
                                peptide %in% protein2varlocat_saav$peptide) %>%
  left_join(., genevariant2index) %>%
  relocate(gene_variant) %>%
  mutate(duplicated_variant = duplicated(gene_variant)) %>%
  relocate(duplicated_variant,
           .after = peptide)

# summarize SAAVs (gene+variant location)
# i.e. take the median abundance of all the peptides mapping to the same SAAV

pep_premat_sumgenarwo6 <- pep_premat_variantwo6 %>%
  group_by(gene_variant) %>%
  summarise_at(vars(starts_with("x")), 
               .funs = function(x){median(x, 
                                          na.rm = TRUE)}) %>%
  ungroup()

full_info_sumgenvar <- pep_premat_sumgenarwo6 %>%
  left_join(., 
            protein2varlocat_saav) %>%
  filter(variant_type == "saav")

pep_matvariantswo6or <- pep_premat_sumgenarwo6 %>%
  column_to_rownames("gene_variant") %>%
  as.matrix()

pep_matvariantswo6 <- pep_premat_sumgenarwo6 %>%
  filter(str_detect(gene_variant, 
                    "CAP1_", 
                    negate = TRUE)) %>%
  column_to_rownames("gene_variant") %>%
  as.matrix()

pep_matvariantswo6nona <- pep_matvariantswo6 %>%
  na.omit()
```

### Generate summarized results tables

```{r}
write_tsv(protein2varlocat_saav,
          here("results/proteogenomics/cellreps_2021/all_peptides_mapping_to_saavs_gbm.tsv"))

write_tsv(pep_premat_sumgenarwo6,
          here("results/proteogenomics/cellreps_2021/summarized_median_abundance_of_saavs.tsv"))

write_tsv(full_info_sumgenvar,
          here("results/proteogenomics/cellreps_2021/summarized_median_abundance_of_saavs_full.tsv"))
```

The variant-containing peptides identified and accurately quantified and summarized are able to account for `r nrow(pep_premat_sumgenarwo6)` single amino acid mutations (SAAVs).

Among these identified SAAVs, `r nrow(pep_matvariantswo6nona)` were identified in all mixtures and samples.

## Heatmap of all SAAVs observed

```{r}
library(circlize)
```

```{r}
library(ComplexHeatmap)
```

```{r}
split_patstage <- str_split_fixed(string = colnames(pep_matvariantswo6nona), 
                                  pattern = "[:digit:]+", 
                                  n = 2)

patient <- str_extract(string = colnames(pep_matvariantswo6nona),
                       pattern = "[:digit:]+")

patientstage <- colnames(pep_matvariantswo6nona) %>%
  str_replace(., 
              pattern = "prim", 
              replacement = "Init") %>%
  str_replace(., 
              pattern = "rec", 
              replacement = "Rec")

stage <- split_patstage[,2] %>%
  if_else(condition = . == "prim",
          true = "Initial",
          false = "Recurrent")

pep_matvariantswo6or_hm <- pep_matvariantswo6or
pep_matvariantswo6nona_hm <- pep_matvariantswo6nona

colnames(pep_matvariantswo6or_hm) <- patientstage
colnames(pep_matvariantswo6nona_hm) <- patientstage
```

```{r fig.height=20, fig.width=7}
col_fun_or = colorRamp2(c(min(pep_matvariantswo6or_hm, 
                              na.rm = TRUE), 
                       median(pep_matvariantswo6or_hm, 
                              na.rm = TRUE), 
                       max(pep_matvariantswo6or_hm, 
                           na.rm = TRUE)), 
                     c("#2a9d8f", 
                       "white", 
                       "red"))

Heatmap(pep_matvariantswo6or_hm, 
        col = col_fun_or,
        column_split = stage, 
        row_names_gp = gpar(fontsize = 5),
        column_names_gp = gpar(fontsize = 5),
        cluster_rows = FALSE, cluster_columns = FALSE,
        heatmap_legend_param = list(
        title = "log2-\nNormalized\nAbundance"
    ))
```

All identified SAAVs

```{r fig.height=20, fig.width=7}
col_fun_or = colorRamp2(c(min(pep_matvariantswo6or_hm, 
                              na.rm = TRUE), 
                       median(pep_matvariantswo6or_hm, 
                              na.rm = TRUE), 
                       max(pep_matvariantswo6or_hm, 
                           na.rm = TRUE)), 
                     c("#2a9d8f", 
                       "white", 
                       "red"))

Heatmap(pep_matvariantswo6or_hm, 
        col = col_fun_or,
        column_split = patient, 
        row_names_gp = gpar(fontsize = 5.5),
        column_names_gp = gpar(fontsize = 7),
        cluster_rows = FALSE, cluster_columns = FALSE,
        heatmap_legend_param = list(
        title = "log2-\nNormalized\nAbundance"
    ))
```

## Heatmap of SAAVs observed on every sample

```{r fig.height=7.5, fig.width=7}
col_fun_sel = colorRamp2(c(min(pep_matvariantswo6or_hm, 
                               na.rm = TRUE), 
                       median(pep_matvariantswo6or_hm, 
                              na.rm = TRUE), 
                       max(pep_matvariantswo6or_hm, 
                           na.rm = TRUE)), 
                     c("#2a9d8f", 
                       "white", 
                       "red"))

Heatmap(pep_matvariantswo6nona_hm, 
        col = col_fun_sel,
        column_split = stage, 
        row_names_gp = gpar(fontsize = 7),
        column_names_gp = gpar(fontsize = 7),
        heatmap_legend_param = list(
        title = "log2-\nNormalized\nAbundance"
    ))
```

# Limma on SAAVs

## Prep design matrix

```{r}
design_limmawo6_saav <- model.matrix(~patient+stage)

sample_annotationwo6 <- sample_annotationwo6 %>%
  mutate(paired_id_semi = paste(patient, 
                                recurrence, 
                                sep = ""))

rownames(design_limmawo6_saav) <- sample_annotationwo6$paired_id_semi
```

## Prep abundance matrices {.unnumbered .unlisted}

```{r}
colnames(pep_matvariantswo6nona)
```

```{r}
rownames(design_limmawo6_saav)
```

## SAAV-level model fit {.unnumbered .unlisted}

```{r echo=TRUE, warning=FALSE, message=FALSE}
limmafit <- lmFit(pep_matvariantswo6nona, 
                  design = design_limmawo6_saav, 
                  method = 'robust')

limmafit <- eBayes(limmafit)

fit_mod_saavs <- topTable(limmafit, 
                        coef = "stageRecurrent", 
                        number = Inf, 
                        adjust.method = "BH") %>%
  mutate(Protein = rownames(.),
         Limma = "Robust - w Patient effect")

compar_tab_saavs <-  fit_mod_saavs %>% 
  rownames_to_column("gene_variant") 
```

```{r}
increased_recwo6_saavs <- compar_tab_saavs %>%
                    filter(logFC > 0,
                           adj.P.Val < 0.05) %>% 
  pull(gene_variant)

decreased_recwo6_saavs <- compar_tab_saavs %>%
                    filter(logFC < 0,
                           adj.P.Val < 0.05) %>% 
  pull(gene_variant)

```

```{r}
length(increased_recwo6_saavs)
```

```{r}
length(decreased_recwo6_saavs)
```

Only 1 SAAV is observed as differentially abundant between Recurrent and Initial tumors.

## Volcano SAAVs

```{r}
diff_saavs <- compar_tab_saavs %>%
                    filter(adj.P.Val < 0.05)

size <- 1.5

volcano_saavs <- ggplot(data = compar_tab_saavs,
                      mapping = aes(x = logFC, 
                                    y = -log10(adj.P.Val))) +
      geom_point(data = compar_tab_saavs %>% 
                   filter(logFC > 0,
                          adj.P.Val < 0.05),
                 mapping = aes(x = logFC, 
                               y = -log10(adj.P.Val)), 
                 color = "red",
                 size = size)+
      geom_point(data = compar_tab_saavs %>% 
                   filter(logFC < -0,
                          adj.P.Val < 0.05),
                 mapping = aes(x = logFC, 
                               y = -log10(adj.P.Val)), 
                 color = "red",
                 size = size) +
      geom_point(data = compar_tab_saavs %>% 
                   filter(logFC > 0,
                          adj.P.Val > 0.05),
                 mapping = aes(x = logFC, 
                               y = -log10(adj.P.Val)), 
                 color = "#2a9d8f",
                 size = size) +
      geom_point(data = compar_tab_saavs %>% 
                   filter(logFC < -0,
                          adj.P.Val > 0.05),
                 mapping = aes(x = logFC, 
                               y = -log10(adj.P.Val)), 
                 color = "#2a9d8f",
                 size = size) +
      geom_hline(yintercept = -log10(0.05),
                 color = "red", 
                 linetype = "dashed") +
      ggrepel::geom_text_repel(data = diff_saavs,
                               aes(label = gene_variant), 
                               size = 4) +
      xlab("logFC - Recurrent / Primary") + 
      theme(axis.text.x = element_text(hjust = 0.5, 
                                       vjust = 0, 
                                       size = 6, 
                                       angle = 360),
            axis.text.y = element_text(hjust = 0.95, 
                                       vjust = 0.2, 
                                       size = 8),
            panel.background = element_blank(),
            panel.grid.major = element_blank(),
            panel.border = element_rect(colour = "black", 
                                        fill = NA, 
                                        size = 0.5),
            axis.title = element_text(size = 8),
            legend.text = element_text(size = 6),
            legend.title = element_text(size = 8),
            legend.key.height = unit(3, 
                                    'mm'),
            legend.key.width = unit(3, 
                                   'mm'),
            legend.position = "bottom") 
```

```{r fig.width=8, fig.height=6}
volcano_saavs
```

```{r echo=FALSE}
ggsave(plot = volcano_saavs, 
       filename = here::here("figures/volcanosaavs_proteomics_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 50)

ggsave(plot = volcano_saavs, 
       filename = here::here("figures/volcanosaavs_proteomics_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 63,
       height = 50)
```

## Boxplot of differentially abundant SAAV

```{r}
prelongsaav <- pep_matvariantswo6nona %>%
  as.data.frame() %>%
  rownames_to_column("gene_variant") %>%
   pivot_longer(cols = starts_with("x"),
                         names_to = c("patient"),
                         values_to = "Abundance") %>%
  mutate(Stage = str_extract(patient, 
                             "[^0-9]+$")) %>%
  mutate(Patient = str_remove_all(patient, 
                                  "[^0-9]+$")) %>%
  filter(gene_variant == increased_recwo6_saavs)
```

```{r}
diff_abund_saav <- ggplot(prelongsaav, 
                    aes(x = Stage, 
                        y = Abundance, 
                        fill = Stage, 
                        cex.axis = 1.5)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", 
               stackdir = "center", 
               dotsize = 0.2) +
  # Box plot with jittered points
  # 0.2 : degree of jitter in x direction
  # geom_jitter(shape=16, position=position_jitter(0.2))
  ylab("log2-Normalized Abundance") +
  geom_signif(
    comparisons = list(c("Initial", 
                         "Recurrence")),
    map_signif_level = TRUE
  ) + 
  stat_compare_means(method="t.test") +
  ggtitle(increased_recwo6_saavs) +
  theme(axis.text.x = element_text(hjust = 0.5, 
                                   vjust = 0, 
                                   size = 6, 
                                   angle = 360),
        axis.text.y = element_text(hjust = 0.95, 
                                   vjust = 0.2, 
                                   size = 8),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", 
                                    fill = NA, 
                                    size = 0.5),
        axis.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.key.height = unit(3, 
                                'mm'),
        legend.key.width = unit(3, 
                               'mm'),
        legend.position = "bottom")

```

## Boxplot DA SAAV

```{r}
  slim_data2saav <- pivot_wider(prelongsaav %>% dplyr::select(-patient), 
                          values_from = Abundance, 
                          names_from = Stage)

  slim_data3saav <- slim_data2saav %>% 
    mutate(Abs_diff = rec-prim) %>%
    mutate(rank = row_number(Abs_diff)) %>% 
    dplyr::rename(Initial = prim, 
                  Recurrent = rec) %>%
    ungroup()

```

```{r fig.width=7, fig.height=3.5}
ggpaired(slim_data3saav,
         cond1 = "Initial",
         cond2 = "Recurrent", 
         fill = "condition",
         ylab = "log2-Normalized Abundance",
         label = NULL,
         repel = TRUE, 
         point.size = 0.2, 
         line.size = 0.1) +
      theme(axis.text.x = element_text(hjust = 0.5, 
                                       vjust = 0,
                                       size = 5, 
                                       angle = 360),
            axis.text.y = element_text(hjust = 0.95, 
                                       vjust = 0.2, 
                                       size = 5),
            panel.background = element_blank(),
            panel.grid.major = element_blank(),
            panel.border = element_rect(colour = "black", 
                                        fill = NA, 
                                        size = 0.5),
            axis.title = element_text(size = 8),
            legend.text = element_text(size = 6),
            legend.title = element_text(size = 8),
            legend.key.height= unit(3, 
                                    'mm'),
            legend.key.width= unit(3, 
                                   'mm'),
            legend.position = "bottom",
            strip.text = element_text(hjust = 0.5, 
                                      vjust = 0, 
                                      size = 8, 
                                      angle = 360)) + 
  ggtitle(increased_recwo6_saavs)
```

```{r}
paired_box_saavs <- ggpaired(slim_data3saav,
         cond1 = "Initial",
         cond2 = "Recurrent", 
         fill = "condition",
         ylab = "log2-Normalized Abundance",
         label = NULL,
         repel = TRUE, 
         point.size = 0.2, 
         line.size = 0.1) +
      theme(axis.text.x = element_text(hjust = 0.5, 
                                       vjust = 0, 
                                       size = 5, 
                                       angle = 360),
            axis.text.y = element_text(hjust = 0.95, 
                                       vjust = 0.2, 
                                       size = 5),
            panel.background = element_blank(),
            panel.grid.major = element_blank(),
            panel.border = element_rect(colour = "black", 
                                        fill = NA, 
                                        size = 0.5),
            axis.title = element_text(size = 8),
            legend.text = element_text(size = 6),
            legend.title = element_text(size = 8),
            legend.key.height = unit(3, 
                                    'mm'),
            legend.key.width = unit(3, 
                                   'mm'),
            legend.position = "bottom",
            strip.text = element_text(hjust = 0.5, 
                                      vjust = 0, 
                                      size = 8, 
                                      angle = 360)) 
```

```{r echo=FALSE}
ggsave(plot = paired_box_saavs, 
       filename = here::here("figures/paired_box_saav_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 61,
       height = 50)

ggsave(plot = paired_box_saavs, 
       filename = here::here("figures/paired_box_saav_gbm.eps"), 
       device = "eps",
       units = "mm",
       width = 61,
       height = 50)
```