### 2023-08-25 reports/gbm_recurrence_mine_single_cell_data_preprocessing.qmd
single_cell_asah1_expression_mining

---
title: "GBM Recurrence - Mine Single Cell data"
author: "Miguel Cosenza-Contreras"
format:
  gfm:
    toc: true 
    toc-depth: 4 # default is 3
    number-sections: true
    execute:
      output: true
always_allow_html: true
editor: source
---

```{r}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

source(here::here("scr/helper_functions.R"))

## Required packages ----
library(tidyverse)
library(mixOmics)
library(fs)
library(kableExtra)
library(sva)
library(limma)
library(naniar)
library(missForest)
library(clusterProfiler)
library(ReactomePA)
library(org.Hs.eg.db)
library(DT)
library(here)
library(janitor)
library(drawProteins)
library(seqinr)
library(ggpubr)
library(ggrepel)
library(extrafont)
extrafont::loadfonts(device = "win")

theme_set(theme(axis.text.x = element_text(hjust = 0.5, 
                                           vjust = 0, 
                                           size = 6, 
                                           angle = 360),
                axis.text.y = element_text(hjust = 0.5, 
                                           vjust = 0, 
                                           size = 6),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", 
                                            fill = NA, 
                                            size = 1.5),
                axis.title=element_text(size = 8),
                legend.title = element_text(size = 8),
                legend.key.height = unit(3, 
                                         'mm'),
                legend.key.width = unit(3, 
                                        'mm'),
                legend.position = "bottom",
                text = element_text(family = "Helvetica")))
```

# Initial data loading and wrangling

Load the matrix and annotation

```{r}
annot_1 <- read_csv(here::here("data/revision_single_cell_data/annot_Human_R_GBM_Full.csv"))

annot_2_ncbi <- read_csv(here::here("data/revision_single_cell_data/from_ncbi/GSM4972210_annot.Human.GBM.R1_2_3_4_4nc.csv.gz"))

# matrix data

matrix_ncbi <- read_csv(here::here("data/revision_single_cell_data/from_ncbi/GSM4972210_Human.GBM.R1_2_3_4_4nc.filtered.gene.bc.matrix.csv"))
```

```{r}
View(annot_1)
#View(annot_2_ncbi)
#View(matrix_ncbi)
```

```{r}
#head(str(matrix_ncbi))
#names(matrix_ncbi)[1]
```

# Normalize (CPM)

```{r}
library(scuttle)
```

# log2 transform

```{r}
matrix_ncbi <- matrix_ncbi %>%
  dplyr::rename('gene' = '...1') %>%
  column_to_rownames("gene") %>% 
  as.matrix()
```

```{r}
cpm_mat <- calculateCPM(matrix_ncbi)
mat_cpm_log2 <- log2(cpm_mat + 1)
rm(cpm_mat)
```

```{r}
mat_cpm_log2[1:15, 1:15]
```

# Wrangle

```{r}
mat_cpm_df <- mat_cpm_log2 %>% 
  as.data.frame() %>%
  rownames_to_column("gene") %>%
  relocate(gene, everything()) 
```

```{r}
# ASAH1 expression
mat_log2_df_asah1 <- mat_cpm_df %>%
  filter(gene == "ASAH1")
```

```{r}
long_log2_df_asah1 <- mat_log2_df_asah1 %>%
  pivot_longer(cols = -gene, 
               names_to = "cell", 
               values_to = "asah1_expression") 
```

Merge with annotation

```{r}
long_log2_df_asah1_annot <- long_log2_df_asah1 %>%
  left_join(annot_1, by = c("cell" = "cell"))
```

```{r}
#View(long_log2_df_asah1_annot)
```

```{r}
saveRDS(long_log2_df_asah1_annot,
        here::here("rds/long_cpm_scrnaseq_asah1_annot.rds"))
```