### 2023-08-26 reports/gbm_recurrence_mine_single_cell_data_visualization.qmd
sc_asah1_expression_violin

---
title: "GBM Recurrence - Mine Single Cell data"
author: "Miguel Cosenza-Contreras"
format:
  gfm:
    toc: true 
    toc-depth: 4 # default is 3
    number-sections: true
    execute:
      output: true
always_allow_html: true
editor: source
---

```{r}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

## Required packages ----
library(tidyverse)
library(ggpubr)
library(ggrepel)
library(cowplot)

theme_set(theme(axis.text.x = element_text(hjust = 0.5, 
                                           vjust = 0, 
                                           size = 12, 
                                           angle = 360),
                axis.text.y = element_text(hjust = 0.5, 
                                           vjust = 0, 
                                           size = 12),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", 
                                            fill = NA, 
                                            linewidth = 1.5),
                axis.title=element_text(size = 12),
                legend.title = element_text(size = 12),
                legend.text = element_text(size = 11),
                legend.key.height = unit(4, 
                                         'mm'),
                legend.key.width = unit(4, 
                                        'mm'),
                legend.key.size = unit(10,
                                        'mm'),
                legend.position = "bottom"))
```

# Initial data loading and wrangling

Load the annotation with ASAH1 expression data

```{r}
annotated_with_asah1 <- readRDS(here::here("rds/long_cpm_scrnaseq_asah1_annot.rds"))
```

# Generate scatter plot

```{r}
umap_asah1_expr <- ggplot(annotated_with_asah1, 
       aes(x = x, 
           y = y, 
           color = asah1_expression)) +
  geom_point(alpha = 0.5) +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "ASAH1 expression\n[log2(CPM)]") +
  theme(legend.position = "bottom")
```

```{r}
cell_type_centroid <- annotated_with_asah1 %>%
  group_by(cluster) %>%
  summarise(x = mean(x), 
            y = mean(y)) %>%
  ungroup()
```

```{r}
umap_cell_type <- ggplot(annotated_with_asah1, 
       aes(x = x, 
           y = y, 
           color = cluster)) +
  geom_point(alpha = 0.5) +
  geom_text(data = cell_type_centroid, 
            aes(x = x, 
                y = y, 
                label = cluster), 
            size = 5,
            color = "black") +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "Cell type") +
  theme(legend.position = "bottom") 
```

```{r}
umap_grid <- plot_grid(umap_cell_type, 
                       umap_asah1_expr, 
                       ncol = 2)
```

```{r fig.width=20, fig.height=10}
umap_grid
```

```{r}
ggsave(here::here("figures/umap_grid_asah1_expression_scRNAseq.pdf"), 
       umap_grid, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_asah1_expression_scRNAseq.png"), 
       umap_grid, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

# Violin plot

```{r}
annotated_with_asah14violin <- annotated_with_asah1 %>%
  filter(asah1_expression > 2.5) 
```

```{r}
violin_asahl_expr_celltype <- ggplot(annotated_with_asah14violin, 
       aes(x = cluster, 
           y = asah1_expression,
           fill = cluster)) +
  geom_violin() +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "Cell type", 
       y = "ASAH1 expression [log2(CPM)]") +
  theme(legend.position = "none",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6)
        ) 
```

```{r}
violin_asahl_expr_celltype
```

```{r}
umap_grid_w_violin <- plot_grid(umap_cell_type, 
                                umap_asah1_expr, 
                                violin_asahl_expr_celltype,
                                ncol = 2)
```

```{r fig.width=20, fig.height=10}
umap_grid_w_violin
```

```{r}
ggsave(here::here("figures/umap_grid_asah1_expression_violin_scRNAseq.pdf"), 
       umap_grid_w_violin, 
       width = 34, 
       height = 34,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_asah1_expression_violin_scRNAseq.png"), 
       umap_grid_w_violin, 
       width = 34, 
       height = 34,
       units = "cm")
```

```{r}
dev.off()
```