### 2023-08-26 reports/gbm_recurrence_cox_phm_incl_age_sex_revision.qmd
cox_phm_w_sex_age_covar_revision

---
title: "GBM Recurrence - Cox Proportional Hazards Model"
author: "Justin Sing - Miguel Cosenza-Contreras"
format:
  gfm:
    toc: true 
    toc-depth: 4 # default is 3
    number-sections: true
    execute:
      output: true
always_allow_html: true
editor: source
---

```{r}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

## Required packages ----
library(tidyverse)
library(ggpubr)
library(ggrepel)
library(cowplot)
library(here)
library(survival)
library(gt)

theme_set(theme(axis.text.x = element_text(hjust = 0.5, 
                                           vjust = 0, 
                                           size = 12, 
                                           angle = 360),
                axis.text.y = element_text(hjust = 0.5, 
                                           vjust = 0, 
                                           size = 12),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", 
                                            fill = NA, 
                                            linewidth = 1.5),
                axis.title=element_text(size = 12),
                legend.title = element_text(size = 12),
                legend.text = element_text(size = 11),
                legend.key.height = unit(4, 
                                         'mm'),
                legend.key.width = unit(4, 
                                        'mm'),
                legend.key.size = unit(10,
                                        'mm'),
                legend.position = "bottom"))
```

# Initial data loading and wrangling

Load the annotation with ASAH1 expression data

```{r}
ttr_surv_data <- read_tsv(here::here("data/elisa_plasma_proteomics/ttr_survival_data.tsv"))
```

# Survival analysis

## Cox-PHM - including only sex (original analysis)

Fitting the model:

```{r}
## COXPH ----
#ttr_surv_data
fit <- coxph(Surv(Time.to.reccurence..days., status) ~ sex +  ASAH1.ng.ml + SYNM.ng.ml + GPNMB.ng.ml + MMP.9.ng.ml, data = ttr_surv_data)
```

Model summary:

```{r}
print(summary(fit))
```

Observation: Sex does not seem to represent a significant variable in the model.

### Genarate tabular and graphical summaries

```{r}
library(finalfit)

dependent_var <- "Surv(Time.to.reccurence..days., status)"
explanatory_vars <- c("ASAH1.ng.ml", "SYNM.ng.ml", "GPNMB.ng.ml", "MMP.9.ng.ml")

ttr_surv_data %>% 
    coxphmulti(dependent_var, explanatory_vars) %>% 
    cox.zph() %>% 
    {zph_result <<- .} %>% 
    plot(var=3)

#zph_result

ttr_surv_data %>%
  finalfit(dependent_var, explanatory_vars) %>%
  gt()

ttr_surv_data %>%
  hr_plot(dependent_var, explanatory_vars, 
          dependent_label="Time to Recurrence")

ttr_surv_data %>%
  ff_plot(dependent_var, explanatory_vars, 
          dependent_label="Time to Recurrence")
```

## Cox-PHM - including only sex and age (revision version analysis)

This section was created as part of the revision process, to check for the effect of age and sex in the model.

```{r}
## COXPH ----
#ttr_surv_data
fit2 <- coxph(Surv(Time.to.reccurence..days., status) ~ sex + Age.at.surgery..years. + ASAH1.ng.ml + SYNM.ng.ml + GPNMB.ng.ml + MMP.9.ng.ml, data = ttr_surv_data)
```

Model summary:

```{r}
print(summary(fit2))
```

```{r}
?coxphmulti
```

Observation: p-value for age is 0.0878 in the model; while ASAH1 concentration is stil below 0.05.

```{r}
dependent_var <- "Surv(Time.to.reccurence..days., status)"
explanatory_vars2 <- c("ASAH1.ng.ml", "SYNM.ng.ml", "GPNMB.ng.ml", "MMP.9.ng.ml",
                      "sex", "Age.at.surgery..years.") # includes sex and age

ttr_surv_data %>% 
    coxphmulti(dependent_var, explanatory_vars2) %>% 
    cox.zph() %>% 
    {zph_result2 <<- .} %>% 
    plot(var=3)

#zph_result2

ttr_surv_data %>%
  finalfit(dependent_var, explanatory_vars2) %>%
  gt()

ttr_surv_data %>%
  hr_plot(dependent_var, explanatory_vars2, dependent_label="Time to Recurrence")

ttr_surv_data %>%
  ff_plot(dependent_var, explanatory_vars2, dependent_label="Time to Recurrence")
```

When including Age and Sex in the model, ASAH1 concentrations are still significantly associated with the time to recurrence. Age and sex are not significant in the model.

```{r}
ggsave(here::here("figures/cox_phm_summary_rev_version_w_age_n_sex.pdf"), 
       ttr_surv_data %>%
  ff_plot(dependent_var, explanatory_vars2, dependent_label="Time to Recurrence"), 
       width = 20, 
       height = 20,
       units = "cm")
```

```{r}
ggsave(here::here("figures/cox_phm_summary_rev_version_w_age_n_sex.png"), 
       ttr_surv_data %>%
  ff_plot(dependent_var, explanatory_vars2, dependent_label="Time to Recurrence"), 
       width = 20, 
       height = 20,
       units = "cm")
```

```{r}
dev.off()
```