### 2023-08-26 reports/gbm_recurrence_proteogenomics_mindy_asso_w_ttr.qmd
exploration_assoc_mindy_saav_w_ttr

---
title: "GBM Recurrence Proteogenomics - MINDY assoc. with TTR"
author: "Miguel Cosenza-Contreras"
format:
  gfm:
    toc: true 
    toc-depth: 4 # default is 3
    number-sections: true
    execute:
      output: true
always_allow_html: true
editor: source
---

```{r}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

source(here::here("scr/helper_functions.R"))

## Required packages ----
library(tidyverse)
library(mixOmics)
library(fs)
library(kableExtra)
library(sva)
library(limma)
library(naniar)
library(missForest)
library(clusterProfiler)
library(ReactomePA)
library(org.Hs.eg.db)
library(DT)
library(here)
library(janitor)
library(drawProteins)
library(seqinr)
library(ggpubr)
library(ggrepel)
library(extrafont)
extrafont::loadfonts(device = "win")

theme_set(theme(axis.text.x = element_text(hjust = 0.5, 
                                           vjust = 0, 
                                           size = 6, 
                                           angle = 360),
                axis.text.y = element_text(hjust = 0.5, 
                                           vjust = 0, 
                                           size = 6),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", 
                                            fill = NA, 
                                            size = 1.5),
                axis.title=element_text(size = 8),
                legend.title = element_text(size = 8),
                legend.key.height = unit(3, 
                                         'mm'),
                legend.key.width = unit(3, 
                                        'mm'),
                legend.position = "bottom",
                text = element_text(family = "Helvetica")))

```

# Get MINDY expression data

Load SAAV abundance matrix

```{r}
saav_abund_mat <- read_tsv(here("results/proteogenomics/cellreps_2021/summarized_median_abundance_of_saavs.tsv"))
```

```{r}
saav_abund_long <- saav_abund_mat %>% 
    pivot_longer(cols = -gene_variant, 
                 names_to = "sample_label", 
                 values_to = "MINDY1_T433K") %>% 
    mutate(recurrence = case_when(str_detect(sample_label, "prim") ~ "prim",
                                  str_detect(sample_label, "rec") ~ "rec",
                                  TRUE ~ "NA"),
           patient = case_when(str_detect(sample_label, "prim") ~ str_replace(sample_label, "prim", ""),
                                 str_detect(sample_label, "rec") ~ str_replace(sample_label, "rec", ""),
                                 TRUE ~ "NA")) %>%
    mutate(paired_id = paste(patient, 
                             recurrence, 
                             sep = "_")) %>%
    dplyr::distinct() %>%
    filter(gene_variant %in% c("FAM63A_T433K")) 
```

# Associate MYNDY expression with Time to Recurrence

## Load and prepare patient data

```{r echo=FALSE}
patient_file <- here("data/Patients_information_FINAL.xlsx")
patient_table <- readxl::read_xlsx(path = patient_file)
```

```{r}
patients_proteomics <- patient_table %>%
  filter(`Mass Spectrometry` == "X") %>%
  mutate(latency = parse_number(`Latency (days)`), 
         status = case_when(str_detect(`Latency (days)`, "alive") ~ 1,
                            TRUE ~ 2)) %>%
  dplyr::select(Sex, 
                `Age at surgery (years)`, 
                `Time-to-reccurence (days)`,
                latency,
                status,
                `Type of resection`,
                `Tumor localization`, 
                Treatment,
                sample_id = `LAB-ID`) %>%
  mutate(`Time-to-reccurence (days)` = as.numeric(`Time-to-reccurence (days)`))
```

## Sample annotation

```{r}
sample_annotation <- read_csv(here("data/sample_annotation.csv"))

sample_annotation2 <- sample_annotation %>%
  mutate(patient = paste("x", 
                         patient, 
                         sep = ""),
         recurrence = case_when(recurrence == "initial" ~ "prim",
                                recurrence == "recurrent" ~ "rec",
                                TRUE ~ recurrence)) %>%
  mutate(paired_id = paste(patient, 
                           recurrence, 
                           sep = "_")) %>%
  filter(recurrence %in% c("prim", 
                           "rec"))
```

## What's the association of MINDY1_T433K at Initial stage with time to recurrence?

```{r}
quant_annot_intr_init <- saav_abund_long %>%
    left_join(sample_annotation2) %>%
    left_join(patients_proteomics) %>%
    filter(!is.na(`Time-to-reccurence (days)`)) %>% 
    dplyr::mutate(MINDY1_T433K_abundance = ifelse(MINDY1_T433K >= mean(MINDY1_T433K), 
                                 "high", 
                                 "low"))

labid2patientn <- quant_annot_intr_init %>% 
  dplyr::select(patient, sample_id) %>% 
  dplyr::distinct()

patients_proteomics <- left_join(labid2patientn, 
                                 patients_proteomics)
```

```{r}
library(survival)
library(survminer)
```

```{r}
fit1 <- survfit(Surv(`Time-to-reccurence (days)`) ~ MINDY1_T433K_abundance, 
                data = quant_annot_intr_init)
print(fit1)
# Change color, linetype by strata, risk.table color by strata
surv_p1 <- ggsurvplot(fit1,
          pval = TRUE, conf.int = TRUE,
          risk.table = FALSE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          ) +
  labs(x = "Time to Recurrence (Days)", 
       y = "Time to Recurrence Probability")

```

```{r}
surv_p1
```

# Cox proportional hazards model on protein expression vs Time to Recurrence

## Checking Assumptions

```{r}
View(quant_annot_intr_init)
names(quant_annot_intr_init)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=18, fig.width=21}
# fit <- coxph(Surv(Time.to.reccurence..days., status) ~ sex + ASAH1.ng.ml + SYNM.ng.ml + GPNMB.ng.ml + MMP.9.ng.ml, data = ttr_surv_data)
# print(summary(fit))

library(finalfit)

quant_annot_intr_init2 <- quant_annot_intr_init %>%
  mutate(ttr_days = `Time-to-reccurence (days)`)

dependent_var <- "Surv(`Time-to-reccurence (days)`, status)"
explanatory_vars <- c("MINDY1_T433K", "Sex", "ttr_days")

dependent_var
explanatory_vars

quant_annot_intr_init2 %>% 
    coxphmulti(dependent_var, explanatory_vars) %>% 
    cox.zph() %>% 
    {zph_result <<- .}

```

## Model results

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4.5, fig.width=8}
library(globaltest)

quant_annot_intr_init2 %>%
  ff_plot(dependent_var, 
          explanatory_vars, 
          dependent_label = "Time to Recurrence")
```

```{r}
explanatory_vars_univar <- c("MINDY1_T433K")
```

```{r}
quant_annot_intr_init2 %>%
  ff_plot(dependent_var, 
          explanatory_vars_univar, 
          dependent_label = "Time to Recurrence")
```

```{r}
ggsave(here::here("figures/cox_phm_mindy_T433K_summary_rev_version_univariate.pdf"), 
       quant_annot_intr_init2 %>%
  ff_plot(dependent_var, 
          explanatory_vars_univar, 
          dependent_label = "Time to Recurrence"), 
       width = 20, 
       height = 10,
       units = "cm")
```

```{r}
ggsave(here::here("figures/cox_phm_mindy_T433K_summary_rev_version_w_sex_n_age.pdf"), 
       quant_annot_intr_init2 %>%
  ff_plot(dependent_var, 
          explanatory_vars, 
          dependent_label = "Time to Recurrence"), 
       width = 20, 
       height = 10,
       units = "cm")
```