### 2023-10-24 reports/gbm_recurrence_mine_single_cell_data_visualization.qmd

---
title: "GBM Recurrence - Mine Single Cell data"
author: "Miguel Cosenza-Contreras"
format:
  gfm:
    toc: true 
    toc-depth: 4 # default is 3
    number-sections: true
    execute:
      output: true
always_allow_html: true
editor: source
---

```{r}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)

## Required packages ----
library(tidyverse)
library(ggpubr)
library(ggrepel)
library(cowplot)
library(scales)    # Install ggnewscale package
library(ggnewscale)    
library(ggrastr)

theme_set(theme(axis.text.x = element_text(hjust = 0.5, 
                                           vjust = 0, 
                                           size = 12, 
                                           angle = 360),
                axis.text.y = element_text(hjust = 0.5, 
                                           vjust = 0, 
                                           size = 12),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey"),
                panel.border = element_rect(colour = "black", 
                                            fill = NA, 
                                            linewidth = 1.5),
                axis.title=element_text(size = 12),
                legend.title = element_text(size = 12),
                legend.text = element_text(size = 11),
                legend.key.height = unit(4, 
                                         'mm'),
                legend.key.width = unit(4, 
                                        'mm'),
                legend.key.size = unit(10,
                                        'mm'),
                legend.position = "bottom"))
```

# Initial data loading and wrangling

Load the annotation with ASAH1 expression data

```{r}
annotated_with_asah1 <- readRDS(here::here("rds/long_cpm_scrnaseq_asah1_annot.rds"))
```

# Generate scatter plot for ASAH1 expression

## rGBM

```{r}
cell_type_centroid <- annotated_with_asah1 %>%
  group_by(cluster) %>%
  summarise(x = mean(x), 
            y = mean(y)) %>%
  ungroup()
```

```{r}
umap_asah1_expr <- ggplot(annotated_with_asah1, 
       aes(x = x, 
           y = y, 
           color = asah1_expression)) +
  ggrastr::rasterize(geom_point(alpha = 0.5)) +
  geom_text(data = cell_type_centroid, 
            aes(x = x, 
                y = y, 
                label = cluster), 
            size = 5,
            color = "black") +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "ASAH1 expression\n[log2(CPM)]",
       title = "Recurrent GBM") +
  theme(legend.position = "bottom")
```

```{r}
cell_type_centroid <- annotated_with_asah1 %>%
  group_by(cluster) %>%
  summarise(x = mean(x), 
            y = mean(y)) %>%
  ungroup()
```

```{r}
umap_cell_type <- ggplot(annotated_with_asah1, 
       aes(x = x, 
           y = y, 
           color = cluster)) +
  ggrastr::rasterize(geom_point(alpha = 0.5)) +
  geom_text(data = cell_type_centroid, 
            aes(x = x, 
                y = y, 
                label = cluster), 
            size = 5,
            color = "black") +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "Cell type",
       title = "Recurrent GBM") +
  theme(legend.position = "bottom") 
```

```{r}
umap_grid <- plot_grid(umap_cell_type, 
                       umap_asah1_expr, 
                       ncol = 2)
```

```{r fig.width=20, fig.height=10}
#umap_grid
```

```{r}
ggsave(here::here("figures/umap_grid_asah1_expression_scRNAseq.pdf"), 
       umap_grid, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_asah1_expression_scRNAseq.png"), 
       umap_grid, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

## Violin plot

```{r}
annotated_with_asah14violin <- annotated_with_asah1 %>%
  filter(asah1_expression > 2.5) 
```

```{r}
violin_asahl_expr_celltype <- ggplot(annotated_with_asah14violin, 
       aes(x = cluster, 
           y = asah1_expression,
           fill = cluster)) +
  geom_violin() +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "Cell type", 
       y = "ASAH1 expression [log2(CPM)]") +
  theme(legend.position = "none",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6)
        ) 
```

```{r}
#violin_asahl_expr_celltype
```

```{r}
umap_grid_w_violin <- plot_grid(umap_cell_type, 
                                umap_asah1_expr, 
                                violin_asahl_expr_celltype,
                                ncol = 2)
```

```{r fig.width=20, fig.height=20}
umap_grid_w_violin
```

```{r}
ggsave(here::here("figures/umap_grid_asah1_expression_violin_scRNAseq.pdf"), 
       umap_grid_w_violin, 
       width = 34, 
       height = 34,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_asah1_expression_violin_scRNAseq.png"), 
       umap_grid_w_violin, 
       width = 34, 
       height = 34,
       units = "cm")
```

```{r}
dev.off()
```

## iGBM

```{r}
annotated_with_asah1_nd <- readRDS(here::here("rds/long_cpm_scrnaseq_asah1_annot_nd.rds"))
```

```{r}
cell_type_centroid_nd <- annotated_with_asah1_nd %>%
  group_by(cluster) %>%
  summarise(x = mean(x), 
            y = mean(y)) %>%
  ungroup()
```

```{r}
umap_cell_type_nd <- ggplot(annotated_with_asah1_nd, 
       aes(x = x, 
           y = y, 
           color = cluster)) +
  ggrastr::rasterize(geom_point(alpha = 0.5)) +
  geom_text(data = cell_type_centroid_nd, 
            aes(x = x, 
                y = y, 
                label = cluster), 
            size = 5,
            color = "black") +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "Cell type",
       title = "Newly diagnosed GBM") +
  theme(legend.position = "bottom") 
```

# Generate scatter plot MPO expression

```{r}
annotated_with_mpo <- readRDS(here::here("rds/long_cpm_scrnaseq_mpo_annot.rds"))
```

```{r}
umap_mpo_expr <- ggplot(annotated_with_mpo, 
       aes(x = x, 
           y = y, 
           color = mpo_expression)) +
  geom_point(alpha = 0.5) +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "MPO expression\n[log2(CPM)]") +
  theme(legend.position = "bottom")
```

```{r}
umap_grid_mpo <- plot_grid(umap_cell_type, 
                           umap_mpo_expr, 
                           ncol = 2)
```

```{r fig.width=20, fig.height=10}
#umap_grid_mpo
```

```{r}
ggsave(here::here("figures/umap_grid_mpo_expression_scRNAseq.pdf"), 
       umap_grid_mpo, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_mpo_expression_scRNAseq.png"), 
       umap_grid_mpo, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

## Violin plot

```{r}
annotated_with_mpo4violin <- annotated_with_mpo %>%
  filter(mpo_expression > 0) 
```

```{r}
violin_mpo_expr_celltype <- ggplot(annotated_with_mpo4violin, 
       aes(x = cluster, 
           y = mpo_expression,
           fill = cluster)) +
  geom_violin() +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "Cell type", 
       y = "MPO expression [log2(CPM)]") +
  theme(legend.position = "none",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6)
        ) 
```

```{r}
#violin_mpo_expr_celltype
```

```{r}
umap_grid_w_violin_mpo <- plot_grid(umap_cell_type, 
                                    umap_mpo_expr, 
                                    violin_mpo_expr_celltype,
                                    ncol = 2)
```

```{r fig.width=20, fig.height=20}
umap_grid_w_violin_mpo
```

```{r}
ggsave(here::here("figures/umap_grid_mpo_expression_violin_scRNAseq.pdf"), 
       umap_grid_w_violin_mpo, 
       width = 34, 
       height = 34,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_mpo_expression_violin_scRNAseq.png"), 
       umap_grid_w_violin_mpo, 
       width = 34, 
       height = 34,
       units = "cm")
```

# Generate scatter plot for CD14 expression

## rGBM

```{r}
annotated_with_cd14 <- readRDS(here::here("rds/long_cpm_scrnaseq_cd14_annot.rds"))
```

```{r}
umap_cd14_expr <- ggplot(annotated_with_cd14, 
       aes(x = x, 
           y = y, 
           color = cd14_expression)) +
  rasterize(geom_point(alpha = 0.5)) +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "CD14 expression\n[log2(CPM)]") +
  theme(legend.position = "bottom")
```

```{r}
umap_grid_cd14 <- plot_grid(umap_cell_type, 
                            umap_cd14_expr, 
                            ncol = 2)
```

```{r fig.width=20, fig.height=10}
umap_grid_cd14
```

```{r}
ggsave(here::here("figures/umap_grid_cd14_expression_scRNAseq.pdf"), 
       umap_grid_cd14, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_cd14_expression_scRNAseq.png"), 
       umap_grid_cd14, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

### Violin plot

```{r}
annotated_with_cd144violin <- annotated_with_cd14 %>%
  filter(cd14_expression > 0) 
```

```{r}
violin_cd14_expr_celltype <- ggplot(annotated_with_cd144violin, 
       aes(x = cluster, 
           y = cd14_expression,
           fill = cluster)) +
  geom_violin() +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "Cell type", 
       y = "CD14 expression [log2(CPM)]") +
  theme(legend.position = "none",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6)
        ) 
```

```{r}
#violin_cd14_expr_celltype
```

```{r}
umap_grid_w_violin_cd14 <- plot_grid(umap_cell_type, 
                                     umap_cd14_expr, 
                                     violin_cd14_expr_celltype,
                                     ncol = 2)
```

```{r fig.width=20, fig.height=20}
umap_grid_w_violin_cd14
```

```{r}
ggsave(here::here("figures/umap_grid_cd14_expression_violin_scRNAseq.pdf"), 
       umap_grid_w_violin_cd14, 
       width = 34, 
       height = 34,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_cd14_expression_violin_scRNAseq.png"), 
       umap_grid_w_violin_cd14, 
       width = 34, 
       height = 34,
       units = "cm")
```

## iGBM

```{r}
annotated_with_cd14_nd <- readRDS(here::here("rds/long_cpm_scrnaseq_cd14_annot_nd.rds"))
```

```{r}
umap_cd14_expr_nd <- ggplot(annotated_with_cd14_nd, 
       aes(x = x, 
           y = y, 
           color = cd14_expression)) +
  rasterize(geom_point(alpha = 0.5)) +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "CD14 expression\n[log2(CPM)]",
       title = "Newly diagnosed GBM") +
  theme(legend.position = "bottom")
```

```{r}
umap_grid_cd14_nd <- plot_grid(umap_cell_type_nd, 
                               umap_cd14_expr_nd, 
                               ncol = 2)
```

```{r fig.width=20, fig.height=10}
umap_grid_cd14_nd
```

```{r}
ggsave(here::here("figures/umap_grid_cd14_expression_scRNAseq_nd.pdf"), 
       umap_grid_cd14_nd, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_cd14_expression_scRNAseq_nd.png"), 
       umap_grid_cd14_nd, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

### Violin plot

```{r}
annotated_with_cd144violin_nd <- annotated_with_cd14_nd %>%
  filter(cd14_expression > 0) 
```

```{r}
violin_cd14_expr_celltype_nd <- ggplot(annotated_with_cd144violin_nd, 
       aes(x = cluster, 
           y = cd14_expression,
           fill = cluster)) +
  geom_violin() +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "Cell type", 
       y = "CD14 expression [log2(CPM)]") +
  theme(legend.position = "none",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6)
        ) 
```

```{r}
#violin_cd14_expr_celltype_nd
```

```{r}
umap_grid_w_violin_cd14_nd <- plot_grid(umap_cell_type, 
                                        umap_cd14_expr_nd, 
                                        violin_cd14_expr_celltype_nd,
                                        ncol = 2)
```

```{r fig.width=20, fig.height=20}
umap_grid_w_violin_cd14_nd
```

```{r}
ggsave(here::here("figures/umap_grid_cd14_expression_violin_scRNAseq_nd.pdf"), 
       umap_grid_w_violin_cd14_nd, 
       width = 34, 
       height = 34,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_cd14_expression_violin_scRNAseq_nd.png"), 
       umap_grid_w_violin_cd14_nd, 
       width = 34, 
       height = 34,
       units = "cm")
```

# Generate scatter plot for IBA1 expression

## rGBM

```{r}
annotated_with_iba1 <- readRDS(here::here("rds/long_cpm_scrnaseq_iba1_annot.rds"))
```

```{r}
umap_iba1_expr <- ggplot(annotated_with_iba1, 
       aes(x = x, 
           y = y, 
           color = iba1_expression)) +
  geom_point(alpha = 0.5) +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "IBA1 expression\n[log2(CPM)]") +
  theme(legend.position = "bottom")
```

```{r}
umap_grid_iba1 <- plot_grid(umap_cell_type, 
                            umap_iba1_expr, 
                            ncol = 2)
```

```{r}
ggsave(here::here("figures/umap_grid_iba1_expression_scRNAseq.pdf"), 
       umap_grid_iba1, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_iba1_expression_scRNAseq.png"), 
       umap_grid_iba1, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

### Violin plot

```{r}
annotated_with_iba14violin <- annotated_with_iba1 %>%
  filter(iba1_expression > 0)

  head(annotated_with_iba14violin)
```

```{r}
violin_iba1_expr_celltype <- ggplot(annotated_with_iba14violin, 
       aes(x = cluster, 
           y = iba1_expression,
           fill = cluster)) +
  geom_violin() +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "Cell type", 
       y = "IBA1 expression [log2(CPM)]") +
  theme(legend.position = "none",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6)
        ) 
```

```{r}
umap_grid_w_violin_iba1 <- plot_grid(umap_cell_type, 
                                     umap_iba1_expr, 
                                     violin_iba1_expr_celltype,
                                     ncol = 2)
```

```{r fig.width=20, fig.height=20}
umap_grid_w_violin_iba1
```

```{r}
ggsave(here::here("figures/umap_grid_iba1_expression_violin_scRNAseq.pdf"), 
       umap_grid_w_violin_iba1, 
       width = 34, 
       height = 34,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_iba1_expression_violin_scRNAseq.png"), 
       umap_grid_w_violin_iba1, 
       width = 34, 
       height = 34,
       units = "cm")
```

## iGBM

```{r}
annotated_with_iba1_nd <- readRDS(here::here("rds/long_cpm_scrnaseq_iba1_annot_nd.rds"))
```

```{r}
umap_iba1_expr_nd <- ggplot(annotated_with_iba1_nd, 
       aes(x = x, 
           y = y, 
           color = iba1_expression)) +
  rasterize(geom_point(alpha = 0.5)) +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "IBA1 expression\n[log2(CPM)]",
       title = "Newly diagnosed GBM") +
  theme(legend.position = "bottom")
```

```{r}
umap_grid_iba1_nd <- plot_grid(umap_cell_type_nd, 
                               umap_iba1_expr_nd, 
                               ncol = 2)
```

```{r fig.width=20, fig.height=10}
umap_grid_iba1_nd
```

```{r}
ggsave(here::here("figures/umap_grid_iba1_expression_scRNAseq_nd.pdf"), 
       umap_grid_iba1_nd, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_iba1_expression_scRNAseq_nd.png"), 
       umap_grid_iba1_nd, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

# Generate scatter plot for ELANE expression

## rGBM

```{r}
annotated_with_elane <- readRDS(here::here("rds/long_cpm_scrnaseq_elane_annot.rds"))
```

```{r}
umap_elane_expr <- ggplot(annotated_with_elane, 
       aes(x = x, 
           y = y, 
           color = elane_expression)) +
  geom_point(alpha = 0.5) +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "ELANE expression\n[log2(CPM)]") +
  theme(legend.position = "bottom")
```

```{r}
umap_grid_elane <- plot_grid(umap_cell_type, 
                             umap_elane_expr, 
                             ncol = 2)
```

```{r fig.width=20, fig.height=10}
#umap_grid_elane
```

```{r}
ggsave(here::here("figures/umap_grid_elane_expression_scRNAseq.pdf"), 
       umap_grid_elane, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_elane_expression_scRNAseq.png"), 
       umap_grid_elane, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

## Violin plot

```{r}
annotated_with_elane4violin <- annotated_with_elane %>%
  filter(elane_expression > 0) 
```

```{r}
violin_elane_expr_celltype <- ggplot(annotated_with_elane4violin, 
       aes(x = cluster, 
           y = elane_expression,
           fill = cluster)) +
  geom_violin() +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "Cell type", 
       y = "ELANE expression [log2(CPM)]") +
  theme(legend.position = "none",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6)
        ) 
```

```{r}
violin_elane_expr_celltype
```

```{r}
umap_grid_w_violin_elane <- plot_grid(umap_cell_type, 
                                      umap_elane_expr, 
                                      violin_elane_expr_celltype,
                                      ncol = 2)
```

```{r fig.width=20, fig.height=20}
umap_grid_w_violin_elane
```

```{r}
ggsave(here::here("figures/umap_grid_elane_expression_violin_scRNAseq.pdf"), 
       umap_grid_w_violin_elane, 
       width = 34, 
       height = 34,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_elane_expression_violin_scRNAseq.png"), 
       umap_grid_w_violin_elane, 
       width = 34, 
       height = 34,
       units = "cm")
```

## iGBM

```{r}
annotated_with_elane_nd <- readRDS(here::here("rds/long_cpm_scrnaseq_elane_annot_nd.rds"))
```

```{r}
umap_elane_expr_nd <- ggplot(annotated_with_elane_nd, 
       aes(x = x, 
           y = y, 
           color = elane_expression)) +
  rasterize(geom_point(alpha = 0.5)) +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "ELANE expression\n[log2(CPM)]",
       title = "Newly diagnosed GBM") +
  theme(legend.position = "bottom")
```

```{r}
umap_grid_elane_nd <- plot_grid(umap_cell_type_nd, 
                                umap_elane_expr_nd, 
                                ncol = 2)
```

```{r fig.width=20, fig.height=10}
umap_grid_elane_nd
```

```{r}
ggsave(here::here("figures/umap_grid_elane_expression_scRNAseq_nd.pdf"), 
       umap_grid_elane_nd, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_elane_expression_scRNAseq_nd.png"), 
       umap_grid_elane_nd, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

# Genearate scatter plot for lipocalin-2 (LCN2) expression

The single-cell dataset does not seem to contain LCN2 expression data.

# Overlay ELANE with ASAH1, MPO and CD14 expression

## rGBM

```{r}
annotated_with_elane_pos <- annotated_with_elane %>%
       mutate(elane_positive = if_else(elane_expression > 0, 
                                       "ELANE +", 
                                       NA)) %>% 
       dplyr::select(-c(gene))

annotated_with_asah1_elane <- left_join(annotated_with_asah1,
                                        annotated_with_elane_pos)

annotated_with_asah1_elane_only_elane <- annotated_with_asah1_elane %>%
       filter(!is.na(elane_positive))

annotated_with_mpo_only <- annotated_with_mpo %>%
       filter(mpo_expression > 0)

annotated_with_cd14_only <- annotated_with_cd14 %>%
       filter(cd14_expression > 0)

```

```{r}
```

```{r}
umap_asah1_elane_expr <- ggplot() +
  rasterize(geom_point(data = annotated_with_asah1_elane, 
       mapping = aes(x = x, 
                     y = y,
                     color = asah1_expression),
       alpha = 0.5
           )) +
  scale_color_gradient(low = "grey", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "ASAH1 expression\n[log2(CPM)]") +
  new_scale_color() + 
  rasterize(geom_point(data = annotated_with_asah1_elane_only_elane, 
       mapping = aes(x = x, 
                     y = y,
                     color = elane_expression),
       alpha = 0.5
           )) +
  scale_color_gradient(low = "grey", high = "green") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "ELANE expression\n[log2(CPM)]") +
  new_scale_color() + 
  rasterize(geom_point(data = annotated_with_mpo_only, 
       mapping = aes(x = x, 
                     y = y,
                     color = mpo_expression),
       alpha = 0.5
           )) +
  scale_color_gradient(low = "grey", high = "blue") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "MPO expression\n[log2(CPM)]",
       title = "Recurrent GBM") +
  geom_text(data = cell_type_centroid, 
            aes(x = x, 
                y = y, 
                label = cluster), 
            size = 5,
            color = "black") 
  #theme(legend.position = "bottom")
```

```{r}
umap_asah1_elane_expr
```

```{r}
umap_grid_overlay_three <- plot_grid(umap_cell_type, 
                                     umap_asah1_elane_expr, 
                                     ncol = 2)
```

```{r}
ggsave(here::here("figures/umap_grid_elane_asah1_mpo_expression_scRNAseq.pdf"), 
       umap_grid_overlay_three, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_elane_asah1_mpo_expression_scRNAseq.png"), 
       umap_grid_overlay_three, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r fig.width=20, fig.height=10}
umap_grid_overlay_three
```

### Violin of overlay

```{r}
annotated_with_elane4violin2 <- dplyr::rename(annotated_with_elane4violin,
                                              gene_expression = elane_expression)
```

```{r}
annotated_with_asah14violin2 <- dplyr::rename(annotated_with_asah14violin,
                                              gene_expression = asah1_expression)
```

```{r}
annotated_with_mpo4violin2 <- dplyr::rename(annotated_with_mpo4violin,
                                            gene_expression = mpo_expression)
```

```{r}
annotated_with_asah1_elane_mpo <- bind_rows(
       annotated_with_elane4violin2,
       annotated_with_asah14violin2,
       annotated_with_mpo4violin2
       ) %>%
       group_by(cluster, gene) %>%
       filter(n() > 1) %>%
       ungroup()
```

```{r}
#View(annotated_with_asah1_elane_mpo)
```

```{r}
violin_three_expr_celltype <- ggplot(annotated_with_asah1_elane_mpo, 
       aes(x = cluster, 
           y = gene_expression,
           fill = cluster)) +
  geom_violin() +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "Cell type", 
       y = "Gene expression [log2(CPM)]") +
  facet_grid(.~gene,
              scale = "free_x") +
  theme(legend.position = "none",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6)
        ) 
```

```{r}
violin_three_expr_celltype
```

```{r}
ggsave(here::here("figures/violin_plots_elane_asah1_mpo_expression_scRNAseq.pdf"), 
       plot = violin_three_expr_celltype, 
       device = "pdf",
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r}
dev.off()
```

## iGBM

```{r}
annotated_with_elane_nd <- readRDS(here::here("rds/long_cpm_scrnaseq_elane_annot_nd.rds"))
```

```{r}
annotated_with_mpo_nd <- readRDS(here::here("rds/long_cpm_scrnaseq_mpo_annot_nd.rds"))
```

```{r}
annotated_with_elane_pos_nd <- annotated_with_elane_nd %>%
       mutate(elane_positive = if_else(elane_expression > 0, 
                                       "ELANE +", 
                                       NA)) %>% 
       dplyr::select(-c(gene))

annotated_with_asah1_elane_nd <- left_join(annotated_with_asah1_nd,
                                           annotated_with_elane_pos_nd)

annotated_with_asah1_elane_only_elane_nd <- annotated_with_asah1_elane_nd %>%
       filter(!is.na(elane_positive))

annotated_with_mpo_only_nd <- annotated_with_mpo_nd %>% 
       filter(mpo_expression > 0)

       annotated_with_mpo_only_nd
```

```{r}
umap_asah1_elane_expr_nd <- ggplot() +
  rasterize(geom_point(data = annotated_with_asah1_elane_nd, 
       mapping = aes(x = x, 
                     y = y,
                     color = asah1_expression),
       alpha = 0.5
           )) +
  scale_color_gradient(low = "grey", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "ASAH1 expression\n[log2(CPM)]") +
  new_scale_color() + 
  rasterize(geom_point(data = annotated_with_asah1_elane_only_elane_nd, 
       mapping = aes(x = x, 
                     y = y,
                     color = elane_expression),
       alpha = 0.5
           )) +
  scale_color_gradient(low = "grey", high = "green") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "ELANE expression\n[log2(CPM)]") +
       new_scale_color() + 
  rasterize(geom_point(data = annotated_with_mpo_only_nd, 
       mapping = aes(x = x, 
                     y = y,
                     color = mpo_expression),
       alpha = 0.5
           )) +
  geom_text(data = cell_type_centroid_nd, 
            aes(x = x, 
                y = y, 
                label = cluster), 
            size = 5,
            color = "black") +
  scale_color_gradient(low = "grey", high = "blue") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "MPO expression\n[log2(CPM)]",
       title = "Newly diagnosed GBM") 
  #theme(legend.position = "bottom")
```

```{r}
umap_asah1_elane_expr_nd
```

## UMAP overlay of rGBM and iGBM

```{r}
umap_grid_overlay_three_nd <- plot_grid(umap_asah1_elane_expr_nd,
                                        umap_asah1_elane_expr, 
                                        ncol = 2)
```

```{r}
umap_grid_overlay_three_nd
```

```{r}
ggsave(here::here("figures/umap_grid_elane_asah1_mpo_expression_scRNAseq_nd.pdf"), 
       umap_grid_overlay_three_nd, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_elane_asah1_mpo_expression_scRNAseq_only_nd.pdf"), 
       umap_asah1_elane_expr_nd, 
       width = 34/2, 
       height = 34/2,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_elane_asah1_mpo_expression_scRNAseq_only_rgbm.pdf"), 
       umap_asah1_elane_expr, 
       width = 25, 
       height = 25,
       units = "cm")
```

# Generate scatter plot for CD68 expression

## rGBM

```{r}
annotated_with_cd68 <- readRDS(here::here("rds/long_cpm_scrnaseq_cd68_annot.rds"))
```

```{r}
umap_cd68_expr <- ggplot(annotated_with_cd68, 
       aes(x = x, 
           y = y, 
           color = cd68_expression)) +
  geom_point(alpha = 0.5) +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "CD68 expression\n[log2(CPM)]") +
  theme(legend.position = "bottom")
```

```{r}
umap_grid_cd68 <- plot_grid(umap_cell_type, 
                            umap_cd68_expr, 
                            ncol = 2)
```

```{r fig.width=20, fig.height=10}
#umap_grid_cd68
```

```{r}
ggsave(here::here("figures/umap_grid_cd68_expression_scRNAseq.pdf"), 
       umap_grid_cd68, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_cd68_expression_scRNAseq.png"), 
       umap_grid_cd68, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

## Violin plot

```{r}
annotated_with_cd684violin <- annotated_with_cd68 %>%
  filter(cd68_expression > 0) 
```

```{r}
violin_cd68_expr_celltype <- ggplot(annotated_with_cd684violin, 
       aes(x = cluster, 
           y = cd68_expression,
           fill = cluster)) +
  geom_violin() +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "Cell type", 
       y = "CD68 expression [log2(CPM)]") +
  theme(legend.position = "none",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6)
        ) 
```

```{r}
#violin_cd68_expr_celltype
```

```{r}
umap_grid_w_violin_cd68 <- plot_grid(umap_cell_type, 
                                     umap_cd68_expr, 
                                     violin_cd68_expr_celltype,
                                     ncol = 2)
```

```{r fig.width=20, fig.height=20}
umap_grid_w_violin_cd68
```

```{r}
ggsave(here::here("figures/umap_grid_cd68_expression_violin_scRNAseq.pdf"), 
       umap_grid_w_violin_cd68, 
       width = 34, 
       height = 34,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_cd68_expression_violin_scRNAseq.png"), 
       umap_grid_w_violin_cd68, 
       width = 34, 
       height = 34,
       units = "cm")
```

## iGBM

```{r}
annotated_with_cd68_nd <- readRDS(here::here("rds/long_cpm_scrnaseq_cd68_annot_nd.rds"))
```

```{r}
umap_cd68_expr_nd <- ggplot(annotated_with_cd68_nd, 
       aes(x = x, 
           y = y, 
           color = cd68_expression)) +
  geom_point(alpha = 0.5) +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "CD68 expression\n[log2(CPM)]",
       title = "Newly diagnosed GBM") +
  theme(legend.position = "bottom")
```

```{r}
umap_grid_cd68_nd <- plot_grid(umap_cell_type_nd, 
                               umap_cd68_expr_nd, 
                               ncol = 2)
```

```{r fig.width=20, fig.height=10}
umap_grid_cd68_nd
```

```{r}
ggsave(here::here("figures/umap_grid_cd68_expression_scRNAseq_nd.pdf"), 
       umap_grid_cd68_nd, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_cd68_expression_scRNAseq_nd.png"), 
       umap_grid_cd68_nd, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

### Violin plot

```{r}
annotated_with_cd684violin_nd <- annotated_with_cd68_nd %>%
  filter(cd68_expression > 0) 
```

```{r}
violin_cd68_expr_celltype_nd <- ggplot(annotated_with_cd684violin_nd, 
       aes(x = cluster, 
           y = cd68_expression,
           fill = cluster)) +
  geom_violin() +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "Cell type", 
       y = "CD68 expression [log2(CPM)]") +
  theme(legend.position = "none",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6)
        ) 
```

```{r}
#violin_cd68_expr_celltype_nd
```

```{r}
umap_grid_w_violin_cd68_nd <- plot_grid(umap_cell_type_nd, 
                                        umap_cd68_expr_nd, 
                                        violin_cd68_expr_celltype_nd,
                                        ncol = 2)
```

```{r fig.width=20, fig.height=20}
umap_grid_w_violin_cd68_nd
```

```{r}
ggsave(here::here("figures/umap_grid_cd68_expression_violin_scRNAseq_nd.pdf"), 
       umap_grid_w_violin_cd68_nd, 
       width = 34, 
       height = 34,
       units = "cm")
```

```{r}
ggsave(here::here("figures/umap_grid_cd68_expression_violin_scRNAseq_nd.png"), 
       umap_grid_w_violin_cd68_nd, 
       width = 34, 
       height = 34,
       units = "cm")
```

# Overlay IBA1 and CD68 expression

## rGBM

```{r}
annotated_with_cd68_pos <- annotated_with_cd68 %>%
       mutate(cd68_positive = if_else(cd68_expression > 0, 
                                      "CD68 +", 
                                      NA)) %>% 
       dplyr::select(-c(gene))

annotated_with_iba1_cd68 <- left_join(annotated_with_iba1,
                                      annotated_with_cd68_pos)

annotated_with_iba1_cd68_only_cd68 <- annotated_with_iba1_cd68 %>%
       filter(!is.na(cd68_positive))

annotated_with_iba1_only <- annotated_with_iba1 %>%
       filter(iba1_expression > 0) 
```

```{r}
umap_iba1_cd68_expr3 <- ggplot() +
  rasterize(
  geom_point(data = annotated_with_iba1_cd68, 
       mapping = aes(x = x, 
                     y = y,
                     color = iba1_expression),
       alpha = 0.2,
       shape = 16
           ),
           dpi = 300
           ) +
  scale_color_gradient(low = "grey", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "IBA1 expression\n[log2(CPM)]") +
  new_scale_color() + 
  rasterize(
  geom_point(data = annotated_with_iba1_cd68_only_cd68, 
       mapping = aes(x = x, 
                     y = y,
                     color = cd68_expression),
       alpha = 0.09,
       shape = 16
           ),
           dpi = 300
           ) +
  scale_color_gradient(low = "grey", high = "green") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "CD68 expression\n[log2(CPM)]",
       title = "Recurrent GBM")
```

```{r}
umap_iba1_cd68_expr3
```

```{r}
ggsave(here::here("figures/umap_grid_iba1_cd68_expression_scRNAseq.pdf"), 
       umap_iba1_cd68_expr3, 
       width = 34/2, 
       height = 34/2,
       units = "cm")
```

### Violin of overlay

```{r}
annotated_with_cd684violin2 <- dplyr::rename(annotated_with_cd684violin,
                                              gene_expression = cd68_expression)
```

```{r}
annotated_with_iba14violin2 <- dplyr::rename(annotated_with_iba14violin,
                                              gene_expression = iba1_expression)
```

```{r}
annotated_with_iba1_cd68 <- bind_rows(
       annotated_with_cd684violin2,
       annotated_with_iba14violin2
       ) %>%
       group_by(cluster, gene) %>%
       filter(n() > 1) %>%
       ungroup()
```

```{r}
#View(annotated_with_iba1_cd68)
```

```{r}
violin_two_expr_celltype <- ggplot(annotated_with_iba1_cd68, 
       aes(x = cluster, 
           y = gene_expression,
           fill = cluster)) +
  geom_violin() +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "Cell type", 
       y = "Gene expression [log2(CPM)]") +
  facet_grid(.~gene,
              scale = "free_x") +
  theme(legend.position = "none",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6)
        ) 
```

```{r}
ggsave(here::here("figures/violin_plots_iba1_cd68_expression_scRNAseq.pdf"), 
       plot = violin_two_expr_celltype, 
       device = "pdf",
       width = 2*(34/3), 
       height = 34/2,
       units = "cm")
```

## iGBM

```{r}
annotated_with_cd68_nd <- readRDS(here::here("rds/long_cpm_scrnaseq_cd68_annot_nd.rds"))
```

```{r}
annotated_with_iba1_nd <- readRDS(here::here("rds/long_cpm_scrnaseq_iba1_annot_nd.rds"))
```

```{r}
annotated_with_cd68_pos_nd <- annotated_with_cd68_nd %>%
       mutate(cd68_positive = if_else(cd68_expression > 0, 
                                      "CD68 +", 
                                      NA)) %>% 
       dplyr::select(-c(gene))

annotated_with_iba1_cd68_nd <- left_join(annotated_with_iba1_nd,
                                         annotated_with_cd68_pos_nd)

annotated_with_iba1_cd68_only_cd68_nd <- annotated_with_iba1_cd68_nd %>%
       filter(!is.na(cd68_positive))

annotated_with_iba1_only_nd <- annotated_with_iba1_nd %>%
       filter(iba1_expression > 0) 
```

```{r}
umap_iba1_cd68_expr3_nd <- ggplot() +
  rasterize(
  geom_point(data = annotated_with_iba1_cd68_nd, 
       mapping = aes(x = x, 
                     y = y,
                     color = iba1_expression),
       alpha = 0.2,
       shape = 16
           ),
           dpi = 300
           ) +
  scale_color_gradient(low = "grey", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "IBA1 expression\n[log2(CPM)]") +
  new_scale_color() + 
  rasterize(
  geom_point(data = annotated_with_iba1_cd68_only_cd68_nd, 
       mapping = aes(x = x, 
                     y = y,
                     color = cd68_expression),
       alpha = 0.09,
       shape = 16
           ),
           dpi = 300
           ) +
  scale_color_gradient(low = "grey", high = "green") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "CD68 expression\n[log2(CPM)]",
       title = "Newly diagnosed GBM")
```

```{r}
umap_iba1_cd68_expr3_nd
```

```{r}
ggsave(here::here("figures/umap_grid_iba1_cd68_expression_scRNAseq_nd.pdf"), 
       umap_iba1_cd68_expr3_nd, 
       width = 34/2, 
       height = 34/2,
       units = "cm")
```

```{r}
umap_iba1_cd68_both <- plot_grid(umap_iba1_cd68_expr3_nd,
                                 umap_iba1_cd68_expr3)
```

```{r}
ggsave(here::here("figures/umap_grid_iba1_cd68_expression_scRNAseq_both.pdf"), 
       umap_iba1_cd68_both, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

## Violin plot of the overlay

```{r}
annot_w_iba1_CD68_igbm_n_rgbm <- bind_rows(
       annotated_with_cd68 %>%
       mutate(gbm_type = "rGBM",
              expression = cd68_expression),
       annotated_with_cd68_nd %>%
       mutate(gbm_type = "iGBM",
              expression = cd68_expression),
       annotated_with_iba1 %>%
       mutate(gbm_type = "rGBM",
              expression = iba1_expression),
       annotated_with_iba1_nd %>%
       mutate(gbm_type = "iGBM",
              expression = iba1_expression)
              ) %>%
                                        mutate(cluster = if_else(
                                          cluster == "DC 1", 
                                          "DC", 
                                          cluster)) %>%
                                        filter(
                                          !cluster %in% c(
                                                 "DC 2",
                                                 "DC 3",
                                                 "DC 4",
                                                 "Mast cells",
                                                 "Plasma B",
                                                 "Regulatory T cells"
                                          )
                                        ) %>%
              filter(expression > 5)

unique(annot_w_iba1_CD68_igbm_n_rgbm$cluster)
```

```{r}
violin_two_expr_celltype_igbm_n_rgbm <- ggplot(annot_w_iba1_CD68_igbm_n_rgbm, 
       aes(x = cluster, 
           y = expression,
           fill = gbm_type)) +
  geom_violin() +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "Cell type", 
       y = "Gene expression [log2(CPM)]") +
  facet_grid(.~gene,
              scale = "free_x") +
  theme(legend.position = "bottom",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6)
        ) 
```

```{r}
violin_two_expr_celltype_igbm_n_rgbm
```

```{r}
ggsave(here::here("figures/violin_plots_iba1_cd68_expression_scRNAseq_igbm_n_rgbm.pdf"), 
       plot = violin_two_expr_celltype_igbm_n_rgbm, 
       device = "pdf",
       width = 2*(34/3), 
       height = 34/3,
       units = "cm")
```

```{r}
ggsave(here::here("figures/violin_plots_iba1_cd68_expression_scRNAseq_igbm_n_rgbm.png"), 
       plot = violin_two_expr_celltype_igbm_n_rgbm, 
       device = "png",
       width = 2*(34/3), 
       height = 34/3,
       units = "cm")
```

```{r}
dev.off()
```

# rGBM vs iGBM comparisons

## ASAH1 expression

Load the data

```{r}
annotated_with_asah1_nd <- readRDS(here::here("rds/long_cpm_scrnaseq_asah1_annot_nd.rds"))
```

### UMAP of cell type based on iGBM (ND = newly diagnosed)

```{r}
cell_type_centroid_nd <- annotated_with_asah1_nd %>%
  group_by(cluster) %>%
  summarise(x = mean(x), 
            y = mean(y)) %>%
  ungroup()
```

```{r}
umap_cell_type_nd <- ggplot(annotated_with_asah1_nd, 
       aes(x = x, 
           y = y, 
           color = cluster)) +
  ggrastr::rasterize(geom_point(alpha = 0.5)) +
  geom_text(data = cell_type_centroid_nd, 
            aes(x = x, 
                y = y, 
                label = cluster), 
            size = 5,
            color = "black") +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "Cell type",
       title = "Newly diagnosed GBM") +
  theme(legend.position = "bottom") 
```

# UMAPs of cell type between iGBM and rGBM

```{r}
umap_grid_wnd <- plot_grid(umap_cell_type, 
                           umap_cell_type_nd, 
                           ncol = 2)
```

```{r}
umap_grid_wnd
```

```{r}
ggsave(here::here("figures/umap_grid_cell_type_iGBM_n_rGBM_scRNAseq.pdf"), 
       umap_grid_wnd, 
       width = 34, 
       height = 34/2,
       units = "cm")
```

## UMAPs with ASAH1 expression between iGBM and rGBM

```{r}
umap_asah1_expr_nd <- ggplot(annotated_with_asah1_nd, 
       aes(x = x, 
           y = y, 
           color = asah1_expression)) +
  ggrastr::rasterize(geom_point(alpha = 0.5)) +
  geom_text(data = cell_type_centroid_nd, 
            aes(x = x, 
                y = y, 
                label = cluster), 
            size = 5,
            color = "black") +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "UMAP_1", 
       y = "UMAP_2", 
       color = "ASAH1 expression\n[log2(CPM)]",
       title = "Newly diagnosed GBM") +
  theme(legend.position = "bottom")
```

```{r}
umap_grid_asah1_nd <- plot_grid(umap_asah1_expr, 
                                umap_asah1_expr_nd, 
                                ncol = 2)
```

```{r}
umap_grid_asah1_nd
```

## Violin plots of ASAH1 expression between iGBM and rGBM

```{r}
# merge tables from annotated abundances in iGBM and rGBM
#View(annotated_with_asah1_nd)
#View(annotated_with_asah1)

str(annotated_with_asah1)
str(annotated_with_asah1_nd)
```

```{r}
annot_w_asah1_igbm_n_rgbm <- bind_rows(annotated_with_asah1 %>%
                                        mutate(gbm_type = "rGBM"),
                                       annotated_with_asah1_nd %>%
                                        mutate(gbm_type = "iGBM")) %>%
                                        mutate(cluster = if_else(
                                          cluster == "DC 1", 
                                          "DC", 
                                          cluster)) %>%
                                        filter(
                                          !cluster %in% c(
                                                 "DC 2",
                                                 "DC 3",
                                                 "DC 4",
                                                 "Mast cells",
                                                 "Plasma B",
                                                 "Regulatory T cells"
                                          )
                                        )
```

```{r}
annot_w_asah1_igbm_n_rgbm_1 <- annot_w_asah1_igbm_n_rgbm %>%
  filter(asah1_expression > 2.5)
```
```{r}
violin_asahl_expr_celltype <- ggplot(annot_w_asah1_igbm_n_rgbm_1, 
       aes(x = cluster, 
           y = asah1_expression,
           fill = gbm_type)) +
  geom_violin() +
  #scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "Cell type", 
       y = "ASAH1 expression [log2(CPM)]") +
  theme(legend.position = "bottom",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6)
        ) 
```

```{r}
violin_asahl_expr_celltype
```

```{r}
# fit a linear model to compare ASAH1 expression between iGBM and rGBM
# within each cluster

#View(annot_w_asah1_igbm_n_rgbm)

```

## Violin plots for ASAH1, ELANE and MPO expression between iGBM and rGBM

```{r}
#head(annot_w_asah1_elane_mpo_igbm_n_rgbm)

annot_w_asah1_elane_mpo_igbm_n_rgbm <- bind_rows(
       annotated_with_asah1 %>%
       mutate(gbm_type = "rGBM",
              expression = asah1_expression),
       annotated_with_asah1_nd %>%
       mutate(gbm_type = "iGBM",
              expression = asah1_expression),
       annotated_with_elane %>%
       mutate(gbm_type = "rGBM",
              expression = elane_expression),
       annotated_with_elane_nd %>%
       mutate(gbm_type = "iGBM",
              expression = elane_expression),
       annotated_with_mpo %>%
       mutate(gbm_type = "rGBM",
              expression = mpo_expression),
       annotated_with_mpo_nd %>%
       mutate(gbm_type = "iGBM",
              expression = mpo_expression)
              ) %>%
                                        mutate(cluster = if_else(
                                          cluster == "DC 1", 
                                          "DC", 
                                          cluster)) %>%
                                        filter(
                                          !cluster %in% c(
                                                 "DC 2",
                                                 "DC 3",
                                                 "DC 4",
                                                 "Mast cells",
                                                 "Plasma B",
                                                 "Regulatory T cells"
                                          )
                                        )

unique(annot_w_asah1_elane_mpo_igbm_n_rgbm$cluster)
```

```{r}
annot_w_asah1_elane_mpo_igbm_n_rgbm1 <- annot_w_asah1_elane_mpo_igbm_n_rgbm %>%
  filter(expression > 2.5)
```

```{r}
violin_three_expr_celltype <- ggplot(annot_w_asah1_elane_mpo_igbm_n_rgbm1, 
       aes(x = cluster, 
           y = expression,
           fill = gbm_type)) +
  geom_violin() +
  #scale_color_gradient(low = "yellow", high = "red") +
  facet_wrap(.~gene,
              scale = "free_x",
              ncol = 3) +
  labs(x = "Cell type", 
       y = "Gene expression [log2(CPM)]") +
  theme(legend.position = "bottom",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6)
        ) 
```

```{r}
violin_three_expr_celltype
```

```{r}
ggsave(here::here("figures/violin_plots_iba1_cd68_expression_scRNAseq_igbm_n_rgbm.pdf"), 
       plot = violin_three_expr_celltype, 
       device = "pdf",
       width = 2*(34/3), 
       height = 34/3,
       units = "cm")
```

## ASAH1 expression in CD14+ cells

### rGBM

Filter CD14+ cells:

```{r}
max(annotated_with_cd14$cd14_expression)

cd14_positive_rgbm <- annotated_with_cd14 %>%
  filter(cluster %in% c(
       "Monocytes",
       "prol. TAM",
       "TAM 1",
       "TAM 2"
  )) %>% 
  pull(cell)

annotated_with_asah1_cd14_posit_rgbm <- annotated_with_asah1 %>%
       filter(cell %in% c(cd14_positive_rgbm))
```

### iGBM

Filter CD14+ cells:

```{r}
max(annotated_with_cd14_nd$cd14_expression)

cd14_positive_igbm <- annotated_with_cd14_nd %>%
  filter(cluster %in% c(
       "Monocytes",
       "prol. TAM",
       "TAM 1",
       "TAM 2"
  )) %>% 
  pull(cell)

annotated_with_asah1_cd14_posit_igbm <- annotated_with_asah1_nd %>%
       filter(cell %in% c(cd14_positive_igbm)) 
```

### Violin plot of ASAH1 expression in CD14+ cells

```{r}
annotated_with_asah1_cd14_posit <- bind_rows(
       annotated_with_asah1_cd14_posit_rgbm %>%
       mutate(gbm_type = "rGBM"),
       annotated_with_asah1_cd14_posit_igbm %>%
       mutate(gbm_type = "iGBM")
       ) %>%
       group_by(cluster, gbm_type) %>%
       filter(n() > 1) %>%
       ungroup() %>% 
       filter(asah1_expression > 2.5) %>%
       mutate(cluster = if_else(
         cluster == "DC 1", 
         "DC", 
         cluster)) %>%
       filter(
         !cluster %in% c(
                "DC 2",
                "DC 3",
                "DC 4",
                "Mast cells",
                "Plasma B",
                "Regulatory T cells",
                "B cells",
                "DC",
                "NK cells",
                "T cells"
         )
       )

       annotated_with_asah1_cd14_posit
```

```{r}
violin_asahl_expr_celltype_cd14 <- ggplot(annotated_with_asah1_cd14_posit, 
       aes(x = cluster, 
           y = asah1_expression,
           fill = gbm_type)) +
  geom_violin(width = 0.6) +
  geom_boxplot(width = 0.4, ) +
  stat_compare_means(aes(group = gbm_type), 
                     label = "p.format") +
  labs(x = "Cell type", 
       y = "ASAH1 expression [log2(CPM)]") +
  theme(legend.position = "bottom",
        # tilt the x-axis labels 45 degrees
        axis.text.x = element_text(angle = -45, 
                                   hjust = 0,
                                   vjust = 1.6))
```

```{r}
dev.off()
```

```{r}
violin_asahl_expr_celltype_cd14
```

```{r}
ggsave(here::here("figures/violin_plots_asahl_expr_celltype_cd14_scRNAseq_igbm_n_rgbm.pdf"), 
       plot = violin_asahl_expr_celltype_cd14, 
       device = "pdf",
       width = (2*(34/3))/2, 
       height = 34/3,
       units = "cm")
```